### 2022-04-15 README.Rmd
Initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# si

<!-- badges: start -->
<!-- badges: end -->

The goal of this 'si' repo is to demonstrate Spatial Interaction models from first principles.
At present it just contains prose and reproducible code demonstrating how they work.
In future it may be turned into software, e.g. an R package, for others wanting to develop spatial interaction models.

### 2022-04-15 README.qmd
Initial commit

---
# format: html
format: gfm
number-sections: true
title: "si: code and prose exploring Spatial Interaction models"
bibliography: references.bib
execute:
  cache: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

<!-- badges: start -->

<!-- badges: end -->

The goal of this 'si' repo is to demonstrate Spatial Interaction models from first principles. At present it just contains prose and reproducible code demonstrating how they work. In future it may be turned into software, e.g. an R package, for others wanting to develop spatial interaction models.

# What are SIMs?

Spatial Interaction Models (SIMs) are mathematical models for estimating movement between spatial entities developed by Alan Wilson in the late 1960s and early 1970, with considerable uptake and refinement for transport modelling since then @boyce_forecasting_2015.
There are four main types of traditional SIMs [@wilson_family_1971]:

-   Unconstrained

-   Production-constrained

-   Attraction-constrained

-   Doubly-constrained

Traditional 'gravity model' based SIMs can be defined as follows, based on the original definition [@wilson_family_1971]:

$$
T_{i j}=K \frac{W_{i}^{(1)} W_{j}^{(2)}}{c_{i j}^{n}}
$$

"where $T_{i j}$ is a measure of the interaction between zones $i$ and $W_{i}^{(1)}$ is a measure of the 'mass term' associated with zone $z_i$, $W_{j}^{(2)}$ is a measure of the 'mass term' associated with zone $z_j$, and $c_{ij}$ is a measure of the distance, or generalised cost of travel, between zone $i$ and zone $j$".
$K$ is a 'constant of proportionality' and $n$ is a parameter to be estimated.

Redefining the $W$ terms as $m$ and $n$ for origins and destinations respectively [@simini_universal_2012], this classic definition of the 'gravity model' can be written as follows:

$$
T_{i j}=K \frac{m_{i} n_{j}}{c_{i j}^{n}}
$$

For the purposes of this project, we will focus on production-constrained SIMs.
These can be defined as follows [@wilson_family_1971]:

$$
T_{ij} = A_iO_in_jf(c_{ij})
$$

where $A$ is a balancing factor defined as:

$$
A_{i}=\frac{1}{\sum_{j} m_{j} \mathrm{f}\left(c_{i j}\right)}
$$

$O_i$ is analogous to the travel demand in zone $i$, which can be roughly approximated by its population.

<!-- $$ -->
<!-- T_{ij} = \frac{m_{i}^{\alpha} n_{j} ^{\beta}} {f(r_{ij})} -->
<!-- $$ -->

<!-- ## Radiation models -->

<!-- A more recent type of SIM is the radiation model @simini_universal_2012: -->

<!-- $$ -->
<!-- \langle T_{ij} \rangle = T_i \frac{m_{i} n_{j}} {(m_i + s_{ij})(m_i + n_j + s_{ij}) } -->
<!-- $$ -->
<!-- where $s_{ij}$ is defined as the total population living within a circle, the centre of which lies in the centroid of zone $i$ and the radius of which is the distance between zones $i$ and $j$. Thus, the greater the population living within the commute distance, the lower the estimated flow rate. -->

```{r}

```

# Implementation in R

```{r, message=FALSE}
library(tmap)
library(pct)
library(tidyverse)
```

```{r}
zones = get_pct_zones(region = "west-yorkshire", geography = "msoa")
centroids = get_pct_centroids(region = "west-yorkshire", geography = "msoa")
zones = zones %>% 
  filter(lad_name == "Leeds")
centroids = centroids %>% 
  filter(lad_name == "Leeds")
tm_shape(zones) + tm_polygons("all", palette = "viridis")
```

```{r}
od_df = od::points_to_od(centroids)
od_sfc = od::odc_to_sfc(od_df[3:6])
sf::st_crs(od_sfc) = 4326
od_df$length = sf::st_length(od_sfc)
od_df = od_df %>% transmute(
  O, D, length = as.numeric(length) / 1000,
  flow = NA, fc = NA
  )
od_df = sf::st_sf(od_df, geometry = od_sfc, crs = 4326)
```

An unconstrained spatial interaction model can be written as follows, with a more-or-less arbitrary value for `beta` which can be optimised later:

```{r}
#| label: unconstrained
beta = 0.3
i = 1
j = 2
for(i in seq(nrow(zones))) {
  for(j in seq(nrow(zones))) {
    O = zones$all[i]
    n = zones$all[j]
    ij = which(od_df$O == zones$geo_code[i] & od_df$D == zones$geo_code[j])
    od_df$fc[ij] = exp(-beta * od_df$length[ij])
    od_df$flow[ij] = O * n * od_df$fc[ij]
  }
}
od_top = od_df %>% 
  filter(O != D) %>% 
  top_n(n = 2000, wt = flow)

tm_shape(zones) +
  tm_borders() +
  tm_shape(od_top) +
  tm_lines("flow")
```

We can plot the 'distance decay' curve associated with this SIM is as follows:

```{r}
summary(od_df$fc)
od_df %>% 
  ggplot() +
  geom_point(aes(length, fc))
```

We can make this production constrained as follows:

```{r}
#| label: constrained
od_dfj = left_join(
  od_df,
  zones %>% select(O = geo_code, all) %>% sf::st_drop_geometry()
)
od_dfj = od_dfj %>% 
  group_by(O) %>% 
  mutate(flow_constrained = flow / sum(flow) * first(all)) %>%
  ungroup()
sum(od_dfj$flow_constrained) == sum(zones$all)
od_top = od_dfj %>% 
  filter(O != D) %>% 
  top_n(n = 2000, wt = flow_constrained)

tm_shape(zones) +
  tm_borders() +
  tm_shape(od_top) +
  tm_lines("flow_constrained")
```

# Validation

```{r validation}
od_census = get_od()
od_census_min = od_census %>% 
  rename(O = geo_code1, D = geo_code2) %>% 
  filter(O %in% zones$geo_code & D %in% zones$geo_code)
od_dfjc = inner_join(od_dfj %>% select(-all), od_census_min)
od_dfjc %>% 
  ggplot() +
  geom_point(aes(all, flow_constrained))
cor(od_dfjc$all, od_dfjc$flow_constrained)^2
```

# Implementation in Python

TBC.

# Resources

-   Reproducible code implementing SIMs in R (based on the outdated `sp` class system): https://rpubs.com/adam_dennett/257231

-

### 2022-04-15 R/si.R
Add example data
Former-commit-id: e60d3036374f63662fbe82d96ec3f3fa5509803c

@@ -1,0 +1,3 @@
+si = function(origins) {
+    
+}

### 2022-04-15 data-raw/zones.R

@@ -1,0 +1,17 @@
+## code to prepare `zones` dataset goes here
+library(pct)
+library(tidyverse)
+zones = get_pct_zones(region = "west-yorkshire", geography = "msoa")
+centroids = get_pct_centroids(region = "west-yorkshire", geography = "msoa")
+zones = zones %>% 
+  filter(lad_name == "Leeds")
+centroids = centroids %>% 
+  filter(lad_name == "Leeds")
+usethis::use_data(zones, overwrite = TRUE)
+usethis::use_data(centroids, overwrite = TRUE)
+
+od_census = get_od()
+od_census_min = od_census %>% 
+  rename(O = geo_code1, D = geo_code2) %>% 
+  filter(O %in% zones$geo_code & D %in% zones$geo_code)
+usethis::use_data(od_census_min)

### 2022-04-15 R/si.R
Update example data

@@ -1,3 +1,8 @@
-si = function(origins) {
-    
+#' Spatial Interactions
+#''
+#' Estimate movement between origins and destinations with a spatial interaction model
+#' @examples
+#' origins = zones
+si = function(origins, destinations, max_euclidean_distance = 20, fun = lm) {
+
 }

### 2022-04-15 data-raw/zones.R
Update example data
Former-commit-id: 04fc02a7f8c9f8d8bf82f4772b1ae1ff1a632c20

@@ -4,9 +4,11 @@ library(tidyverse)
 zones = get_pct_zones(region = "west-yorkshire", geography = "msoa")
 centroids = get_pct_centroids(region = "west-yorkshire", geography = "msoa")
 zones = zones %>% 
-  filter(lad_name == "Leeds")
+  filter(lad_name == "Leeds") %>%
+  select(1:13)
 centroids = centroids %>% 
-  filter(lad_name == "Leeds")
+  filter(lad_name == "Leeds") %>%
+  select(1:13)
 usethis::use_data(zones, overwrite = TRUE)
 usethis::use_data(centroids, overwrite = TRUE)
 

### 2022-04-15 R/si.R
our vscode test pass
Former-commit-id: 2d44d1798eed1688261a92df7d973c1a49f68a34

@@ -1,8 +1,58 @@
-#' Spatial Interactions
-#''
-#' Estimate movement between origins and destinations with a spatial interaction model
+#' Spatial Interaction Model
+#'
+#' Estimate movement between origins and destinations 
+#' with a spatial interaction model
+#'
+#' @param origins `sf` object representing origin locations/zones
+#' @param destinations `sf` object representing destination locations/zones
+#' @param max_euclidean_distance Maximum distance in meters between OD pairs.
+#'   Only OD pairs that are this distance apart or less will be returned.
+#' @export
 #' @examples
-#' origins = zones
-si = function(origins, destinations, max_euclidean_distance = 20, fun = lm) {
+#' origins = centroids
+#' destinations = centroids
+#' odsf = si_to_od(origins, destinations)
+#' plot(odsf)
+#' 
+si_to_od = function(origins, destinations, max_euclidean_distance = 10000) {
+  if(!identical(origins, destinations)) {
+        stop("Origins and destinations are different, not implemented yet")
+    }
+  od_df = od::points_to_od(origins)
+  od_df$distance_euclidean = geodist::geodist(
+      x = od_df[c("ox", "oy")],
+      y = od_df[c("dx", "dy")],
+      paired = TRUE
+      )
+  if(max(od_df$distance_euclidean) > max_euclidean_distance) {
+      nrow_before = nrow(od_df)
+      od_df = od_df[od_df$distance_euclidean <= max_euclidean_distance, ]
+      nrow_after = nrow(od_df)
+      pct_kept = round(nrow_after / nrow_before * 100)
+      message(
+          nrow_after,
+          " OD pairs remaining after removing those with a distance greater than ", # nolint
+          max_euclidean_distance, " meters", ":\n",
+          pct_kept, " of all possible OD pairs"
+          )
+  }
+  # join origin attributes
+  origins_to_join = sf::st_drop_geometry(origins)
+  names(origins_to_join) = paste0("origin_", names(origins_to_join))
+  names(origins_to_join)[1] = names(od_df)[1]
+  od_dfj = dplyr::inner_join(od_df, origins_to_join)
+  # join destination attributes
+  destinations_to_join = sf::st_drop_geometry(destinations)
+  names(destinations_to_join) = paste0("destination_", names(destinations_to_join)) # nolint
[…]

### 2022-04-15 R/si.R
si_model mvp

@@ -1,6 +1,7 @@
-#' Spatial Interaction Model
+#' Prepare OD
 #'
-#' Estimate movement between origins and destinations 
+#' Prepares an OD data frame that next could be used to
+#' estimate movement between origins and destinations 
 #' with a spatial interaction model
 #'
 #' @param origins `sf` object representing origin locations/zones
@@ -40,12 +41,12 @@ si_to_od = function(origins, destinations, max_euclidean_distance = 10000) {
   origins_to_join = sf::st_drop_geometry(origins)
   names(origins_to_join) = paste0("origin_", names(origins_to_join))
   names(origins_to_join)[1] = names(od_df)[1]
-  od_dfj = dplyr::inner_join(od_df, origins_to_join)
+  od_dfj = dplyr::inner_join(od_df, origins_to_join, by = "O")
   # join destination attributes
   destinations_to_join = sf::st_drop_geometry(destinations)
   names(destinations_to_join) = paste0("destination_", names(destinations_to_join)) # nolint
   names(destinations_to_join)[1] = names(od_df)[2]
-  od_dfj = dplyr::inner_join(od_dfj, destinations_to_join)
+  od_dfj = dplyr::inner_join(od_dfj, destinations_to_join, by = "D")
   names(od_dfj)
 
   # convert to sfc:
@@ -56,3 +57,39 @@ si_to_od = function(origins, destinations, max_euclidean_distance = 10000) {
   sf::st_sf(od_dfj, geometry = od_sfc)
 }
 
+#' Run a spatial interaction model
+#'
+#' Executes a spatial iteraction model based on an OD data frame 
+#' and user-specified function
+#' 
+#' @param od A data frame representing origin-destination data, e.g. as created by
+#'   [si_to_od()]
+#' @param fun A function that calculates the interaction (e.g. the number of trips)
+#'   between each OD pair
+#' @param var_p Use this argument to run a 'production constrained' SIM.
+#'   Character string corresponding to column in the `od`
+#'   dataset that constrains the total 'interaction' (e.g. n. trips) for all OD pairs
+#'   such that the total for each zone of origin cannot go above this value.
+#' @param ... Arguments passed to `fun`
+#' @export
+#' @examples
+#' od = si_to_od(zones, zones)
[…]

### 2022-04-15 R/si.R
Update with tests passing

@@ -83,7 +83,7 @@ si_to_od = function(origins, destinations, max_euclidean_distance = 10000) {
 #' plot(od$distance_euclidean, od_output$res)
 #' od_pconst = si_model(od, fun = fun, beta = 0.3, O = "origin_all", n = "destination_all", d = "distance_euclidean", var_p = origin_all)
 #' plot(od_pconst$distance_euclidean, od_pconst$res_constrained)
-#' plot(od_pconst["res_constrained"], logz = TRUE)
+#' plot(od_pconst["res"], logz = TRUE)
 si_model = function(od, fun, var_p, ...) {
     od$res = fun(od, ...)
     if (!missing(var_p)) {

### 2022-04-15 R/data.R
Document
Former-commit-id: 6e1bd768d8502a40ef4a6e8d9ed1e2affd01f447

@@ -1,0 +1,20 @@
+#' Example zones and centroids
+#'
+#' `zones` and `centroids` represent administrative zones between which
+#' flows are to be estimated.
+#' @note The schema data can be (re-)generated using code in `data-raw`
+#' @docType data
+#' @keywords datasets
+#' @name zones
+#' @aliases centroids
+NULL
+
+#' Example OD dataset
+#'
+#' Example OD dataset from the 2011 UK Census
+#' @note Regenerate the data with scripts in the `data-raw` directory.
+#'
+#' @docType data
+#' @keywords datasets
+#' @name od_census_min
+NULL

### 2022-04-15 README.qmd
updates readme

@@
 title: "si: code and prose exploring Spatial Interaction models"
-bibliography: references.bib
+bibliography: data-raw/references.bib
 execute:
@@
 ```{r, include=FALSE}
-knitr::opts_chunk$set(cache = TRUE)
+knitr::opts_chunk$set(cache = TRUE,
+  fig.path = "man/figures/README-")
 ```

### 2022-04-15 README.qmd
Update docs, tests will pass!

@@
 ```{r}
+#| label: inputs
 zones = get_pct_zones(region = "west-yorkshire", geography = "msoa")
@@
 ```{r}
+#| label: distance_decay
 summary(od_df$fc)

### 2022-04-15 R/si.R
fixes two notes

@@ -72,23 +72,26 @@ si_to_od = function(origins, destinations, max_euclidean_distance = 10000) {
 #'   such that the total for each zone of origin cannot go above this value.
 #' @param ... Arguments passed to `fun`
 #' @export
+#' @importFrom rlang .data
 #' @examples
 #' od = si_to_od(zones, zones)
 #' fun_dd = function(od, d = "distance_euclidean", beta = 0.3) exp(-beta * od[[d]] / 1000)
 #' od_dd = si_model(od, fun = fun_dd)
 #' plot(od$distance_euclidean, od_dd$res)
 #' fun = function(od, O, n, d, beta) od[[O]] * od[[n]] * exp(-beta * od[[d]]/1000)
-#' od_output = si_model(od, fun = fun, beta = 0.3, O = "origin_all", n = "destination_all", d = "distance_euclidean")
+#' od_output = si_model(od, fun = fun, beta = 0.3, O = "origin_all", 
+#'   n = "destination_all", d = "distance_euclidean")
 #' head(od_output)
 #' plot(od$distance_euclidean, od_output$res)
-#' od_pconst = si_model(od, fun = fun, beta = 0.3, O = "origin_all", n = "destination_all", d = "distance_euclidean", var_p = origin_all)
+#' od_pconst = si_model(od, fun = fun, beta = 0.3, O = "origin_all", n = "destination_all",
+#'   d = "distance_euclidean", var_p = origin_all)
 #' plot(od_pconst$distance_euclidean, od_pconst$res_constrained)
 #' plot(od_pconst["res"], logz = TRUE)
 si_model = function(od, fun, var_p, ...) {
     od$res = fun(od, ...)
     if (!missing(var_p)) {
-        od_grouped = dplyr::group_by(od, O)
-        od_grouped = dplyr::mutate(od_grouped, res = res / sum(res) * mean( {{var_p}} ))
+        od_grouped = dplyr::group_by(od, .data$O)
+        od_grouped = dplyr::mutate(od_grouped, res = .data$res / sum(.data$res) * mean( {{var_p}} ))
         od = dplyr::ungroup(od_grouped)
     }
     od

### 2022-04-15 README.qmd

@@
   cache: true
+  fig.path: "man/figures/README-"
 ---

### 2022-04-15 README.qmd
Re-add figures
Former-commit-id: dfe28103db676f8a40a96cf9410ada47fb2e85a7

--- 2022-04-15 README.qmd
@@
 execute:
-  cache: true
+  cache: false
+  fig.path: "man/figures/README-"
 ---

### 2022-04-15 R/si.R
Fix minor issue in docs

@@ -85,7 +85,7 @@ si_to_od = function(origins, destinations, max_euclidean_distance = 10000) {
 #' plot(od$distance_euclidean, od_output$res)
 #' od_pconst = si_model(od, fun = fun, beta = 0.3, O = "origin_all", n = "destination_all",
 #'   d = "distance_euclidean", var_p = origin_all)
-#' plot(od_pconst$distance_euclidean, od_pconst$res_constrained)
+#' plot(od_pconst$distance_euclidean, od_pconst$res)
 #' plot(od_pconst["res"], logz = TRUE)
 si_model = function(od, fun, var_p, ...) {
     od$res = fun(od, ...)

### 2022-04-15 R/si.R
Comment out plots

@@ -85,8 +85,8 @@ si_to_od = function(origins, destinations, max_euclidean_distance = 10000) {
 #' plot(od$distance_euclidean, od_output$res)
 #' od_pconst = si_model(od, fun = fun, beta = 0.3, O = "origin_all", n = "destination_all",
 #'   d = "distance_euclidean", var_p = origin_all)
-#' plot(od_pconst$distance_euclidean, od_pconst$res)
-#' plot(od_pconst["res"], logz = TRUE)
+#' # plot(od_pconst$distance_euclidean, od_pconst$res)
+#' # plot(od_pconst["res"], logz = TRUE)
 si_model = function(od, fun, var_p, ...) {
     od$res = fun(od, ...)
     if (!missing(var_p)) {

### 2022-04-15 R/si.R
Document

@@ -85,8 +85,8 @@ si_to_od = function(origins, destinations, max_euclidean_distance = 10000) {
 #' plot(od$distance_euclidean, od_output$res)
 #' od_pconst = si_model(od, fun = fun, beta = 0.3, O = "origin_all", n = "destination_all",
 #'   d = "distance_euclidean", var_p = origin_all)
-#' # plot(od_pconst$distance_euclidean, od_pconst$res)
-#' # plot(od_pconst["res"], logz = TRUE)
+#' plot(od_pconst$distance_euclidean, od_pconst$res)
+#' plot(od_pconst["res"], logz = TRUE)
 si_model = function(od, fun, var_p, ...) {
     od$res = fun(od, ...)
     if (!missing(var_p)) {

### 2022-04-15 README.qmd
Actions badge
Former-commit-id: a6fe4071ab57225a8e27a5445dbb5d0f79cec1fe
Update README with more resources

@@
 <!-- badges: start -->
-
+[![R-CMD-check](https://github.com/robinlovelace/si/workflows/R-CMD-check/badge.svg)](https://github.com/robinlovelace/si/actions)
 <!-- badges: end -->
@@
 
-TBC.
-
 # Resources
 
--   Reproducible code implementing SIMs in R (based on the outdated `sp` class system): https://rpubs.com/adam_dennett/257231
+- Reproducible [code](https://rpubs.com/adam_dennett/257231) implementing SIMs in R by Adam Dennet, based on the outdated `sp` class system, and based on [`sf` classes](https://github.com/adamdennett/LondonSchoolsSIM/blob/main/Schools.Rmd)
 
--
+- The [`spflow` R package](https://github.com/LukeCe/spflow) on CRAN
+
+- The [`spint` Python package](https://spint.readthedocs.io/en/latest/)

### 2022-04-15 README.qmd
Add sims vignette, close #1
Former-commit-id: ee3925f59c11c8c63fe76fd3afb016bf477359fb

@@
 title: "si: code and prose exploring Spatial Interaction models"
-bibliography: data-raw/references.bib
+bibliography: vignettes/references.bib
 execute:
@@
 
-The goal of this 'si' repo is to demonstrate Spatial Interaction models from first principles.
-At present it just contains prose and reproducible code demonstrating how they work.
-In future it may be turned into software, e.g.
-an R package, for others wanting to develop spatial interaction models.
+The goal of {si} is to demonstrate Spatial Interaction models from first principles (see the `sims` vignette for a more detailed description of SIMs and implementation in R without using this package for reference).
+The package also contains reproducible code and documentation demonstrating how they work.
 
-# What are SIMs?
-
-Spatial Interaction Models (SIMs) are mathematical models for estimating movement between spatial entities developed by Alan Wilson in the late 1960s and early 1970, with considerable uptake and refinement for transport modelling since then @boyce_forecasting_2015.
-There are four main types of traditional SIMs [@wilson_family_1971]:
+Spatial Interaction Models (SIMs) estimate movement between spatial entities.
+They have been central to transport modelling since the 1960s, with many refinements over the [decades](https://www.researchgate.net/publication/284345182_Forecasting_urban_travel_Past_present_and_future).
+There are four main [types](https://doi.org/10.1016/j.jtrangeo.2015.12.008) of traditional SIM:
 
@@
 
-Traditional 'gravity model' based SIMs can be defined as follows, based on the original definition [@wilson_family_1971]:
+The focus of this package is currently on unconstrained and production-constrained SIMs.
 
-$$
-T_{i j}=K \frac{W_{i}^{(1)} W_{j}^{(2)}}{c_{i j}^{n}}
-$$
+# Implementations in other languages
 
-"where $T_{i j}$ is a measure of the interaction between zones $i$ and $W_{i}^{(1)}$ is a measure of the 'mass term' associated with zone $z_i$, $W_{j}^{(2)}$ is a measure of the 'mass term' associated with zone $z_j$, and $c_{ij}$ is a measure of the distance, or generalised cost of travel, between zone $i$ and zone $j$".
-$K$ is a 'constant of proportionality' and $n$ is a parameter to be estimated.
-
-Redefining the $W$ terms as $m$ and $n$ for origins and destinations respectively [@simini_universal_2012], this classic definition of the 'gravity model' can be written as follows:
-
-$$
-T_{i j}=K \frac{m_{i} n_{j}}{c_{i j}^{n}}
-$$
-
-For the purposes of this project, we will focus on production-constrained SIMs.
-These can be defined as follows [@wilson_family_1971]:
-
-$$
-T_{ij} = A_iO_in_jf(c_{ij})
-$$
-
-where $A$ is a balancing factor defined as:
-
-$$
-A_{i}=\frac{1}{\sum_{j} m_{j} \mathrm{f}\left(c_{i j}\right)}
-$$
-
-$O_i$ is analogous to the travel demand in zone $i$, which can be roughly approximated by its population.
-
-<!-- $$ -->
-<!-- T_{ij} = \frac{m_{i}^{\alpha} n_{j} ^{\beta}} {f(r_{ij})} -->
-<!-- $$ -->
-
-<!-- ## Radiation models -->
-
-<!-- A more recent type of SIM is the radiation model @simini_universal_2012: -->
-
-<!-- $$ -->
-<!-- \langle T_{ij} \rangle = T_i \frac{m_{i} n_{j}} {(m_i + s_{ij})(m_i + n_j + s_{ij}) } -->
-<!-- $$ -->
-<!-- where $s_{ij}$ is defined as the total population living within a circle, the centre of which lies in the centroid of zone $i$ and the radius of which is the distance between zones $i$ and $j$.
-Thus, the greater the population living within the commute distance, the lower the estimated flow rate.
--->
-
-```{r}
-
-```
-
-# Implementation in R
-
-```{r, message=FALSE}
-library(tmap)
-library(pct)
-library(tidyverse)
-```
-
-```{r}
-#| label: inputs
-zones = get_pct_zones(region = "west-yorkshire", geography = "msoa")
-centroids = get_pct_centroids(region = "west-yorkshire", geography = "msoa")
-zones = zones %>% 
-  filter(lad_name == "Leeds")
-centroids = centroids %>% 
-  filter(lad_name == "Leeds")
-tm_shape(zones) + tm_polygons("all", palette = "viridis")
-```
-
-```{r}
-od_df = od::points_to_od(centroids)
-od_sfc = od::odc_to_sfc(od_df[3:6])
-sf::st_crs(od_sfc) = 4326
-od_df$length = sf::st_length(od_sfc)
-od_df = od_df %>% transmute(
-  O, D, length = as.numeric(length) / 1000,
-  flow = NA, fc = NA
-  )
-od_df = sf::st_sf(od_df, geometry = od_sfc, crs = 4326)
-```
-
-An unconstrained spatial interaction model can be written as follows, with a more-or-less arbitrary value for `beta` which can be optimised later:
-
-```{r}
-#| label: unconstrained
-beta = 0.3
-i = 1
-j = 2
-for(i in seq(nrow(zones))) {
-  for(j in seq(nrow(zones))) {
-    O = zones$all[i]
-    n = zones$all[j]
-    ij = which(od_df$O == zones$geo_code[i] & od_df$D == zones$geo_code[j])
-    od_df$fc[ij] = exp(-beta * od_df$length[ij])
-    od_df$flow[ij] = O * n * od_df$fc[ij]
-  }
-}
-od_top = od_df %>% 
-  filter(O != D) %>% 
-  top_n(n = 2000, wt = flow)
-
-tm_shape(zones) +
-  tm_borders() +
-  tm_shape(od_top) +
-  tm_lines("flow")
-```
-
-We can plot the 'distance decay' curve associated with this SIM is as follows:
-
-```{r}
-#| label: distance_decay
-summary(od_df$fc)
-od_df %>% 
-  ggplot() +
-  geom_point(aes(length, fc))
-```
-
-We can make this production constrained as follows:
-
-```{r}
-#| label: constrained
-od_dfj = left_join(
-  od_df,
-  zones %>% select(O = geo_code, all) %>% sf::st_drop_geometry()
-)
-od_dfj = od_dfj %>% 
-  group_by(O) %>% 
-  mutate(flow_constrained = flow / sum(flow) * first(all)) %>%
-  ungroup()
-sum(od_dfj$flow_constrained) == sum(zones$all)
-od_top = od_dfj %>% 
-  filter(O != D) %>% 
-  top_n(n = 2000, wt = flow_constrained)
-
-tm_shape(zones) +
-  tm_borders() +
-  tm_shape(od_top) +
-  tm_lines("flow_constrained")
-```
-
-# Validation
-
-```{r validation}
-od_census = get_od()
-od_census_min = od_census %>% 
-  rename(O = geo_code1, D = geo_code2) %>% 
-  filter(O %in% zones$geo_code & D %in% zones$geo_code)
-od_dfjc = inner_join(od_dfj %>% select(-all), od_census_min)
-od_dfjc %>% 
-  ggplot() +
-  geom_point(aes(all, flow_constrained))
-cor(od_dfjc$all, od_dfjc$flow_constrained)^2
-```
-
-# Implementation in Python
+We would be interested to hear how the approach presented in this package compared with other implementations such as those presented in the links below.
+If anyone would like to try the approach or implement it in another language feel free to get in touch via the issue tracker.
 

### 2022-04-15 README.qmd
Add readme with minimal example and description for #1
Former-commit-id: f2a35a8d2dac6764957ca2bd1b7c32a44e3ac588

@@
 format: gfm
-number-sections: true
 title: "si: code and prose exploring Spatial Interaction models"
@@
 
-The goal of {si} is to demonstrate Spatial Interaction models from first principles (see the `sims` vignette for a more detailed description of SIMs and implementation in R without using this package for reference).
+The goal of {si} is to demonstrate Spatial Interaction models from first principles (see the [`sims` vignette](https://robinlovelace.github.io/si/articles/sims.html) for a more detailed description of SIMs and implementation in R without using this package for reference).
 The package also contains reproducible code and documentation demonstrating how they work.
@@
 -   Unconstrained
-
 -   Production-constrained
-
 -   Attraction-constrained
-
 -   Doubly-constrained
@@
 
-# Implementations in other languages
+<!-- # Implementations in other languages -->
+## si basics
+
+Install the package as follows:
+
+```r
+install.packages("remotes") # if not already installed
+```
+
+```{r installation}
+remotes::install_github("robinlovelace/si")
+```
+
+Run a basic SIM as follows:
+
+```{r distance plot}
+library(si)
+od = si_to_od(origins = zones, destinations = zones, max_euclidean_distance = 5000)
+gravity_model = function(od, beta) {
+  od[["origin_all"]] * od[["destination_all"]] *
+    exp(-beta * od[["distance_euclidean"]] / 1000)
+} 
+od_res = si_model(od, fun = gravity_model, beta = 0.3, var_p = origin_all)
+plot(od_res$distance_euclidean, od_res$res)
+```
+
+What just happened?
+As the example above shows, the package allows/encourages you to define and use your own functions to estimate the amount of interaction/movement between places.
+The resulting estimates of interaction, returned in the column `res` and plotted with distance in the graphic above, resulted from our choice of spatial interaction model inputs, allowing a wide range of alternative approaches to be implemented.
+This flexibility is a key aspect of the package, enabling small and easily modified functions to be implemented and tested.
+
+The output `si_model()` is a geographic object which can be plotted as a map:
+
+```{r map}
+plot(od_res["res"], logz = TRUE)
+```
+
+Perhaps the most important function in the package is therefore `si_to_od()`, which transforms geographic entities (typically zones but points and other geographic features could be used) into a data frame representing the full combination of origin-destination pairs that are less than `max_euclidean_distance` meters apart.
+A common saying in data science is that 80% of the effort goes into the pre-processing stage.
+This is equally true for spatial interaction modelling as it is for other types of data intensive analysis/modelling work.
+So what does this function return?
+
+```{r}
+names(od)
+```
+
+As shown in the output above, the function allows you to use any variable in the origin or destination data in the function, with column names appended with `origin` and `destination`.
+
+Note: support for OD datasets in which the origins and destinations are different objects is work in progress (see [#3](https://github.com/Robinlovelace/si/issues/3)).
+
+## Feedback
 
@@
 
-# Resources
+## Alternative approaches
+
+The {si} package is nascent. Anyone looking for more mature approaches could check the following:
 

### 2022-04-15 R/si.R
Add missing %
Former-commit-id: b95d68286bf5015b2dc3a5ef37297a3c3f3d33e0

@@ -34,7 +34,7 @@ si_to_od = function(origins, destinations, max_euclidean_distance = 10000) {
           nrow_after,
           " OD pairs remaining after removing those with a distance greater than ", # nolint
           max_euclidean_distance, " meters", ":\n",
-          pct_kept, " of all possible OD pairs"
+          pct_kept, "% of all possible OD pairs"
           )
   }
   # join origin attributes

### 2022-04-15 README.qmd
Portable file names

@@
 
-```{r distance plot}
+```{r distance}
 library(si)

### 2022-04-16 R/si.R
Tidy up get started vignette
Former-commit-id: eee851b1e4b47d5ed69ce5cacb7fe6c985d0d6f6

@@ -15,7 +15,7 @@
 #' odsf = si_to_od(origins, destinations)
 #' plot(odsf)
 #' 
-si_to_od = function(origins, destinations, max_euclidean_distance = 10000) {
+si_to_od = function(origins, destinations, max_euclidean_distance = Inf) {
   if(!identical(origins, destinations)) {
         stop("Origins and destinations are different, not implemented yet")
     }

### 2022-04-16 README.qmd
Rename param, shorten code

@@
 library(si)
-od = si_to_od(origins = zones, destinations = zones, max_euclidean_distance = 5000)
+od = si_to_od(origins = zones, destinations = zones, max_dist = 5000)
 gravity_model = function(od, beta) {
@@
 
-Perhaps the most important function in the package is therefore `si_to_od()`, which transforms geographic entities (typically zones but points and other geographic features could be used) into a data frame representing the full combination of origin-destination pairs that are less than `max_euclidean_distance` meters apart.
+Perhaps the most important function in the package is therefore `si_to_od()`, which transforms geographic entities (typically zones but points and other geographic features could be used) into a data frame representing the full combination of origin-destination pairs that are less than `max_dist` meters apart.
 A common saying in data science is that 80% of the effort goes into the pre-processing stage.

### 2022-04-16 R/si.R

@@ -6,8 +6,9 @@
 #'
 #' @param origins `sf` object representing origin locations/zones
 #' @param destinations `sf` object representing destination locations/zones
-#' @param max_euclidean_distance Maximum distance in meters between OD pairs.
-#'   Only OD pairs that are this distance apart or less will be returned.
+#' @param max_dist Euclidean distance in meters (numeric).
+#'   Only OD pairs that are this distance apart or less will be returned
+#'   and therefore included in the SIM.
 #' @export
 #' @examples
 #' origins = centroids
@@ -15,7 +16,7 @@
 #' odsf = si_to_od(origins, destinations)
 #' plot(odsf)
 #' 
-si_to_od = function(origins, destinations, max_euclidean_distance = Inf) {
+si_to_od = function(origins, destinations, max_dist = Inf) {
   if(!identical(origins, destinations)) {
         stop("Origins and destinations are different, not implemented yet")
     }
@@ -25,15 +26,15 @@ si_to_od = function(origins, destinations, max_euclidean_distance = Inf) {
       y = od_df[c("dx", "dy")],
       paired = TRUE
       )
-  if(max(od_df$distance_euclidean) > max_euclidean_distance) {
+  if(max(od_df$distance_euclidean) > max_dist) {
       nrow_before = nrow(od_df)
-      od_df = od_df[od_df$distance_euclidean <= max_euclidean_distance, ]
+      od_df = od_df[od_df$distance_euclidean <= max_dist, ]
       nrow_after = nrow(od_df)
       pct_kept = round(nrow_after / nrow_before * 100)
       message(
           nrow_after,
           " OD pairs remaining after removing those with a distance greater than ", # nolint
-          max_euclidean_distance, " meters", ":\n",
+          max_dist, " meters", ":\n",
           pct_kept, "% of all possible OD pairs"
           )
   }

### 2022-04-16 R/si.R
More tidying, remove coords from od results
Former-commit-id: 7d997bf1b2e1dd3285e25101667a497f9b5c5811

@@ -42,6 +42,9 @@ si_to_od = function(origins, destinations, max_dist = Inf) {
   origins_to_join = sf::st_drop_geometry(origins)
   names(origins_to_join) = paste0("origin_", names(origins_to_join))
   names(origins_to_join)[1] = names(od_df)[1]
+  od_sfc = od::odc_to_sfc(od_df[3:6])
+  sf::st_crs(od_sfc) = 4326 # todo: add CRS argument?
+  od_df = od_df[-c(3:6)]
   od_dfj = dplyr::inner_join(od_df, origins_to_join, by = "O")
   # join destination attributes
   destinations_to_join = sf::st_drop_geometry(destinations)
@@ -49,11 +52,6 @@ si_to_od = function(origins, destinations, max_dist = Inf) {
   names(destinations_to_join)[1] = names(od_df)[2]
   od_dfj = dplyr::inner_join(od_dfj, destinations_to_join, by = "D")
   names(od_dfj)
-
-  # convert to sfc:
-
-  od_sfc = od::odc_to_sfc(od_df[3:6])
-  sf::st_crs(od_sfc) = 4326 # todo: add CRS argument?
   # create and return sf object
   sf::st_sf(od_dfj, geometry = od_sfc)
 }

### 2022-04-16 R/data.R
New line

@@ -1,12 +1,16 @@
 #' Example zones and centroids
 #'
-#' `zones` and `centroids` represent administrative zones between which
+#' `si_zones` and `si_centroids` represent administrative zones between which
 #' flows are to be estimated.
 #' @note The schema data can be (re-)generated using code in `data-raw`
 #' @docType data
 #' @keywords datasets
-#' @name zones
-#' @aliases centroids
+#' @name si_zones
+#' @aliases si_centroids
+#' @examples
+#' si_zones
+#' plot(si_zones$geometry)
+#' plot(si_centroids$geometry, add = TRUE)
 NULL
 
 #' Example OD dataset
@@ -16,5 +20,7 @@ NULL
 #'
 #' @docType data
 #' @keywords datasets
-#' @name od_census_min
+#' @name si_od_census
+#' @examples
+#' head(si_od_census)
 NULL

### 2022-04-16 data-raw/zones.R
New line
Former-commit-id: 43509f18d40cb2d73af972066351c2eef8b86c83

@@ -3,17 +3,17 @@ library(pct)
 library(tidyverse)
 zones = get_pct_zones(region = "west-yorkshire", geography = "msoa")
 centroids = get_pct_centroids(region = "west-yorkshire", geography = "msoa")
-zones = zones %>% 
+si_zones = zones %>% 
   filter(lad_name == "Leeds") %>%
   select(1:13)
-centroids = centroids %>% 
+si_centroids = centroids %>% 
   filter(lad_name == "Leeds") %>%
   select(1:13)
-usethis::use_data(zones, overwrite = TRUE)
-usethis::use_data(centroids, overwrite = TRUE)
+usethis::use_data(si_zones, overwrite = TRUE)
+usethis::use_data(si_centroids, overwrite = TRUE)
 
 od_census = get_od()
-od_census_min = od_census %>% 
+si_od_census = od_census %>% 
   rename(O = geo_code1, D = geo_code2) %>% 
   filter(O %in% zones$geo_code & D %in% zones$geo_code)
-usethis::use_data(od_census_min)
+usethis::use_data(si_od_census)

### 2022-04-16 R/si.R
Change name of example datasets, close #2
Former-commit-id: e914f3bb9645e305a3f816282e7dfdd160dd46d7

@@ -11,11 +11,10 @@
 #'   and therefore included in the SIM.
 #' @export
 #' @examples
-#' origins = centroids
-#' destinations = centroids
-#' odsf = si_to_od(origins, destinations)
+#' origins = si_centroids
+#' destinations = si_centroids
+#' odsf = si_to_od(origins, destinations, max_dist = 2000)
 #' plot(odsf)
-#' 
 si_to_od = function(origins, destinations, max_dist = Inf) {
   if(!identical(origins, destinations)) {
         stop("Origins and destinations are different, not implemented yet")
@@ -73,7 +72,7 @@ si_to_od = function(origins, destinations, max_dist = Inf) {
 #' @export
 #' @importFrom rlang .data
 #' @examples
-#' od = si_to_od(zones, zones)
+#' od = si_to_od(si_zones, si_zones, max_dist = 4000)
 #' fun_dd = function(od, d = "distance_euclidean", beta = 0.3) exp(-beta * od[[d]] / 1000)
 #' od_dd = si_model(od, fun = fun_dd)
 #' plot(od$distance_euclidean, od_dd$res)

### 2022-04-16 README.qmd
Change name of example datasets, close #2

@@
 
-```{r installation}
+```r
 remotes::install_github("robinlovelace/si")
@@
 library(si)
-od = si_to_od(origins = zones, destinations = zones, max_dist = 5000)
+od = si_to_od(
+  origins = si_zones,
+  destinations = si_zones,
+  max_dist = 5000
+  )
 gravity_model = function(od, beta) {

### 2022-04-16 README.qmd
Update readme, move vignette

@@
 
-The goal of {si} is to demonstrate Spatial Interaction models from first principles (see the [`sims` vignette](https://robinlovelace.github.io/si/articles/sims.html) for a more detailed description of SIMs and implementation in R without using this package for reference).
-The package also contains reproducible code and documentation demonstrating how they work.
-
-Spatial Interaction Models (SIMs) estimate movement between spatial entities.
-They have been central to transport modelling since the 1960s, with many refinements over the [decades](https://www.researchgate.net/publication/284345182_Forecasting_urban_travel_Past_present_and_future).
-There are four main [types](https://doi.org/10.1016/j.jtrangeo.2015.12.008) of traditional SIM:
-
--   Unconstrained
--   Production-constrained
--   Attraction-constrained
--   Doubly-constrained
-
-The focus of this package is currently on unconstrained and production-constrained SIMs.
+The goal of {si} is to provide a simple and flexible framework for developing spatial interaction models (SIMs), which estimate the amount of movement between spatial entities.
 
@@
 
-Perhaps the most important function in the package is therefore `si_to_od()`, which transforms geographic entities (typically zones but points and other geographic features could be used) into a data frame representing the full combination of origin-destination pairs that are less than `max_dist` meters apart.
+The `si_to_od()` function transforms geographic entities (typically polygons and points) into a data frame representing the full combination of origin-destination pairs that are less than `max_dist` meters apart.
 A common saying in data science is that 80% of the effort goes into the pre-processing stage.
@@
 
-## Alternative approaches
+## Further reading
 
-The {si} package is nascent. Anyone looking for more mature approaches could check the following:
+For details on what SIMs are and how they have been defined mathematically and in code from first principles, see the [`sims` vignette](https://robinlovelace.github.io/si/articles/sims-first-principles.html).
 
-- Reproducible [code](https://rpubs.com/adam_dennett/257231) implementing SIMs in R by Adam Dennet, based on the outdated `sp` class system, and based on [`sf` classes](https://github.com/adamdennett/LondonSchoolsSIM/blob/main/Schools.Rmd)
+To dive straight into using {si} to develop SIMs, see the [`si` Get started vignette](https://robinlovelace.github.io/si/articles/si.html).
+
+For a detailed introduction to SIMs, support by reproducible R code, see Adam Dennett's [2018 paper](https://doi.org/10.37970/aps.v2i2.38).
+
+## Other SIM packages
 
 - The [`spflow` R package](https://github.com/LukeCe/spflow) on CRAN
-
 - The [`spint` Python package](https://spint.readthedocs.io/en/latest/)

### 2022-04-17 R/si.R
Destination examples, progress on #3

@@ -11,15 +11,28 @@
 #'   and therefore included in the SIM.
 #' @export
 #' @examples
-#' origins = si_centroids
-#' destinations = si_centroids
-#' odsf = si_to_od(origins, destinations, max_dist = 2000)
+#' library(sf)
+#' origins = si_centroids[c(1, 2, 99), ]
+#' destinations = origins
+#' plot(origins$geometry)
+#' odsf = si_to_od(origins, destinations, max_dist = 1200)
+#' plot(odsf)
+#' # note: result contains intrazonal flows represented by linestrings
+#' # with a length of 0, e.g.:
+#' sf::st_coordinates(odsf$geometry[1])
+#' # With different destinations compared with origins
+#' library(sf)
+#' origins = si_centroids[c(2, 99), c(1, 6, 7)]
+#' destinations = si_centroids[1, c(1, 6, 8)]
+#' odsf = si_to_od(origins, destinations)
+#' nrow(odsf) # no intrazonal flows
 #' plot(odsf)
 si_to_od = function(origins, destinations, max_dist = Inf) {
-  if(!identical(origins, destinations)) {
-        stop("Origins and destinations are different, not implemented yet")
-    }
-  od_df = od::points_to_od(origins)
+  if(identical(origins, destinations)) {
+    od_df = od::points_to_od(origins)
+  } else {
+    od_df = od::points_to_od(origins, destinations)
+  }
   od_df$distance_euclidean = geodist::geodist(
       x = od_df[c("ox", "oy")],
       y = od_df[c("dx", "dy")],
@@ -37,20 +50,21 @@ si_to_od = function(origins, destinations, max_dist = Inf) {
           pct_kept, "% of all possible OD pairs"
           )
   }
+  od_sfc = od::odc_to_sfc(od_df[3:6])
+  sf::st_crs(od_sfc) = 4326 # todo: add CRS argument?
+  od_df = od_df[-c(3:6)]
+  
   # join origin attributes
   origins_to_join = sf::st_drop_geometry(origins)
   names(origins_to_join) = paste0("origin_", names(origins_to_join))
   names(origins_to_join)[1] = names(od_df)[1]
-  od_sfc = od::odc_to_sfc(od_df[3:6])
-  sf::st_crs(od_sfc) = 4326 # todo: add CRS argument?
-  od_df = od_df[-c(3:6)]
   od_dfj = dplyr::inner_join(od_df, origins_to_join, by = "O")
   # join destination attributes
[…]

### 2022-04-18 README.qmd
improves the readme

@@
 
-<!-- # Implementations in other languages -->
-## si basics
+## Installation
 
@@
 
+<!-- # Implementations in other languages -->
+## si basics
+
 Run a basic SIM as follows:
@@
 library(si)
+# prepare OD data
 od = si_to_od(
-  origins = si_zones,
-  destinations = si_zones,
-  max_dist = 5000
+  origins = si_zones,        # origin locations
+  destinations = si_zones,   # destination locations
+  max_dist = 5000            # maximum distance between OD pairs
   )
+# specify a function
 gravity_model = function(od, beta) {
@@
 } 
+# perform SIM
 od_res = si_model(od, fun = gravity_model, beta = 0.3, var_p = origin_all)
+# visualize the results
 plot(od_res$distance_euclidean, od_res$res)

### 2022-04-18 R/si_model.R
Create si_predict/si_calculate for #5
Former-commit-id: 01a92c2c5e8b24f8656ecdbe82de24278b61bc51

@@ -1,4 +1,4 @@
-#' Run a spatial interaction model
+#' Calculate flow using a pre-existing function
 #'
 #' Executes a spatial interaction model based on an OD data frame 
 #' and user-specified function
@@ -7,7 +7,7 @@
 #'   [si_to_od()]
 #' @param fun A function that calculates the interaction (e.g. the number of trips)
 #'   between each OD pair
-#' @param var_p Use this argument to run a 'production constrained' SIM.
+#' @param constraint_p Use this argument to run a 'production constrained' SIM.
 #'   Character string corresponding to column in the `od`
 #'   dataset that constrains the total 'interaction' (e.g. n. trips) for all OD pairs
 #'   such that the total for each zone of origin cannot go above this value.
@@ -17,22 +17,43 @@
 #' @examples
 #' od = si_to_od(si_zones, si_zones, max_dist = 4000)
 #' fun_dd = function(od, d = "distance_euclidean", beta = 0.3) exp(-beta * od[[d]] / 1000)
-#' od_dd = si_model(od, fun = fun_dd)
+#' od_dd = si_calculate(od, fun = fun_dd)
 #' plot(od$distance_euclidean, od_dd$res)
 #' fun = function(od, O, n, d, beta) od[[O]] * od[[n]] * exp(-beta * od[[d]]/1000)
-#' od_output = si_model(od, fun = fun, beta = 0.3, O = "origin_all", 
+#' od_output = si_calculate(od, fun = fun, beta = 0.3, O = "origin_all", 
 #'   n = "destination_all", d = "distance_euclidean")
 #' head(od_output)
 #' plot(od$distance_euclidean, od_output$res)
-#' od_pconst = si_model(od, fun = fun, beta = 0.3, O = "origin_all", n = "destination_all",
-#'   d = "distance_euclidean", var_p = origin_all)
+#' od_pconst = si_calculate(od, fun = fun, beta = 0.3, O = "origin_all", n = "destination_all",
+#'   d = "distance_euclidean", constraint_p = origin_all)
 #' plot(od_pconst$distance_euclidean, od_pconst$res)
 #' plot(od_pconst["res"], logz = TRUE)
-si_model = function(od, fun, var_p, ...) {
+si_calculate = function(od, fun, constraint_p, ...) {
   od$res = fun(od, ...)
-  if (!missing(var_p)) {
+  if (!missing(constraint_p)) {
[…]

### 2022-04-18 README.qmd
Update README
Former-commit-id: 293606045649ce3ad94064f4ff308656a3761679

--- 2022-04-16 README.qmd
@@
 } 
-od_res = si_model(od, fun = gravity_model, beta = 0.3, var_p = origin_all)
+# perform SIM
+od_res = si_calculate(od, fun = gravity_model, beta = 0.3, var_p = origin_all)
+# visualize the results
 plot(od_res$distance_euclidean, od_res$res)
@@
 
-The output `si_model()` is a geographic object which can be plotted as a map:
+The output `si_calculate)` is a geographic object which can be plotted as a map:
 

### 2022-04-21 README.qmd
Change output variable name for #9
Former-commit-id: 68bf56d2039a28ccdd978670f21764ca145dd7a4

@@
 # visualize the results
-plot(od_res$distance_euclidean, od_res$res)
+plot(od_res$distance_euclidean, od_res$interaction)
 ```

### 2022-04-21 R/si_model.R
Change output variable name for #9

@@ -18,21 +18,25 @@
 #' od = si_to_od(si_zones, si_zones, max_dist = 4000)
 #' fun_dd = function(od, d = "distance_euclidean", beta = 0.3) exp(-beta * od[[d]] / 1000)
 #' od_dd = si_calculate(od, fun = fun_dd)
-#' plot(od$distance_euclidean, od_dd$res)
+#' plot(od$distance_euclidean, od_dd$interaction)
 #' fun = function(od, O, n, d, beta) od[[O]] * od[[n]] * exp(-beta * od[[d]]/1000)
 #' od_output = si_calculate(od, fun = fun, beta = 0.3, O = "origin_all", 
 #'   n = "destination_all", d = "distance_euclidean")
 #' head(od_output)
-#' plot(od$distance_euclidean, od_output$res)
+#' plot(od$distance_euclidean, od_output$interaction)
 #' od_pconst = si_calculate(od, fun = fun, beta = 0.3, O = "origin_all", n = "destination_all",
 #'   d = "distance_euclidean", constraint_p = origin_all)
-#' plot(od_pconst$distance_euclidean, od_pconst$res)
-#' plot(od_pconst["res"], logz = TRUE)
+#' plot(od_pconst$distance_euclidean, od_pconst$interaction)
+#' plot(od_pconst["interaction"], logz = TRUE)
 si_calculate = function(od, fun, constraint_p, ...) {
-  od$res = fun(od, ...)
+  od$interaction = fun(od, ...)
   if (!missing(constraint_p)) {
     od_grouped = dplyr::group_by(od, .data$O)
-    od_grouped = dplyr::mutate(od_grouped, res = .data$res / sum(.data$res) * mean( {{constraint_p}} ))
+    od_grouped = dplyr::mutate(
+      od_grouped,
+      interaction = .data$interaction /
+        sum(.data$interaction) * mean( {{constraint_p}} )
+      )
     od = dplyr::ungroup(od_grouped)
   }
   od
@@ -48,11 +52,13 @@ si_calculate = function(od, fun, constraint_p, ...) {
 #' m = lm(od$origin_all ~ od$origin_bicycle)
 #' od_updated = si_predict(od, m)
 si_predict = function(od, model, constraint_p) {
-  od$res = stats::predict(model, od)
+  od$interaction = stats::predict(model, od)
   if (!missing(constraint_p)) {
     od_grouped = dplyr::group_by(od, .data$O)
     od_grouped = dplyr::mutate(
[…]

### 2022-04-21 README.qmd
Update readme

--- 2022-04-18 README.qmd
@@
 # perform SIM
-od_res = si_calculate(od, fun = gravity_model, beta = 0.3, var_p = origin_all)
+od_res = si_calculate(od, fun = gravity_model, beta = 0.3, constraint_p = origin_all)
 # visualize the results
-plot(od_res$distance_euclidean, od_res$res)
+plot(od_res$distance_euclidean, od_res$interaction)
 ```
@@
 ```{r map}
-plot(od_res["res"], logz = TRUE)
+plot(od_res["interaction"], logz = TRUE)
 ```
@@
 - The [`spint` Python package](https://spint.readthedocs.io/en/latest/)
+- The [`gravity`](https://cran.r-project.org/package=gravity) R package

### 2022-04-21 R/si_model.R
Switch to tidyeval for #5

@@ -16,20 +16,22 @@
 #' @importFrom rlang .data
 #' @examples
 #' od = si_to_od(si_zones, si_zones, max_dist = 4000)
-#' fun_dd = function(od, d = "distance_euclidean", beta = 0.3) exp(-beta * od[[d]] / 1000)
-#' od_dd = si_calculate(od, fun = fun_dd)
+#' fun_dd = function(d = "distance_euclidean", beta = 0.3) exp(-beta * d / 1000)
+#' fun_dd(d = (1:5)*1000)
+#' od_dd = si_calculate(od, fun = fun_dd, d = distance_euclidean)
 #' plot(od$distance_euclidean, od_dd$interaction)
-#' fun = function(od, O, n, d, beta) od[[O]] * od[[n]] * exp(-beta * od[[d]]/1000)
-#' od_output = si_calculate(od, fun = fun, beta = 0.3, O = "origin_all", 
-#'   n = "destination_all", d = "distance_euclidean")
+#' fun = function(O, n, d, beta) O * n * exp(-beta * d / 1000)
+#' od_output = si_calculate(od, fun = fun, beta = 0.3, O = origin_all, 
+#'   n = destination_all, d = distance_euclidean)
 #' head(od_output)
 #' plot(od$distance_euclidean, od_output$interaction)
-#' od_pconst = si_calculate(od, fun = fun, beta = 0.3, O = "origin_all", n = "destination_all",
-#'   d = "distance_euclidean", constraint_p = origin_all)
+#' od_pconst = si_calculate(od, fun = fun, beta = 0.3, O = origin_all, n = destination_all,
+#'   d = distance_euclidean, constraint_p = origin_all)
 #' plot(od_pconst$distance_euclidean, od_pconst$interaction)
 #' plot(od_pconst["interaction"], logz = TRUE)
 si_calculate = function(od, fun, constraint_p, ...) {
-  od$interaction = fun(od, ...)
+  dots = rlang::enquos(...)
+  od = mutate(od, interaction = fun(!!!dots))
   if (!missing(constraint_p)) {
     od_grouped = dplyr::group_by(od, .data$O)
     od_grouped = dplyr::mutate(

### 2022-04-21 R/si_model.R
Complete switch to tidyeval, progress on #5

@@ -25,13 +25,13 @@
 #'   n = destination_all, d = distance_euclidean)
 #' head(od_output)
 #' plot(od$distance_euclidean, od_output$interaction)
-#' od_pconst = si_calculate(od, fun = fun, beta = 0.3, O = origin_all, n = destination_all,
-#'   d = distance_euclidean, constraint_p = origin_all)
+#' od_pconst = si_calculate(od, fun = fun, beta = 0.3, O = origin_all,
+#'   n = destination_all, d = distance_euclidean, constraint_p = origin_all)
 #' plot(od_pconst$distance_euclidean, od_pconst$interaction)
 #' plot(od_pconst["interaction"], logz = TRUE)
 si_calculate = function(od, fun, constraint_p, ...) {
   dots = rlang::enquos(...)
-  od = mutate(od, interaction = fun(!!!dots))
+  od = dplyr::mutate(od, interaction = fun(!!!dots))
   if (!missing(constraint_p)) {
     od_grouped = dplyr::group_by(od, .data$O)
     od_grouped = dplyr::mutate(

### 2022-04-22 README.qmd
Update with tidyeval example
Former-commit-id: 854122a5dca7e0601224aeb74d6d244f34de0660

@@
 library(si)
+library(dplyr)
 # prepare OD data
@@
 # specify a function
-gravity_model = function(od, beta) {
-  od[["origin_all"]] * od[["destination_all"]] *
-    exp(-beta * od[["distance_euclidean"]] / 1000)
+gravity_model = function(beta, d, m, n) {
+  m * n * exp(-beta * d / 1000)
 } 
 # perform SIM
-od_res = si_calculate(od, fun = gravity_model, beta = 0.3, constraint_p = origin_all)
+od_res = si_calculate(
+  od,
+  fun = gravity_model,
+  d = distance_euclidean,
+  m = origin_all,
+  n = destination_all,
+  constraint_p = origin_all,
+  beta = 0.3
+  )
 # visualize the results
 plot(od_res$distance_euclidean, od_res$interaction)
+```
+
+Note: this approach also allows you to use {si} functions in {dplyr} pipelines:
+
+```{r}
+od_res = od %>% 
+  si_calculate(fun = gravity_model, 
+               m = origin_all,
+               n = destination_all,
+               d = distance_euclidean,
+               beta = 0.3)
+od_res %>% 
+  select(interaction) %>% 
+  plot()
 ```

### 2022-04-22 README.qmd
Update with logz map
Former-commit-id: 6a9eac73e9a4846114354f8c2020aced72cea4ff

@@
 
+What just happened?
+We created an 'OD data frame' from geographic origins and destinations, and then estimated a simple gravity model based on the population in origin and destination zones and a custom distance decay function.
+As the example above shows, the package allows/encourages you to define and use your own functions to estimate the amount of interaction/movement between places.
+
 Note: this approach also allows you to use {si} functions in {dplyr} pipelines:
@@
 od_res %>% 
-  select(interaction) %>% 
-  plot()
+  select(interaction)
 ```
 
-What just happened?
-As the example above shows, the package allows/encourages you to define and use your own functions to estimate the amount of interaction/movement between places.
-The resulting estimates of interaction, returned in the column `res` and plotted with distance in the graphic above, resulted from our choice of spatial interaction model inputs, allowing a wide range of alternative approaches to be implemented.
+The resulting estimates of interaction, returned in the column `interaction` and plotted with distance in the graphic above, resulted from our choice of spatial interaction model inputs, allowing a wide range of alternative approaches to be implemented.
 This flexibility is a key aspect of the package, enabling small and easily modified functions to be implemented and tested.

### 2022-04-22 README.qmd
Update with nse the same as se

@@
 What just happened?
-We created an 'OD data frame' from geographic origins and destinations, and then estimated a simple gravity model based on the population in origin and destination zones and a custom distance decay function.
+We created an 'OD data frame' with the function `si_to_od()` from geographic origins and destinations, and then estimated a simple 'production constrained' (with the `constraint_p` argument) gravity model based on the population in origin and destination zones and a custom distance decay function with `si_calculate()`.
 As the example above shows, the package allows/encourages you to define and use your own functions to estimate the amount of interaction/movement between places.
 
-Note: this approach also allows you to use {si} functions in {dplyr} pipelines:
+The approach is also 'tidy', allowing use of {si} functions in {dplyr} pipelines:
 
@@
                d = distance_euclidean,
+               constraint_p = origin_all,
                beta = 0.3)

### 2022-04-22 README.qmd

@@
   fig.path: "man/figures/README-"
+  message: false
 ---
@@
 knitr::opts_chunk$set(cache = TRUE,
-  fig.path = "man/figures/README-")
+  fig.path = "man/figures/README-",
+  message = FALSE)
 ```

### 2022-04-22 README.qmd
Tweak readme
Former-commit-id: c6f5f113465bfe6929388bb3bbc8fedae03ec16f

@@
 format: gfm
-title: "si: code and prose exploring Spatial Interaction models"
 bibliography: vignettes/references.bib
@@
 
-The goal of {si} is to provide a simple and flexible framework for developing spatial interaction models (SIMs), which estimate the amount of movement between spatial entities.
+The goal of {si} is to provide a simple, '[tidy](https://dplyr.tidyverse.org/)', and flexible framework for developing spatial interaction models (SIMs), which estimate the amount of movement between spatial entities.
 

### 2022-04-22 data-raw/zones.R
Set-up for #11
Former-commit-id: 48c3670ff89fe067e84eee0fe4bc7009d2959894

@@ -16,4 +16,4 @@ od_census = get_od()
 si_od_census = od_census %>% 
   rename(O = geo_code1, D = geo_code2) %>% 
   filter(O %in% zones$geo_code & D %in% zones$geo_code)
-usethis::use_data(si_od_census)
+usethis::use_data(si_od_census, overwrite = TRUE)

### 2022-04-23 R/si_model.R
Add output_col argument, close #9

@@ -12,6 +12,8 @@
 #'   dataset that constrains the total 'interaction' (e.g. n. trips) for all OD pairs
 #'   such that the total for each zone of origin cannot go above this value.
 #' @param ... Arguments passed to `fun`
+#' @param output_col Character string containing the name of the new output
+#'   column. `"interaction"` by default.
 #' @export
 #' @importFrom rlang .data
 #' @examples
@@ -29,15 +31,17 @@
 #'   n = destination_all, d = distance_euclidean, constraint_p = origin_all)
 #' plot(od_pconst$distance_euclidean, od_pconst$interaction)
 #' plot(od_pconst["interaction"], logz = TRUE)
-si_calculate = function(od, fun, constraint_p, ...) {
+#' od_dd = si_calculate(od, fun = fun_dd, d = distance_euclidean, output_col = "res")
+#' head(od_dd$res)
+si_calculate = function(od, fun, constraint_p, ..., output_col = "interaction") {
   dots = rlang::enquos(...)
-  od = dplyr::mutate(od, interaction = fun(!!!dots))
+  od = dplyr::mutate(od, {{output_col}} := fun(!!!dots))
   if (!missing(constraint_p)) {
     od_grouped = dplyr::group_by(od, .data$O)
     od_grouped = dplyr::mutate(
       od_grouped,
-      interaction = .data$interaction /
-        sum(.data$interaction) * mean( {{constraint_p}} )
+      output_col = .data[[output_col]] /
+        sum(.data[[output_col]]) * mean( {{constraint_p}} )
       )
     od = dplyr::ungroup(od_grouped)
   }
@@ -53,14 +57,14 @@ si_calculate = function(od, fun, constraint_p, ...) {
 #' od = si_to_od(si_zones, si_zones, max_dist = 4000)
 #' m = lm(od$origin_all ~ od$origin_bicycle)
 #' od_updated = si_predict(od, m)
-si_predict = function(od, model, constraint_p) {
-  od$interaction = stats::predict(model, od)
+si_predict = function(od, model, constraint_p, output_col = "interaction") {
+  od[[output_col]] = stats::predict(model, od)
   if (!missing(constraint_p)) {
     od_grouped = dplyr::group_by(od, .data$O)
     od_grouped = dplyr::mutate(
       od_grouped,
-      interaction = .data$interaction /
-        sum(.data$interaction) * mean( {{constraint_p}} )
[…]

### 2022-04-23 R/globals.R
Add output_col argument, close #9
Former-commit-id: 549cb5e6ada9953f7a3169c8fa75044205695e04

@@ -1,0 +1,1 @@
+utils::globalVariables(c(":="))

### 2022-04-23 data-raw/zones.R
Update si_od_census object

@@ -15,5 +15,5 @@ usethis::use_data(si_centroids, overwrite = TRUE)
 od_census = get_od()
 si_od_census = od_census %>% 
   rename(O = geo_code1, D = geo_code2) %>% 
-  filter(O %in% zones$geo_code & D %in% zones$geo_code)
+  filter(O %in% si_zones$geo_code & D %in% si_zones$geo_code)
 usethis::use_data(si_od_census, overwrite = TRUE)

### 2022-04-23 README.qmd
Rename constraint_p to constraint_production #12
Former-commit-id: 4529e7d21166bac415501a34da0050828d3dfdb6

@@
   n = destination_all,
-  constraint_p = origin_all,
+  constraint_production = origin_all,
   beta = 0.3
@@
 What just happened?
-We created an 'OD data frame' with the function `si_to_od()` from geographic origins and destinations, and then estimated a simple 'production constrained' (with the `constraint_p` argument) gravity model based on the population in origin and destination zones and a custom distance decay function with `si_calculate()`.
+We created an 'OD data frame' with the function `si_to_od()` from geographic origins and destinations, and then estimated a simple 'production constrained' (with the `constraint_production` argument) gravity model based on the population in origin and destination zones and a custom distance decay function with `si_calculate()`.
 As the example above shows, the package allows/encourages you to define and use your own functions to estimate the amount of interaction/movement between places.
@@
                d = distance_euclidean,
-               constraint_p = origin_all,
+               constraint_production = origin_all,
                beta = 0.3)

### 2022-04-23 R/si_model.R

@@ -7,7 +7,7 @@
 #'   [si_to_od()]
 #' @param fun A function that calculates the interaction (e.g. the number of trips)
 #'   between each OD pair
-#' @param constraint_p Use this argument to run a 'production constrained' SIM.
+#' @param constraint_production Use this argument to run a 'production constrained' SIM.
 #'   Character string corresponding to column in the `od`
 #'   dataset that constrains the total 'interaction' (e.g. n. trips) for all OD pairs
 #'   such that the total for each zone of origin cannot go above this value.
@@ -28,20 +28,20 @@
 #' head(od_output)
 #' plot(od$distance_euclidean, od_output$interaction)
 #' od_pconst = si_calculate(od, fun = fun, beta = 0.3, O = origin_all,
-#'   n = destination_all, d = distance_euclidean, constraint_p = origin_all)
+#'   n = destination_all, d = distance_euclidean, constraint_production = origin_all)
 #' plot(od_pconst$distance_euclidean, od_pconst$interaction)
 #' plot(od_pconst["interaction"], logz = TRUE)
 #' od_dd = si_calculate(od, fun = fun_dd, d = distance_euclidean, output_col = "res")
 #' head(od_dd$res)
-si_calculate = function(od, fun, constraint_p, ..., output_col = "interaction") {
+si_calculate = function(od, fun, constraint_production, ..., output_col = "interaction") {
   dots = rlang::enquos(...)
   od = dplyr::mutate(od, {{output_col}} := fun(!!!dots))
-  if (!missing(constraint_p)) {
+  if (!missing(constraint_production)) {
     od_grouped = dplyr::group_by(od, .data$O)
     od_grouped = dplyr::mutate(
       od_grouped,
       output_col = .data[[output_col]] /
-        sum(.data[[output_col]]) * mean( {{constraint_p}} )
+        sum(.data[[output_col]]) * mean( {{constraint_production}} )
       )
     od = dplyr::ungroup(od_grouped)
   }
@@ -57,14 +57,14 @@ si_calculate = function(od, fun, constraint_p, ..., output_col = "interaction") 
 #' od = si_to_od(si_zones, si_zones, max_dist = 4000)
 #' m = lm(od$origin_all ~ od$origin_bicycle)
 #' od_updated = si_predict(od, m)
-si_predict = function(od, model, constraint_p, output_col = "interaction") {
+si_predict = function(od, model, constraint_production, output_col = "interaction") {
[…]

### 2022-04-23 R/si_model.R
Add constraint_ args, close #12
Former-commit-id: 5c15aea59322d149ef9aed0657e34d2ccd0beb5f

@@ -7,10 +7,17 @@
 #'   [si_to_od()]
 #' @param fun A function that calculates the interaction (e.g. the number of trips)
 #'   between each OD pair
-#' @param constraint_production Use this argument to run a 'production constrained' SIM.
-#'   Character string corresponding to column in the `od`
-#'   dataset that constrains the total 'interaction' (e.g. n. trips) for all OD pairs
-#'   such that the total for each zone of origin cannot go above this value.
+#' @param constraint_production Character representing column in `od`.
+#'   This argument, when set, ensures that the outputs are 'production constrained':
+#'   the total 'interaction' (e.g. n. trips) for all OD pairs is set such that
+#'   the total for each zone of origin cannot go above this value.
+#' @param constraint_attraction Character representing column in `od`.
+#'   This argument, when set, ensures that the outputs are 'attraction constrained':
+#'   the total 'interaction' (e.g. n. trips) for all OD pairs is set such that
+#'   the sum of trips to destination is equal to the mean value per destination.
+#' @param constraint_total Single number representing the total interaction.
+#'   This argument, when set, ensures that the sum of the interaction
+#'   calculated will equal the value given.
 #' @param ... Arguments passed to `fun`
 #' @param output_col Character string containing the name of the new output
 #'   column. `"interaction"` by default.
@@ -33,17 +40,27 @@
 #' plot(od_pconst["interaction"], logz = TRUE)
 #' od_dd = si_calculate(od, fun = fun_dd, d = distance_euclidean, output_col = "res")
 #' head(od_dd$res)
-si_calculate = function(od, fun, constraint_production, ..., output_col = "interaction") {
+#' od_dd = si_calculate(od, fun = fun_dd, d = distance_euclidean, constraint_total = 10)
+#' sum(od_dd$interaction)
+si_calculate = function(
+    od,
+    fun,
+    constraint_production,
+    constraint_attraction,
+    constraint_total,
+    output_col = "interaction",
+    ...
+    ) {
   dots = rlang::enquos(...)
   od = dplyr::mutate(od, {{output_col}} := fun(!!!dots))
   if (!missing(constraint_production)) {
-    od_grouped = dplyr::group_by(od, .data$O)
-    od_grouped = dplyr::mutate(
-      od_grouped,
-      output_col = .data[[output_col]] /
[…]

### 2022-04-24 R/si_model.R
Disallow negative values with constraint_total, update text
Former-commit-id: df1513f304e26f9c0e7af7acd3f7c719879567b1

@@ -109,6 +109,10 @@ constrain_production = function(od, output_col, constraint_production) {
 }
 
 constrain_total = function(od, output_col, constraint_total) {
+  if(min(od[[output_col]]) < 0) {
+    message("Negative values in output, setting them to zero")
+    od[[output_col]][od[[output_col]] < 0] = 0
+  }
   od[[output_col]] = od[[output_col]] / sum(od[[output_col]]) * constraint_total
   od
 }

### 2022-04-25 R/si_model.R
Implement and test production constrained SIMs
Former-commit-id: 0021274be9a743f07725eab8ddaeaa4d54bd0c4f

@@ -57,7 +57,7 @@ si_calculate = function(
     od = constrain_production(od, output_col, {{constraint_production}})
   }
   if (!missing(constraint_attraction)) {
-    message("Not yet implemented")
+    od = constrain_attraction(od, output_col, {{constraint_attraction}})
   }
   if (!missing(constraint_total)) {
     od = constrain_total(od, output_col, constraint_total)
@@ -88,7 +88,7 @@ si_predict = function(
     od = constrain_production(od, output_col, {{constraint_production}})
   }
   if (!missing(constraint_attraction)) {
-    message("Not yet implemented")
+    od = constrain_attraction(od, output_col, {{constraint_attraction}})
   }
   if (!missing(constraint_total)) {
     od = constrain_total(od, output_col, constraint_total)
@@ -99,6 +99,18 @@ si_predict = function(
 constrain_production = function(od, output_col, constraint_production) {
   # todo: should the grouping var (the first column, 1) be an argument?
   od_grouped = dplyr::group_by_at(od, 1)
+  od_grouped = dplyr::mutate(
+    od_grouped,
+    {{output_col}} := .data[[output_col]] /
+      sum(.data[[output_col]]) * mean( {{constraint_production}} )
+  )
+  od = dplyr::ungroup(od_grouped)
+  od
+}
+
+constrain_attraction = function(od, output_col, constraint_production) {
+  # todo: should the grouping var (the first column, 2) be an argument?
+  od_grouped = dplyr::group_by_at(od, 2)
   od_grouped = dplyr::mutate(
     od_grouped,
     {{output_col}} := .data[[output_col]] /

### 2022-04-25 R/si_model.R
Fix up attraction constrained SIMs (testing)

@@ -108,13 +108,13 @@ constrain_production = function(od, output_col, constraint_production) {
   od
 }
 
-constrain_attraction = function(od, output_col, constraint_production) {
+constrain_attraction = function(od, output_col, constraint_attraction) {
   # todo: should the grouping var (the first column, 2) be an argument?
   od_grouped = dplyr::group_by_at(od, 2)
   od_grouped = dplyr::mutate(
     od_grouped,
     {{output_col}} := .data[[output_col]] /
-      sum(.data[[output_col]]) * mean( {{constraint_production}} )
+      sum(.data[[output_col]]) * mean( {{constraint_attraction}} )
   )
   od = dplyr::ungroup(od_grouped)
   od

### 2022-04-25 README.qmd
Update README to mention mobility
Thanks @luukvdmeer


Former-commit-id: 99a0772bb3957ae3d2ff737f87d95d46f862ba05

@@
 
-- The [`spflow` R package](https://github.com/LukeCe/spflow) on CRAN
+- The [`spflow` R package](https://github.com/LukeCe/spflow)
 - The [`spint` Python package](https://spint.readthedocs.io/en/latest/)
 - The [`gravity`](https://cran.r-project.org/package=gravity) R package
+- The [`mobility`](https://covid-19-mobility-data-network.github.io/mobility/index.html) R package

### 2022-04-25 README.qmd
Mention scikit-mobility
Former-commit-id: b1e2c2508482d8077850c96107c31ca3af63be0a

--- 2022-04-23 README.qmd
@@
 
-- The [`spflow` R package](https://github.com/LukeCe/spflow) on CRAN
+- The [`spflow` R package](https://github.com/LukeCe/spflow)
 - The [`spint` Python package](https://spint.readthedocs.io/en/latest/)
 - The [`gravity`](https://cran.r-project.org/package=gravity) R package
+- The [`mobility`](https://covid-19-mobility-data-network.github.io/mobility/index.html) R package
+- The [gravity functions in the scikit-mobility](https://scikit-mobility.github.io/scikit-mobility/reference/models.html#module-skmob.models.gravity) Python package

### 2022-04-25 data-raw/si-classes.Rmd
Add docs to test compatibility with other classes for #13

---
title: "Using the si package with different input classes"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# remotes::install_github("robinlovelace/si")
library(si)
# ?si_predict
```

```{r}
od = si_to_od(si_zones, si_zones, max_dist = 4000)
m = lm(od$origin_all ~ od$origin_bicycle)
od_updated = si_predict(od, m)
class(od_updated)
od_dt = data.table::data.table(od)
class(od_dt)
od_updated_dt = si_predict(od_dt, m)
class(od_updated_dt)
identical(od_updated$interaction, od_updated_dt$interaction)
```

### 2022-04-25 R/data.R
Add new example dataset for #3

@@ -24,3 +24,15 @@ NULL
 #' @examples
 #' head(si_od_census)
 NULL
+
+#' Example destinations dataset: pubs in Leeds
+#'
+#' Example dataset from Leeds, UK
+#' @note Regenerate the data with scripts in the `data-raw` directory.
+#'
+#' @docType data
+#' @keywords datasets
+#' @name si_pubs
+#' @examples
+#' head(si_pubs)
+NULL

### 2022-04-25 data-raw/zones.R

@@ -1,6 +1,8 @@
 ## code to prepare `zones` dataset goes here
 library(pct)
 library(tidyverse)
+library(tmap)
+tmap_mode("view")
 zones = get_pct_zones(region = "west-yorkshire", geography = "msoa")
 centroids = get_pct_centroids(region = "west-yorkshire", geography = "msoa")
 si_zones = zones %>% 
@@ -17,3 +19,13 @@ si_od_census = od_census %>%
   rename(O = geo_code1, D = geo_code2) %>% 
   filter(O %in% si_zones$geo_code & D %in% si_zones$geo_code)
 usethis::use_data(si_od_census, overwrite = TRUE)
+
+# Get pubs in Leeds
+source("https://github.com/cyipt/actdev/raw/main/code/get_pois.R")
+q = "SELECT * FROM 'points' WHERE amenity IN ('pub')"
+pubs_west_yorkshire = get_pois(region_name = "west yorkshire", q = q)
+pubs_leeds = pubs_west_yorkshire[si::si_zones, ]
+plot(pubs_leeds)
+si_pubs = pubs_leeds %>% 
+  select(-other_tags)
+usethis::use_data(si_pubs)

### 2022-04-25 README.qmd
Document destinations in README, close #3

@@
   constraint_production = origin_all,
-  beta = 0.3
+  beta = 0.9
   )
@@
 
-Note: support for OD datasets in which the origins and destinations are different objects is work in progress (see [#3](https://github.com/Robinlovelace/si/issues/3)).
+The approach works equally well when origin and destination points are different, as shown in the following example which estimates the number of trips to pubs in Leeds, assuming that for every trip to the pub there are 50 trips to work:
+
+```{r pubs}
+zones_pubs = si_zones %>% 
+  mutate(to_pubs = all / 50)
+pubs_example = si_pubs %>% 
+  filter(grepl(pattern = "Chemic|Nag", x = name))
+pubs_example$size = c(100, 80)
+od_to_pubs = si_to_od(zones_pubs, pubs_example)
+od_to_pubs_result = od_to_pubs %>% 
+  si_calculate(fun = gravity_model, 
+               m = origin_to_pubs,
+               n = destination_size,
+               d = distance_euclidean,
+               beta = 0.5,
+               constraint_production = origin_to_pubs)
+od_to_pubs_result %>% 
+  select(O, D, destination_name, interaction)
+```
+
+We can plot the top 20 desire lines between zone centroids and the 2 pubs in the example dataset as follows:
+
+```{r pubmap}
+library(tmap)
+tmap_mode("view")
+od_to_pubs_result %>% 
+  top_n(n = 50, wt = interaction) %>% 
+  tm_shape() +
+  tm_lines(col = "interaction", palette = "viridis", scale = 2)
+```
 

### 2022-04-25 README.qmd
Tweak readem

@@
 
-The goal of {si} is to provide a simple, '[tidy](https://dplyr.tidyverse.org/)', and flexible framework for developing spatial interaction models (SIMs), which estimate the amount of movement between spatial entities.
+The goal of {si} is to provide a simple, [tidy](https://www.tidyverse.org/), and flexible framework for developing spatial interaction models (SIMs).
+SIMs estimate the amount of movement between spatial entities and can be used for many things, including to support evidence-based investment in sustainable transport infrastructure and prioritisation of location options for public services.
 

### 2022-05-15 R/si_to_od.R
Document bipartite SIMs, close #17
Former-commit-id: 523d3f2aea9f5db01c73f1e63e66f443d6dd9c33

@@ -2,8 +2,20 @@
 #'
 #' Prepares an OD data frame that next could be used to
 #' estimate movement between origins and destinations 
-#' with a spatial interaction model
-#'
+#' with a spatial interaction model.
+#' 
+#' In most origin-destination datasets the spatial entities that constitute
+#' origins (typically administrative zones) also represent destinations.
+#' In this 'unipartite' case `origins` and `destinations` should be passed
+#' the same object, an `sf` data frame representing administrative zones.
+#' 
+#' 'Bipartite' datasets, by contrast, represent
+#' "spatial interaction systems where origins cannot act as
+#' destinations and vice versa" (Hasova et al. [2022](https://lenkahas.com/files/preprint.pdf)). 
+#' 
+#'  a different
+#' `sf` object can be passed to the `destinations` argument.
+#' 
 #' @param origins `sf` object representing origin locations/zones
 #' @param destinations `sf` object representing destination locations/zones
 #' @param max_dist Euclidean distance in meters (numeric).

### 2022-05-15 README.qmd
Document bipartite SIMs, close #17

@@
 SIMs estimate the amount of movement between spatial entities and can be used for many things, including to support evidence-based investment in sustainable transport infrastructure and prioritisation of location options for public services.
+
+Unlike many software tools designed to support spatial interaction modelling, {si} does not define (or even encourage use of) any particular functional forms or modelling frameworks for predicting movement between origins and destinations.
+Instead, it provides a framework enabling you to use model function forms or models of your choosing, ensuring flexibility and encouraging flexibility.
 
@@
 
-The output `si_calculate)` is a geographic object which can be plotted as a map:
+The output of `si_calculate()` is a geographic object that can be plotted with `sf`'s plot method (or other geographic data visualisation packages):
 
@@
 
-```{r}
-names(od)
-```
+As documented in the [si vignette](https://robinlovelace.github.io/si/articles/si.html), the function allows you to use any variable in the origin or destination data by joining all attributes onto the OD data frame, with column names appended with `origin` and `destination`.
 
-As shown in the output above, the function allows you to use any variable in the origin or destination data in the function, with column names appended with `origin` and `destination`.
-
-The approach works equally well when origin and destination points are different, as shown in the following example which estimates the number of trips to pubs in Leeds, assuming that for every trip to the pub there are 50 trips to work:
+The approach works equally well for 'bipartite' SIMs, in which origin and destination points are different (Hasova et al. [2022](https://lenkahas.com/files/preprint.pdf)).
+The following example implements a bipartite SIM that estimates the number of trips from administrative zones to pubs in Leeds:
 
 ```{r pubs}
+# Set n. trips to pubs, assuming that for every trip to the pub there are
+# 50 trips to work (this would be validated/tested/modelled in empirical work)
 zones_pubs = si_zones %>% 
@@
 
-```{r pubmap}
+```{r pubmap, eval=FALSE}
 library(tmap)
@@
 ```
+
+![](https://user-images.githubusercontent.com/1825120/168485930-1adf9970-7dc1-4417-891e-2d5a1ec5e170.png)
 

### 2022-06-13 README.qmd
Update name {si} -> {sim}

@@
 
-The goal of {si} is to provide a simple, [tidy](https://www.tidyverse.org/), and flexible framework for developing spatial interaction models (SIMs).
+The goal of {sim} is to provide a simple, [tidy](https://www.tidyverse.org/), and flexible framework for developing spatial interaction models (SIMs).
 SIMs estimate the amount of movement between spatial entities and can be used for many things, including to support evidence-based investment in sustainable transport infrastructure and prioritisation of location options for public services.
 
-Unlike many software tools designed to support spatial interaction modelling, {si} does not define (or even encourage use of) any particular functional forms or modelling frameworks for predicting movement between origins and destinations.
+Unlike many software tools designed to support spatial interaction modelling, {sim} does not define (or even encourage use of) any particular functional forms or modelling frameworks for predicting movement between origins and destinations.
 Instead, it provides a framework enabling you to use model function forms or models of your choosing, ensuring flexibility and encouraging flexibility.
@@
 
-The approach is also 'tidy', allowing use of {si} functions in {dplyr} pipelines:
+The approach is also 'tidy', allowing use of {sim} functions in {dplyr} pipelines:
 
@@
 
-To dive straight into using {si} to develop SIMs, see the [`si` Get started vignette](https://robinlovelace.github.io/si/articles/si.html).
+To dive straight into using {sim} to develop SIMs, see the [`si` Get started vignette](https://robinlovelace.github.io/si/articles/si.html).
 

### 2022-06-13 README.qmd
Update name in library calls #20

@@
 ```{r distance}
-library(si)
+library(sim)
 library(dplyr)

### 2022-06-13 README.qmd
Update all URLs #20

@@
 <!-- badges: start -->
-[![R-CMD-check](https://github.com/robinlovelace/si/workflows/R-CMD-check/badge.svg)](https://github.com/robinlovelace/si/actions)
+[![R-CMD-check](https://github.com/Robinlovelace/sim/workflows/R-CMD-check/badge.svg)](https://github.com/Robinlovelace/sim/actions)
 <!-- badges: end -->
@@
 ```r
-remotes::install_github("robinlovelace/si")
+remotes::install_github("Robinlovelace/sim")
 ```

### 2022-06-13 README.qmd
Update urls

--- 2022-06-13 README.qmd
@@
 <!-- badges: start -->
-[![R-CMD-check](https://github.com/robinlovelace/si/workflows/R-CMD-check/badge.svg)](https://github.com/robinlovelace/si/actions)
+[![R-CMD-check](https://github.com/Robinlovelace/simm/workflows/R-CMD-check/badge.svg)](https://github.com/Robinlovelace/simm/actions)
 <!-- badges: end -->
@@
 ```r
-remotes::install_github("robinlovelace/si")
+remotes::install_github("Robinlovelace/simm")
 ```
@@
 
-As documented in the [si vignette](https://robinlovelace.github.io/si/articles/si.html), the function allows you to use any variable in the origin or destination data by joining all attributes onto the OD data frame, with column names appended with `origin` and `destination`.
+As documented in the [si vignette](https://robinlovelace.github.io/sim/articles/sim.html), the function allows you to use any variable in the origin or destination data by joining all attributes onto the OD data frame, with column names appended with `origin` and `destination`.
 
@@
 
-For details on what SIMs are and how they have been defined mathematically and in code from first principles, see the [`sims` vignette](https://robinlovelace.github.io/si/articles/sims-first-principles.html).
+For details on what SIMs are and how they have been defined mathematically and in code from first principles, see the [`sims` vignette](https://robinlovelace.github.io/sim/articles/simms-first-principles.html).
 
-To dive straight into using {sim} to develop SIMs, see the [`si` Get started vignette](https://robinlovelace.github.io/si/articles/si.html).
+To dive straight into using {sim} to develop SIMs, see the [`si` Get started vignette](https://robinlovelace.github.io/sim/articles/sim.html).
 

### 2022-06-13 README.qmd
Change name, close #20

@@
 <!-- # Implementations in other languages -->
-## si basics
+## sim basics
 
@@
 
-As documented in the [si vignette](https://robinlovelace.github.io/sim/articles/sim.html), the function allows you to use any variable in the origin or destination data by joining all attributes onto the OD data frame, with column names appended with `origin` and `destination`.
+As documented in the [sim vignette](https://robinlovelace.github.io/sim/articles/sim.html), the function allows you to use any variable in the origin or destination data by joining all attributes onto the OD data frame, with column names appended with `origin` and `destination`.
 

### 2022-06-13 data-raw/zones.R

@@ -24,7 +24,7 @@ usethis::use_data(si_od_census, overwrite = TRUE)
 source("https://github.com/cyipt/actdev/raw/main/code/get_pois.R")
 q = "SELECT * FROM 'points' WHERE amenity IN ('pub')"
 pubs_west_yorkshire = get_pois(region_name = "west yorkshire", q = q)
-pubs_leeds = pubs_west_yorkshire[si::si_zones, ]
+pubs_leeds = pubs_west_yorkshire[sim::si_zones, ]
 plot(pubs_leeds)
 si_pubs = pubs_leeds %>% 
   select(-other_tags)

### 2022-06-13 data-raw/si-classes.Rmd
Update name in library calls #20
Update all URLs #20
Update urls
Rename pkg with changer, close #20
Former-commit-id: 6c6fe93a3cbf6dac23e5eae172f3e6a4e4332fba

@@
 ---
-title: "Using the si package with different input classes"
+title: "Using the sim package with different input classes"
 output: github_document
@@
 ```{r setup}
-# remotes::install_github("robinlovelace/si")
-library(si)
+# remotes::install_github("robinlovelace/sim")
+library(sim)
 # ?si_predict

### 2022-06-13 R/si_model.R

@@ -64,7 +64,7 @@ si_calculate = function(
   }
   od
 }
-#' Predict si model based on pre-trained model
+#' Predict sim model based on pre-trained model
 #' 
 #' @param model A model object, e.g. from [lm()] or [glm()]
 #' @inheritParams si_calculate

### 2022-06-13 data-raw/sim-paper.Rmd
Add draft paper content
Former-commit-id: aa491d3dc66d72527c90832a82313d39b48b6086

---
# title: "sim: A Package for Developing Spatial Interaction Models"
# output: rmarkdown::html_vignette
# vignette: >
#   %\VignetteIndexEntry{sim: A Package for Developing Spatial Interaction Models}
#   %\VignetteEngine{knitr::rmarkdown}
#   %\VignetteEncoding{UTF-8}

bibliography: paper.bib

output: rticles::jss_article
documentclass: jss
author:
  - name: FirstName LastName
    orcid: 0000-0000-0000-0000
    affiliation: University/Company
    # use this syntax to add text on several lines
    address: |
      | First line
      | Second line
    email: \email{name@company.com}
    url: http://rstudio.com
  - name: Second Author
    orcid: 0000-0000-0000-0000
    affiliation: 'Affiliation \AND'
    # To add another line, use \AND at the end of the previous one as above
  - name: Third Author
    orcid: 0000-0000-0000-0000
    address: |
      | Department of Statistics and Mathematics,
      | Faculty of Biosciences,
      | Universitat Autònoma de Barcelona
    affiliation: |
      | Universitat Autònoma
      | de Barcelona
    # use a different affiliation in adress field (differently formated here)
    affiliation2: Universitat Autònoma de Barcelona
title:
  formatted: "\\pkg{sim}: A Package for Developing Spatial Interaction Models"
  # If you use tex in the formatted title, also supply version without
  plain:     "sim: A Package for Developing Spatial Interaction Models"
  # For running headers, if needed
  short:     "\\pkg{sim}: Spatial Interaction Models"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [spatial interaction, movement, "\\proglang{R}"]
  plain:     [spatial interaction, movement, R]
preamble: >
  \usepackage{amsmath}
---

```{r echo=FALSE, eval=FALSE}
rbbt::bbt_update_bib(path_rmd = "vignettes/paper.Rmd", "vignettes/paper.bib")
```

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
```

# Introduction

Spatial Interaction Models (SIMs) are mathematical models for estimating movement between spatial entities.
First developed by Alan Wilson in the late 1960s and early 1970, SIMs have become a key tool for transport modelling with substantial practical applications [@boyce_forecasting_2015].
There are four main types of traditional SIMs [@wilson_family_1971]:

-   Unconstrained

-   Production-constrained

-   Attraction-constrained

-   Doubly-constrained

An early and highly influential type of SIM was the 'gravity model', defined by @wilson_family_1971 as follows (in a paper that explored many iterations on this formulation):

$$
T_{i j}=K \frac{W_{i}^{(1)} W_{j}^{(2)}}{c_{i j}^{n}}
$$ "where $T_{i j}$ is a measure of the interaction between zones $i$ and $W_{i}^{(1)}$ is a measure of the 'mass term' associated with zone $z_i$, $W_{j}^{(2)}$ is a measure of the 'mass term' associated with zone $z_j$, and $c_{ij}$ is a measure of the distance, or generalised cost of travel, between zone $i$ and zone $j$".
$K$ is a 'constant of proportionality' and $n$ is a parameter to be estimated.

Redefining the $W$ terms as $m$ and $n$ for origins and destinations respectively [@simini_universal_2012], this classic definition of the 'gravity model' can be written as follows:

$$
T_{i j}=K \frac{m_{i} n_{j}}{c_{i j}^{n}}
$$

For the purposes of this project, we will focus on production-constrained SIMs.
These can be defined as follows [@wilson_family_1971]:

$$
T_{ij} = A_iO_in_jf(c_{ij})
$$

where $A$ is a balancing factor defined as:

$$
A_{i}=\frac{1}{\sum_{j} m_{j} \mathrm{f}\left(c_{i j}\right)}
$$

$O_i$ is analogous to the travel demand in zone $i$, which can be roughly approximated by its population.

More recent innovations in SIMs including the 'radiation model' @simini_universal_2012.
See @lenormand_systematic_2016 for a comparison of alternative approaches.

## Code formatting

In general, don't use Markdown, but use the more precise LaTeX commands instead:

-   \proglang{Java}

-   \pkg{plyr}

One exception is inline code, which can be written inside a pair of backticks (i.e., using the Markdown syntax).

If you want to use LaTeX commands in headers, you need to provide a `short-title` attribute.
You can also provide a custom identifier if necessary.
See the header of Section \ref{r-code} for example.

# \proglang{R} code {#r-code short-title="R code"}

Can be inserted in regular R markdown blocks.

```{r}
x <- 1:10
x
```

## Features specific to \pkg{rticles} {short-title="Features specific to rticles"}

-   Adding short titles to section headers is a feature specific to \pkg{rticles} (implemented via a Pandoc Lua filter). This feature is currently not supported by Pandoc and we will update this template if [it is officially supported in the future](https://github.com/jgm/pandoc/issues/4409).
-   Using the `\AND` syntax in the `author` field to add authors on a new line. This is a specific to the `rticles::jss_article` format.

### 2022-06-13 data-raw/si-classes.Rmd
Update name again, close #20 again

--- 2022-04-25 data-raw/si-classes.Rmd
@@
 ---
-title: "Using the si package with different input classes"
+title: "Using the simodels package with different input classes"
 output: github_document
@@
 ```{r setup}
-# remotes::install_github("robinlovelace/si")
-library(si)
+# remotes::install_github("robinlovelace/simodels")
+library(simodels)
 # ?si_predict

### 2022-06-13 data-raw/sim-paper.Rmd
Update name again, close #20 again

@@
 ---
-# title: "sim: A Package for Developing Spatial Interaction Models"
+# title: "simodels: A Package for Developing Spatial Interaction Models"
 # output: rmarkdown::html_vignette
 # vignette: >
-#   %\VignetteIndexEntry{sim: A Package for Developing Spatial Interaction Models}
+#   %\VignetteIndexEntry{simodels: A Package for Developing Spatial Interaction Models}
 #   %\VignetteEngine{knitr::rmarkdown}
@@
 title:
-  formatted: "\\pkg{sim}: A Package for Developing Spatial Interaction Models"
+  formatted: "\\pkg{simodels}: A Package for Developing Spatial Interaction Models"
   # If you use tex in the formatted title, also supply version without
-  plain:     "sim: A Package for Developing Spatial Interaction Models"
+  plain:     "simodels: A Package for Developing Spatial Interaction Models"
   # For running headers, if needed
-  short:     "\\pkg{sim}: Spatial Interaction Models"
+  short:     "\\pkg{simodels}: Spatial Interaction Models"
 abstract: >

### 2022-06-13 R/si_model.R

@@ -64,7 +64,7 @@ si_calculate = function(
   }
   od
 }
-#' Predict sim model based on pre-trained model
+#' Predict simodels model based on pre-trained model
 #' 
 #' @param model A model object, e.g. from [lm()] or [glm()]
 #' @inheritParams si_calculate

### 2022-06-13 data-raw/zones.R

@@ -24,7 +24,7 @@ usethis::use_data(si_od_census, overwrite = TRUE)
 source("https://github.com/cyipt/actdev/raw/main/code/get_pois.R")
 q = "SELECT * FROM 'points' WHERE amenity IN ('pub')"
 pubs_west_yorkshire = get_pois(region_name = "west yorkshire", q = q)
-pubs_leeds = pubs_west_yorkshire[sim::si_zones, ]
+pubs_leeds = pubs_west_yorkshire[simodels::si_zones, ]
 plot(pubs_leeds)
 si_pubs = pubs_leeds %>% 
   select(-other_tags)

### 2022-06-13 README.qmd
Remove final mentions of si?
Update readme #20

--- 2022-06-13 README.qmd
@@
 <!-- badges: start -->
-[![R-CMD-check](https://github.com/Robinlovelace/sim/workflows/R-CMD-check/badge.svg)](https://github.com/Robinlovelace/sim/actions)
+[![R-CMD-check](https://github.com/robinlovelace/simodels/workflows/R-CMD-check/badge.svg)](https://github.com/robinlovelace/simodels/actions)
 <!-- badges: end -->
 
-The goal of {sim} is to provide a simple, [tidy](https://www.tidyverse.org/), and flexible framework for developing spatial interaction models (SIMs).
+The goal of {simodels} is to provide a simple, [tidy](https://www.tidyverse.org/), and flexible framework for developing spatial interaction models (SIMs).
 SIMs estimate the amount of movement between spatial entities and can be used for many things, including to support evidence-based investment in sustainable transport infrastructure and prioritisation of location options for public services.
 
-Unlike many software tools designed to support spatial interaction modelling, {sim} does not define (or even encourage use of) any particular functional forms or modelling frameworks for predicting movement between origins and destinations.
+Unlike many software tools designed to support spatial interaction modelling, {simodels} does not define (or even encourage use of) any particular functional forms or modelling frameworks for predicting movement between origins and destinations.
 Instead, it provides a framework enabling you to use model function forms or models of your choosing, ensuring flexibility and encouraging flexibility.
@@
 ```r
-remotes::install_github("Robinlovelace/sim")
+remotes::install_github("robinlovelace/simodels")
 ```
@@
 <!-- # Implementations in other languages -->
-## si basics
+## simodel basics
 
@@
 ```{r distance}
-library(sim)
+library(simodels)
 library(dplyr)
@@
 
-The approach is also 'tidy', allowing use of {sim} functions in {dplyr} pipelines:
+The approach is also 'tidy', allowing use of {simodels} functions in {dplyr} pipelines:
 
@@
 
-As documented in the [si vignette](https://robinlovelace.github.io/si/articles/si.html), the function allows you to use any variable in the origin or destination data by joining all attributes onto the OD data frame, with column names appended with `origin` and `destination`.
+The function allows you to use any variable in the origin or destination data by joining all attributes onto the OD data frame, with column names appended with `origin` and `destination`.
 
@@
 
-For details on what SIMs are and how they have been defined mathematically and in code from first principles, see the [`sims` vignette](https://robinlovelace.github.io/si/articles/sims-first-principles.html).
+For details on what SIMs are and how they have been defined mathematically and in code from first principles, see the [`sims` vignette](https://robinlovelace.github.io/simodels/articles/sims-first-principles.html).
 
-To dive straight into using {sim} to develop SIMs, see the [`si` Get started vignette](https://robinlovelace.github.io/si/articles/si.html).
+To dive straight into using {simodels} to develop SIMs, see the [`si` Get started vignette](https://robinlovelace.github.io/simodels/articles/si.html).
 

### 2022-06-13 README.qmd
Document

@@
 
-To dive straight into using {simodels} to develop SIMs, see the [`si` Get started vignette](https://robinlovelace.github.io/simodels/articles/si.html).
+To dive straight into using {simodels} to develop SIMs, see the [Get started vignette](https://robinlovelace.github.io/simodels/articles/simodels.html).
 

### 2022-06-13 R/si_model.R

@@ -64,7 +64,7 @@ si_calculate = function(
   }
   od
 }
-#' Predict simodels model based on pre-trained model
+#' Predict spatial interaction based on pre-trained model
 #' 
 #' @param model A model object, e.g. from [lm()] or [glm()]
 #' @inheritParams si_calculate

### 2022-06-13 R/si_model.R
Add @return values for CRAN https://github.com/ThinkR-open/prepare-for-cran

@@ -21,6 +21,7 @@
 #' @param ... Arguments passed to `fun`
 #' @param output_col Character string containing the name of the new output
 #'   column. `"interaction"` by default.
+#' @return An sf data frame
 #' @export
 #' @importFrom rlang .data
 #' @examples
@@ -69,6 +70,7 @@ si_calculate = function(
 #' @param model A model object, e.g. from [lm()] or [glm()]
 #' @inheritParams si_calculate
 #' @seealso si_calculate
+#' @return An sf data frame
 #' @export 
 #' @examples
 #' od = si_to_od(si_zones, si_zones, max_dist = 4000)

### 2022-06-13 R/si_to_od.R

@@ -25,6 +25,7 @@
 #'   Intrazonal OD pairs represent movement from one
 #'   place in a zone to another place in the same zone.
 #'   `TRUE` by default.
+#' @return An sf data frame
 #' @export
 #' @examples
 #' library(sf)

### 2022-08-31 R/data.R
Add aus_ datasets, close #28

@@ -36,3 +36,27 @@ NULL
 #' @examples
 #' head(si_pubs)
 NULL
+
+#' Example zones dataset: regions of Australia
+#'
+#' Example dataset from Australia
+#' @note Regenerate the data with scripts in the `data-raw` directory.
+#'
+#' @docType data
+#' @keywords datasets
+#' @name zones_aus
+#' @examples
+#' head(zones_aus)
+NULL
+
+#' Example OD dataset: flows between regions in Australia
+#'
+#' Example dataset from Australia
+#' @note Regenerate the data with scripts in the `data-raw` directory.
+#'
+#' @docType data
+#' @keywords datasets
+#' @name od_aus
+#' @examples
+#' head(od_aus)
+NULL

### 2024-02-01 data-raw/si_model.R
Tests

@@ -41,7 +41,6 @@ od_pconst = si_calculate(od, fun = fun, beta = 0.3, O = "origin_all", n = "desti
 plot(od_pconst$distance_euclidean, od_pconst$res)
 plot(od_pconst["res"], logz = TRUE)
 si_model = function(od, fun, var_p, ...) {
-  browser()
   res = fun(...)
   od = dplyr::mutate(od, res = fun(...))
   od$res = fun(od, ...)

### 2024-02-01 R/si_model.R
Fix part of #31: output_col removed

@@ -53,7 +53,7 @@ si_calculate = function(
     ...
     ) {
   dots = rlang::enquos(...)
-  od = dplyr::mutate(od, {{output_col}} := fun(!!!dots))
+  od = dplyr::mutate(od, "{output_col}" := fun(!!!dots))
   if (!missing(constraint_production)) {
     od = constrain_production(od, output_col, {{constraint_production}})
   }
@@ -103,7 +103,7 @@ constrain_production = function(od, output_col, constraint_production) {
   od_grouped = dplyr::group_by_at(od, 1)
   od_grouped = dplyr::mutate(
     od_grouped,
-    {{output_col}} := .data[[output_col]] /
+    "{output_col}" := .data[[output_col]] /
       sum(.data[[output_col]]) * mean( {{constraint_production}} )
   )
   od = dplyr::ungroup(od_grouped)
@@ -115,7 +115,7 @@ constrain_attraction = function(od, output_col, constraint_attraction) {
   od_grouped = dplyr::group_by_at(od, 2)
   od_grouped = dplyr::mutate(
     od_grouped,
-    {{output_col}} := .data[[output_col]] /
+    "{output_col}" := .data[[output_col]] /
       sum(.data[[output_col]]) * mean( {{constraint_attraction}} )
   )
   od = dplyr::ungroup(od_grouped)

### 2024-02-01 R/si_model.R
Update with faster constrain_production

@@ -104,8 +104,24 @@ constrain_production = function(od, output_col, constraint_production) {
   od_grouped = dplyr::mutate(
     od_grouped,
     "{output_col}" := .data[[output_col]] /
-      sum(.data[[output_col]]) * mean( {{constraint_production}} )
+      sum(.data[[output_col]]) * first( {{constraint_production}} )
   )
+  # # With base R and aggregate
+  # output_value = aggregate(
+  #   od[[eval(output_col)]],
+  #   by = list(od[[1]]),
+  #   FUN = function(x) x / sum(x) * od[[]][1]
+  # )
+  # od_grouped |>
+  #   select(origin_all, interaction)
+  # # Assert values are correct for test data
+  # od_grouped |>
+  #   sf::st_drop_geometry() |>
+  #   # group_by(1) |>
+  #   summarise(
+  #     sum = sum(interaction),
+  #     first = first(origin_all)
+  #   )
   od = dplyr::ungroup(od_grouped)
   od
 }

### 2024-02-01 R/si_model.R
Tidy-up

@@ -106,15 +106,9 @@ constrain_production = function(od, output_col, constraint_production) {
     "{output_col}" := .data[[output_col]] /
       sum(.data[[output_col]]) * first( {{constraint_production}} )
   )
-  # # With base R and aggregate
-  # output_value = aggregate(
-  #   od[[eval(output_col)]],
-  #   by = list(od[[1]]),
-  #   FUN = function(x) x / sum(x) * od[[]][1]
-  # )
+  # # Assert values are correct for test data:
   # od_grouped |>
   #   select(origin_all, interaction)
-  # # Assert values are correct for test data
   # od_grouped |>
   #   sf::st_drop_geometry() |>
   #   # group_by(1) |>

### 2024-02-01 R/si_model.R
Minor fixes, tests pass

@@ -104,7 +104,7 @@ constrain_production = function(od, output_col, constraint_production) {
   od_grouped = dplyr::mutate(
     od_grouped,
     "{output_col}" := .data[[output_col]] /
-      sum(.data[[output_col]]) * first( {{constraint_production}} )
+      sum(.data[[output_col]]) * dplyr::first( {{constraint_production}} )
   )
   # # Assert values are correct for test data:
   # od_grouped |>
@@ -126,10 +126,8 @@ constrain_attraction = function(od, output_col, constraint_attraction) {
   od_grouped = dplyr::mutate(
     od_grouped,
     "{output_col}" := .data[[output_col]] /
-      sum(.data[[output_col]]) * mean( {{constraint_attraction}} )
+      sum(.data[[output_col]]) * dplyr::first( {{constraint_attraction}} )
   )
-  od = dplyr::ungroup(od_grouped)
-  od
 }
 
 constrain_total = function(od, output_col, constraint_total) {

### 2024-02-01 R/poinds_to_od_maxdist.R
Working version of si_to_od

@@ -1,0 +1,59 @@
+#' COnvert points to OD with a maxiumum distance
+#'
+#' 
+#' @param p A spatial points object or a matrix of coordinates representing points
+#' @param pd Optional spatial points object or matrix objects representing destinations
+#' @param max_dist Maximum distance in metres
+#' @param max_dest The maximum number of destinations for each origin (numeric).
+#'   Default is Inf. Alternative to max_dist for limiting the number of ODs.
+#' @param intrazonal Should the result only include interzonal OD pairs, in which
+#' the ID of the origin is different from the ID of the destination zone?
+#' `FALSE` by default
+#' @param ids_only Should a data frame with only 2 columns (origin and destination IDs)
+#' be returned? The default is `FALSE`, meaning the result should also contain the
+#' coordinates of the start and end points of each OD pair.
+#' @return An sf data frame
+#' @export
+
+points_to_od_maxdist = function(p, pd = NULL, max_dist = Inf, max_dest = Inf, intrazonal = FALSE, ids_only = FALSE){
+  
+  if(any(sf::st_geometry_type(p) != "POINT")){
+    message("Converting to centroids")
+    suppressWarnings(p <- sf::st_centroid(p))
+  }
+  
+  if(!is.null(pd)){
+    if(any(sf::st_geometry_type(pd) != "POINT")){
+      suppressWarnings(pd <- sf::st_centroid(pd))
+    }
+  } else {
+    pd = p
+  }
+  
+  if(max_dest > nrow(pd)){
+    max_dest = nrow(pd)
+  }
+  
+  nn <- nngeo::st_nn(p, pd, k = max_dest, maxdist = max_dist)
+  res = data.frame(O = rep(p[[1]], lengths(nn)),
+                   D = pd[[1]][unlist(nn, use.names = FALSE)])
+  
+  if(intrazonal){
+    res = res[res$O != res$D,]
+  }
+  
+  if(ids_only){
+    return(res)
+  }
+  
+  # Add Coords
+  p_coords = as.data.frame(sf::st_coordinates(p))
+  names(p_coords) = c("X1","Y1")
+  
+  pd_coords = as.data.frame(sf::st_coordinates(pd))
+  names(pd_coords) = c("X2","Y2")
+  
+  res = cbind(res, p_coords[match(res$O, p[[1]]),])
+  res = cbind(res, pd_coords[match(res$D, pd[[1]]),])
+  res
+}

### 2024-02-01 R/si_model.R

@@ -55,13 +55,13 @@ si_calculate = function(
   dots = rlang::enquos(...)
   od = dplyr::mutate(od, "{output_col}" := fun(!!!dots))
   if (!missing(constraint_production)) {
-    od = constrain_production(od, output_col, {{constraint_production}})
+    od[[output_col]] = constrain(od[[1]], od[[output_col]], od[[constraint_production]])
   }
   if (!missing(constraint_attraction)) {
-    od = constrain_attraction(od, output_col, {{constraint_attraction}})
+    od[[output_col]] = constrain(od[[2]], od[[output_col]], od[[constraint_production]])
   }
   if (!missing(constraint_total)) {
-    od = constrain_total(od, output_col, constraint_total)
+    od[[output_col]] = constrain_total(od, output_col, constraint_total)
   }
   od
 }
@@ -87,10 +87,10 @@ si_predict = function(
 ) {
   od[[output_col]] = stats::predict(model, od)
   if (!missing(constraint_production)) {
-    od = constrain_production(od, output_col, {{constraint_production}})
+    od[[output_col]] = constrain(od[[1]], od[[output_col]], od[[constraint_production]])
   }
   if (!missing(constraint_attraction)) {
-    od = constrain_attraction(od, output_col, {{constraint_attraction}})
+    od[[output_col]] = constrain(od[[1]], od[[output_col]], od[[constraint_attraction]])
   }
   if (!missing(constraint_total)) {
     od = constrain_total(od, output_col, constraint_total)
@@ -98,38 +98,10 @@ si_predict = function(
   od
 }
 
-constrain_production = function(od, output_col, constraint_production) {
-  # todo: should the grouping var (the first column, 1) be an argument?
-  od_grouped = dplyr::group_by_at(od, 1)
-  od_grouped = dplyr::mutate(
-    od_grouped,
-    "{output_col}" := .data[[output_col]] /
-      sum(.data[[output_col]]) * first( {{constraint_production}} )
-  )
-  # # Assert values are correct for test data:
-  # od_grouped |>
-  #   select(origin_all, interaction)
-  # od_grouped |>
-  #   sf::st_drop_geometry() |>
-  #   # group_by(1) |>
-  #   summarise(
-  #     sum = sum(interaction),
-  #     first = first(origin_all)
-  #   )
-  od = dplyr::ungroup(od_grouped)
-  od
-}
-
[…]

### 2024-02-01 R/si_to_od.R

@@ -1,30 +1,29 @@
 #' Prepare OD data frame
 #'
-#' Prepares an OD data frame that next could be used to
-#' estimate movement between origins and destinations 
-#' with a spatial interaction model.
-#' 
+#' Prepares an OD data frame that next could be used to estimate movement
+#' between origins and destinations with a spatial interaction model.
+#'
 #' In most origin-destination datasets the spatial entities that constitute
-#' origins (typically administrative zones) also represent destinations.
-#' In this 'unipartite' case `origins` and `destinations` should be passed
-#' the same object, an `sf` data frame representing administrative zones.
-#' 
-#' 'Bipartite' datasets, by contrast, represent
-#' "spatial interaction systems where origins cannot act as
-#' destinations and vice versa" (Hasova et al. [2022](https://lenkahas.com/files/preprint.pdf)). 
-#' 
-#'  a different
-#' `sf` object can be passed to the `destinations` argument.
-#' 
+#' origins (typically administrative zones) also represent destinations. In this
+#' 'unipartite' case `origins` and `destinations` should be passed the same
+#' object, an `sf` data frame representing administrative zones.
+#'
+#' 'Bipartite' datasets, by contrast, represent "spatial interaction systems
+#' where origins cannot act as destinations and vice versa" (Hasova et al.
+#' [2022](https://lenkahas.com/files/preprint.pdf)).
+#'
+#' a different `sf` object can be passed to the `destinations` argument.
+#'
 #' @param origins `sf` object representing origin locations/zones
 #' @param destinations `sf` object representing destination locations/zones
-#' @param max_dist Euclidean distance in meters (numeric).
-#'   Only OD pairs that are this distance apart or less will be returned
-#'   and therefore included in the SIM.
-#' @param intrazonal Include intrazonal OD pairs?
-#'   Intrazonal OD pairs represent movement from one
-#'   place in a zone to another place in the same zone.
-#'   `TRUE` by default.
+#' @param max_dist Euclidean distance in meters (numeric). Only OD pairs that
+#'   are this distance apart or less will be returned and therefore included in
+#'   the SIM.
[…]

### 2024-02-01 R/poinds_to_od_maxdist.R
add distances to results

@@ -34,9 +34,10 @@ points_to_od_maxdist = function(p, pd = NULL, max_dist = Inf, max_dest = Inf, in
     max_dest = nrow(pd)
   }
   
-  nn <- nngeo::st_nn(p, pd, k = max_dest, maxdist = max_dist)
-  res = data.frame(O = rep(p[[1]], lengths(nn)),
-                   D = pd[[1]][unlist(nn, use.names = FALSE)])
+  nn <- nngeo::st_nn(p, pd, k = max_dest, maxdist = max_dist, returnDist = TRUE)
+  res = data.frame(O = rep(p[[1]], lengths(nn$nn)),
+                   D = pd[[1]][unlist(nn$nn, use.names = FALSE)],
+                   distance_euclidean = unlist(nn$dist, use.names = FALSE))
   
   if(intrazonal){
     res = res[res$O != res$D,]

### 2024-02-01 R/si_to_od.R

@@ -44,28 +44,30 @@
 #' nrow(odsf) # no intrazonal flows
 #' plot(odsf)
 
-si_to_od = function(origins, destinations, max_dist = 10000, max_dest = Inf, intrazonal = TRUE) {
+si_to_od = function(origins, 
+                    destinations, 
+                    max_dist = 10000, 
+                    max_dest = Inf, 
+                    intrazonal = TRUE) {
   
   od_df = points_to_od_maxdist(origins, destinations, max_dist = max_dist, intrazonal = intrazonal, max_dest = max_dest)
   
-  #message("Hi")
-  od_sfc = od::odc_to_sfc(od_df[3:6])
-  sf::st_crs(od_sfc) = 4326 # todo: add CRS argument?
-  od_df = od_df[-c(3:6)]
-  #message("Hi2")
+  od_sfc = od::odc_to_sfc(od_df[4:7])
+  sf::st_crs(od_sfc) = sf::st_crs(origins)
+  od_df = od_df[-c(4:7)]
+  
   # join origin attributes
   origins_to_join = sf::st_drop_geometry(origins)
   names(origins_to_join) = paste0("origin_", names(origins_to_join))
   names(origins_to_join)[1] = names(od_df)[1]
   od_dfj = dplyr::inner_join(od_df, origins_to_join, by = "O")
-  #message("Hi3")
+  
   # join destination attributes
   destinations_to_join = sf::st_drop_geometry(destinations)
   names(destinations_to_join) = paste0("destination_", names(destinations_to_join)) # nolint
   names(destinations_to_join)[1] = names(od_df)[2]
   od_dfj = dplyr::inner_join(od_dfj, destinations_to_join, by = "D")
-  #message("Hi4")
-  # names(od_dfj)
+
   # create and return sf object
   sf::st_sf(od_dfj, geometry = od_sfc)
 }

### 2024-02-01 README.qmd

@@
 remotes::install_github("robinlovelace/simodels")
+```
+
+To get the develoment version do:
+
+```{r}
+devtools::load_all()
 ```
@@
 - The [gravity functions in the scikit-mobility](https://scikit-mobility.github.io/scikit-mobility/reference/models.html#module-skmob.models.gravity) Python package
+
+```{r}
+#| echo: false
+#| eval: false
+# Benmarks
+bench::mark(
+  {od_res = si_calculate(
+  od,
+  fun = gravity_model,
+  d = distance_euclidean,
+  m = origin_all,
+  n = destination_all,
+  constraint_production = origin_all,
+  beta = 0.9
+  )}
+)
+```
+
+```{r}
+#| echo: false
+#| eval: false
+# Add actions:
+usethis::use_github_action("check-standard")
+# Add pkgdown action:
+usethis::use_github_action("pkgdown")
+```

### 2024-07-26 paper/paper.qmd
Daft paper

---
title: "simodels: A package for Spatial Interaction Modelling"
format:
    jss-pdf:
        keep-tex: true
    jss-html: default
# author:
#   - name: Achim Zeileis
#     affiliations:
#       - name: Universität Innsbruck
#         department: Department of Statistics
#         address: Universitätsstr. 15
#         city: Innsbruck
#         country: Austria
#         postal-code: 6020
#       - Journal of Statistical Software
#     orcid: 0000-0003-0918-3766
#     email: Achim.Zeileis@R-project.org
#     url: https://www.zeileis.org/
#   - name: Second  Author
#     affiliations:
#       - Plus Affiliation
# abstract: |
#   This short article illustrates how to write a manuscript for the
#   _Journal of Statistical Software_ (JSS) using its {{< latex >}} style files.
#   Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
#   it is recommended to keep the {{< latex >}} code as simple as possible,
#   i.e., avoid inclusion of packages/commands that are not necessary.
#   For outlining the typical structure of a JSS article some brief text snippets are employed that have been inspired by @ZeileisKleiberJackman2008, discussing count data regression in [R]{.proglang}. Editorial comments and instructions are marked by vertical bars.
abstract: |
  This article introduces the `simodels` package for spatial interaction modelling in R. The package provides functions for estimating spatial interaction models, including unconstrained gravity models and models that are production-, attraction-, or doubly-constrained.

keywords: [JSS, style guide, comma-separated, not capitalized, R]
keywords-formatted: [JSS, style guide, comma-separated, not capitalized, "[R]{.proglang}"]

bibliography: bibliography.bib  
---

<!-- ## Introduction: Count data regression in R {#sec-intro}

::: callout
The introduction is in principle "as usual". However, it should usually embed both the implemented *methods* and the *software* into the respective relevant literature. For the latter both competing and complementary software should be discussed (within the same software environment and beyond), bringing out relative (dis)advantages. All software mentioned should be properly `@cited`'d. (See also [Using BibTeX] for more details on {{< bibtex >}}.)

For writing about software JSS requires authors to use the markup `[]{.proglang}` (programming languages and large programmable systems), `[]{.pkg}` (software packages), back ticks like \`code\` for code (functions, commands, arguments, etc.).

If there is such markup in (sub)section titles (as above), a plain text version has to be provided in the {{< latex >}} command as well. Below we also illustrate how abbrevations should be introduced and citation commands can be employed. See the {{< latex >}} code for more details.
:::

Modeling count variables is a common task in economics and the social sciences. The classical Poisson regression model for count data is often of limited use in these disciplines because empirical count data sets typically exhibit overdispersion and/or an excess number of zeros. The former issue can be addressed by extending the plain Poisson regression model in various directions: e.g., using sandwich covariances or estimating an additional dispersion parameter (in a so-called quasi-Poisson model). Another more formal way is to use a negative binomial (NB) regression. All of these models belong to the family of generalized linear models (GLMs). However, although these models typically can capture overdispersion rather well, they are in many applications not sufficient for modeling excess zeros. Since @Mullahy1986 there is increased interest in zero-augmented models that address this issue by a second model component capturing zero counts. An overview of count data models in econometrics, including hurdle and zero-inflated models, is provided in @CameronTrivedi2013.

In [R]{.proglang} @R, GLMs are provided by the model fitting functions [glm]{.fct} in the [stats]{.pkg} package and [glm.nb]{.fct} in the [MASS]{.pkg} package [@VenablesRipley2002] along with associated methods for diagnostics and inference. The manuscript that this document is based on [@ZeileisKleiberJackman2008] then introduced hurdle and zero-inflated count models in the functions [hurdle]{.fct} and [zeroinfl]{.fct} in the [pcsl]{.pkg} package @Jackman2015. Of course, much more software could be discussed here, including (but not limited to) generalized additive models for count data as available in the [R]{.proglang} packages [mgcv]{.pkg} @Wood2006, [gamlss]{.pkg} @StasinopoulosRigby2007, or [VGAM]{.pkg} @Yee2009.

## Models and software {#sec-models}

The basic Poisson regression model for count data is a special case of the GLM framework @McCullaghNelder1989. It describes the dependence of a count response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution $y_i \sim \mathrm{Pois}(\mu_i)$. The dependence of the conditional mean $E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a log link and a linear predictor

$$
\log(\mu_i) \quad = \quad x_i^\top \beta,
$$ {#eq-mean}

where the regression coefficients $\beta$ are estimated by maximum likelihood (ML) using the iterative weighted least squares (IWLS) algorithm.

::: callout
TODO: Note that around the equation above there should be no spaces (avoided in the {{< latex >}} code by `%` lines) so that "normal" spacing is used and not a new paragraph started.
:::

[R]{.proglang} provides a very flexible implementation of the general GLM framework in the function [glm]{.fct} @ChambersHastie1992 in the [stats]{.pkg} package. Its most important arguments are

```r
glm(formula, data, subset, na.action, weights, offset,
  family = gaussian, start = NULL, control = glm.control(…),
  model = TRUE, y = TRUE, x = FALSE, …)
```

where `formula` plus `data` is the now standard way of specifying regression relationships in [R]{.proglang}/[S]{.proglang} introduced in @ChambersHastie1992. The remaining arguments in the first line (`subset`, `na.action`, `weights`, and `offset`) are also standard for setting up formula-based regression models in [R]{.proglang}/[S]{.proglang}. The arguments in the second line control aspects specific to GLMs while the arguments in the last line specify which components are returned in the fitted model object (of class [glm]{.class} which inherits from [lm]{.class}). For further arguments to [glm]{.fct} (including alternative specifications of starting values) see `?glm`. For estimating a Poisson model `family = poisson` has to be specified.

::: callout
As the synopsis above is a code listing that is not meant to be executed, one can use either the dedicated `{Code}` environment or a simple `{verbatim}` environment for this. Again, spaces before and after should be avoided.

Finally, there might be a reference to a `{table}` such as @tbl-overview. Usually, these are placed at the top of the page (`[t!]`), centered (`\centering`), with a caption below the table, column headers and captions in sentence style, and if possible avoiding vertical lines.
:::

| Type           | Distribution | Method   | Description                                                                                                                                                                                  |
|---------|---------|---------|-----------------------------------------------|
| GLM            | Poisson      | ML       | Poisson regression: classical GLM, estimated by maximum likelihood (ML)                                                                                                                      |
|                |              | Quasi    | "Quasi-Poisson regression'': same mean function, estimated by quasi-ML (QML) or equivalently generalized estimating equations (GEE), inference adjustment via estimated dispersion parameter |
|                |              | Adjusted | "Adjusted Poisson regression'': same mean function, estimated by QML/GEE, inference adjustment via sandwich covariances                                                                      |
|                | NB           | ML       | NB regression: extended GLM, estimated by ML including additional shape parameter                                                                                                            |
| Zero-augmented | Poisson      | ML       | Zero-inflated Poisson (ZIP), hurdle Poisson                                                                                                                                                  |
|                | NB           | ML       | Zero-inflated NB (ZINB), hurdle NB                                                                                                                                                           |
: Overview of various count regression models. The table is usually placed at the top of the page (`[t!]`), centered (`centering`), has a caption below the table, column headers and captions are in sentence style, and if possible vertical lines should be avoided. {#tbl-overview}

## Illustrations {#sec-illustrations}

For a simple illustration of basic Poisson and NB count regression the
`quine` data from the [MASS]{.pkg} package is used. This provides the number
of `Days` that children were absent from school in Australia in a
particular year, along with several covariates that can be employed as regressors. The data can be loaded by

```{R}
#| prompt: true
data("quine", package = "MASS")
```
and a basic frequency distribution of the response variable is displayed in
@fig-quine.

:::{.callout}
For code input and output, the style files provide dedicated environments.
Either the "agnostic" `{CodeInput}` and `{CodeOutput}` can be used
or, equivalently, the environments `{Sinput}` and `{Soutput}` as
produced by [Sweave]{.fct} or [knitr]{.pkg} when using the `render_sweave()`
hook. Please make sure that all code is properly spaced, e.g., using
`y = a + b * x` and _not_ `y=a+b*x`. Moreover, code input should use "the usual" command prompt in the respective software system. For
[R]{.proglang} code, the prompt `R> ` should be used with `+  ` as
the continuation prompt. Generally, comments within the code chunks should be
avoided -- and made in the regular {{< latex >}} text instead. Finally, empty lines before and after code input/output should be avoided (see above).
:::

::: {#fig-quine}

![](article-visualization.pdf)

Frequency distribution for number of days absent from school.
:::

As a first model for the `quine` data, we fit the basic Poisson regression
model. (Note that JSS prefers when the second line of code is indented by two
spaces.)

```{R}
#| prompt: true
m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine, family = poisson)
```

To account for potential overdispersion we also consider a negative binomial
GLM.

```{R}
#| prompt: true
library("MASS")
m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
```

In a comparison with the BIC the latter model is clearly preferred.

```{R}
#| prompt: true
library("MASS")
BIC(m_pois, m_nbin)
```

Hence, the full summary of that model is shown below.

```{R}
#| prompt: true
summary(m_nbin)
```

## Summary and discussion {#sec-summary}

:::{.callout}

As usual…

:::

## Computational details {.unnumbered}

:::{.callout}

If necessary or useful, information about certain computational details
such as version numbers, operating systems, or compilers could be included
in an unnumbered section. Also, auxiliary packages (say, for visualizations,
maps, tables, …) that are not cited in the main text can be credited here.

:::

The results in this paper were obtained using [R]{.proglang}~3.4.1 with the
[MASS]{.pkg}~7.3.47 package. [R]{.proglang} itself and all packages used are available from the Comprehensive [R]{.proglang} Archive Network (CRAN) at
[https://CRAN.R-project.org/].

## Acknowledgments {.unnumbered}

:::{.callout}

All acknowledgments (note the AE spelling) should be collected in this
unnumbered section before the references. It may contain the usual information
about funding and feedback from colleagues/reviewers/etc. Furthermore,
information such as relative contributions of the authors may be added here
(if any).

:::

## References {.unnumbered}

:::{#refs}

:::

{{< pagebreak >}}

## More technical details {#sec-techdetails .unnumbered}

:::{.callout}

Appendices can be included after the bibliography (with a page break). Each
section within the appendix should have a proper section title (rather than
just _Appendix_).

For more technical style details, please check out JSS's style FAQ at
[https://www.jstatsoft.org/pages/view/style#frequently-asked-questions]
which includes the following topics:

- Title vs. sentence case.
- Graphics formatting.
- Naming conventions.
- Turning JSS manuscripts into [R]{.proglang} package vignettes.
- Trouble shooting.
- Many other potentially helpful details…

:::

## Using BibTeX {#sec-bibtex .unnumbered}

:::{.callout}

References need to be provided in a {{< bibtex >}} file (`.bib`). All
references should be made with `@cite` syntax. This commands yield different
formats of author-year citations and allow to include additional details (e.g.,pages, chapters, \dots) in brackets. In case you are not familiar with these
commands see the JSS style FAQ for details.

Cleaning up {{< bibtex >}} files is a somewhat tedious task -- especially
when acquiring the entries automatically from mixed online sources. However,
it is important that informations are complete and presented in a consistent
style to avoid confusions. JSS requires the following format.

- item JSS-specific markup (`\proglang`, `\pkg`, `\code`) should be used in the references.
- item Titles should be in title case.
- item Journal titles should not be abbreviated and in title case.
- item DOIs should be included where available.
- item Software should be properly cited as well. For [R]{.proglang} packages `citation("pkgname")` typically provides a good starting point.

::: -->

Spatial Interaction Models (SIMs) are mathematical models for estimating movement between spatial entities.
First developed in the late 1960s and early 1970, SIMs have become a key tool for transport modelling with substantial practical applications [@boyce_forecasting_2015].

An early and highly influential type of SIM was the 'gravity model', defined by @wilson_family_1971 as follows (in a paper that explored many extensions of this formulation and is influential enough to merit being quoted at length):

$$
T_{i j}=K \frac{W_{i}^{(1)} W_{j}^{(2)}}{c_{i j}^{n}}
$$

"where $T_{i j}$ is a measure of the interaction between zones $i$ and $W_{i}^{(1)}$ is a measure of the 'mass term' associated with zone $z_i$, $W_{j}^{(2)}$ is a measure of the 'mass term' associated with zone $z_j$, and $c_{ij}$ is a measure of the distance, or generalised cost of travel, between zone $i$ and zone $j$".
$K$ is a 'constant of proportionality' and $n$ is a parameter to be estimated.

We follow @simini_universal_2012 in replacing the $W$ terms above with $m$ and $n$ for origins and destinations respectively.
This seminal definition of the unconstrained 'gravity model' can be written in its "multiplicative form" as follows @dennett_modelling_2018, notation that we will use in this paper:

$$
T_{ij} = k m_i^{\mu} n_j^{*} d_{ij} \beta
$$

It is unconstrained because there is no limit on the amount of interaction (we will refer to the more concrete but semantically equivalent term "number of trips") in this paper.
There are four main types of traditional SIMs  [@wilson_family_1971]:

- Unconstrained
- Production-constrained
- Attraction-constrained
- Doubly-constrained

For the purposes of this project, we will focus on production-constrained SIMs.
These can be defined as follows [@wilson_family_1971]:

$$
T_{ij} = A_iO_in_jf(c_{ij})
$$

where $A$ is a balancing factor defined as:

$$
A_{i}=\frac{1}{\sum_{j} m_{j} \mathrm{f}\left(c_{i j}\right)}
$$

$O_i$ is analogous to the travel demand in zone $i$, which can be roughly approximated by its population.

More recent innovations in SIMs including the 'radiation model' @simini_universal_2012.
See @lenormand_systematic_2016 for a comparison of alternative approaches.

## Implementation in R

Before describing the functions in the package, it's worth implementing SIMs from first principles, to gain an understanding of how they work.
The code presented below was written before the functions in the [simodels]{.pkg} package were developed, building on @dennett_modelling_2018.
The aim is to demonstrate a common way of running SIMs, in a for loop, rather than using vectorised operations used in `simodels`.

```{r, message=FALSE}
library(tmap)
library(dplyr)
library(ggplot2)
```

```{r inputs}
zones = simodels::si_zones
centroids = simodels::si_centroids
od = simodels::si_od_census
tm_shape(zones) + tm_polygons("all", palette = "viridis")
```

```{r}
od_df = od::points_to_od(centroids)
od_sfc = od::odc_to_sfc(od_df[3:6])
sf::st_crs(od_sfc) = 4326
od_df$length = sf::st_length(od_sfc)
od_df = od_df %>% transmute(
  O, D, length = as.numeric(length) / 1000,
  flow = NA, fc = NA
  )
od_df = sf::st_sf(od_df, geometry = od_sfc, crs = 4326)
```

An unconstrained spatial interaction model can be written as follows, with a more-or-less arbitrary value for `beta` which can be optimised later:

```{r unconstrained}
beta = 0.3
i = 1
j = 2
for(i in seq(nrow(zones))) {
  for(j in seq(nrow(zones))) {
    O = zones$all[i]
    n = zones$all[j]
    ij = which(od_df$O == zones$geo_code[i] & od_df$D == zones$geo_code[j])
    od_df$fc[ij] = exp(-beta * od_df$length[ij])
    od_df$flow[ij] = O * n * od_df$fc[ij]
  }
}
od_top = od_df %>% 
  filter(O != D) %>% 
  top_n(n = 2000, wt = flow)

tm_shape(zones) +
  tm_borders() +
  tm_shape(od_top) +
  tm_lines("flow")
```

We can plot the 'distance decay' curve associated with this SIM is as follows:

```{r distance_decay}
summary(od_df$fc)
od_df %>% 
  ggplot() +
  geom_point(aes(length, fc))
```

We can make this production constrained as follows:

```{r constrained}
od_dfj = left_join(
  od_df,
  zones %>% select(O = geo_code, all) %>% sf::st_drop_geometry()
)
od_dfj = od_dfj %>% 
  group_by(O) %>% 
  mutate(flow_constrained = flow / sum(flow) * first(all)) %>%
  ungroup()
sum(od_dfj$flow_constrained) == sum(zones$all)
od_top = od_dfj %>% 
  filter(O != D) %>% 
  top_n(n = 2000, wt = flow_constrained)

tm_shape(zones) +
  tm_borders() +
  tm_shape(od_top) +
  tm_lines("flow_constrained")
```

# OD data preparation

# Interaction calculation

# Interaction modelling

# Examples

```{r validation}
od_dfjc = inner_join(od_dfj %>% select(-all), od)
od_dfjc %>% 
  ggplot() +
  geom_point(aes(all, flow_constrained))
cor(od_dfjc$all, od_dfjc$flow_constrained)^2
```

# Conclusions

### 2024-07-26 paper/paper.qmd
[](.pkg) syntax

@@
 abstract: |
-  This article introduces the `simodels` package for spatial interaction modelling in R.
+  This article introduces the [simodels]{.pkg} package for spatial interaction modelling in R.
 The package provides functions for estimating spatial interaction models, including unconstrained gravity models and models that are production-, attraction-, or doubly-constrained.
@@
 The code presented below was written before the functions in the [simodels]{.pkg} package were developed, building on @dennett_modelling_2018.
-The aim is to demonstrate a common way of running SIMs, in a for loop, rather than using vectorised operations used in `simodels`.
+The aim is to demonstrate a common way of running SIMs, in a for loop, rather than using vectorised operations used in [simodels]{.pkg}.
 

### 2024-07-31 paper/paper.qmd
document

@@
 bibliography: bibliography.bib
+execute: 
+  messages: false
+  warning: false 
 ---
@@
 
+## Introduction
+
 Spatial Interaction Models (SIMs) are mathematical models for estimating movement between spatial entities.
@@
 
-## Implementation in R
+The paper is structured as follows.
+@sec-base demonstrates how to implement an unconstrained SIM in base [R]{.proglang}, highlighting the need for a package to streamline the process.
+@sec-od describes the process of converting spatial data to Origin-Destination (OD) data, a key step in SIMs, and the functions in the package that facilitate this process.
+@sec-interaction describes the process of calculating interaction with pre-defined parameters with the `si_calculate()` function.
+
+## Basic implementation of SIMs {#sec-base}
 
@@
 The aim is to demonstrate a common way of running SIMs, in a for loop, rather than using vectorised operations used in [simodels]{.pkg}.
+We will lust the following packages for data manipulation and manipulation, using only base [R]{.proglang} functions for the SIM calculations.
 
-```{r, message=FALSE}
+```{r}
+#| label: load-packages
 library(tmap)
@@
 ```
+
+The package contains a number of datasets that can be used to demonstrate the SIMs, as shown below:
+
+- `si_zones`: a dataset of zones with population data
+- `si_centroids`: a dataset of zone centroids
+- `si_od_census`: a dataset of Origin-Destination (OD) data for evaluating the SIMs
 
@@
 
+We'll use the [od](.pkg) package to convert the geographic data to OD data.
+We set the `length` column to the length of the line between the centroids of the zones, in km, and set the coordinate reference system (CRS) to WGS84 (EPSG:4326).
+
 ```{r}
@@
 od_df$length = sf::st_length(od_sfc)
-od_df = od_df %>% transmute(
+od_df = od_df |>
+  transmute(
     O, D, length = as.numeric(length) / 1000,
@@
 beta = 0.3
-i = 1
-j = 2
 for(i in seq(nrow(zones))) {
@@
 }
-od_top = od_df %>% 
-  filter(O != D) %>% 
-  top_n(n = 2000, wt = flow)
+```
+
+The results of the unconstrained SIM are shown in @fig-unconstrained, with the geographic desire lines showing the interaction between zones and a scatter plot illustrating the 'friction of distance'.
+
+```{r}
+#| label: fig-unconstrained
+#| fig-cap: Unconstrained spatial interaction model
+#| fig-subcap: 
+#|   - "Geographic desire lines showing interaction between zones calculated by the unconstrained SIM"
+#|   - "Scatter plot illustrating the 'friction of distance' in the unconstrained SIM"
+#| layout-ncol: 2
+#| echo: false
+od_top = od_df |> 
+  filter(O != D) |> 
+  arrange(desc(flow)) |>
+  slice(seq(2000)) |>
+  arrange(flow)
 
@@
   tm_lines("flow")
-```
-
-We can plot the 'distance decay' curve associated with this SIM is as follows:
-
-```{r distance_decay}
-summary(od_df$fc)
-od_df %>% 
+od_df |> 
   ggplot() +
@@
   od_df,
-  zones %>% select(O = geo_code, all) %>% sf::st_drop_geometry()
+  zones |> select(O = geo_code, all) |> sf::st_drop_geometry()
 )
-od_dfj = od_dfj %>% 
-  group_by(O) %>% 
-  mutate(flow_constrained = flow / sum(flow) * first(all)) %>%
+od_dfj = od_dfj |> 
+  group_by(O) |> 
+  mutate(flow_constrained = flow / sum(flow) * first(all)) |>
   ungroup()
 sum(od_dfj$flow_constrained) == sum(zones$all)
-od_top = od_dfj %>% 
-  filter(O != D) %>% 
-  top_n(n = 2000, wt = flow_constrained)
+od_top = od_dfj |> 
+  filter(O != D) |> 
+  arrange(desc(flow_constrained)) |>
+  slice(seq(2000)) |>
+  arrange(flow_constrained)
 
@@
 
-# OD data preparation
+## OD data preparation {#sec-od}
 
-# Interaction calculation
+## Interaction calculation {#sec-interaction}
 
-# Interaction modelling
+## Interaction modelling {#sec-models}
 
-# Examples
+## Examples {#sec-examples}
 
 ```{r validation}
-od_dfjc = inner_join(od_dfj %>% select(-all), od)
-od_dfjc %>% 
+od_dfjc = inner_join(od_dfj |> select(-all), od)
+od_dfjc |> 
   ggplot() +
@@
 
-# Conclusions
+## Conclusions

### 2024-08-01 R/data.R
Add new datasets (#34)

@@ -60,3 +60,28 @@ NULL
 #' @examples
 #' head(od_aus)
 NULL
+
+#' Example zones dataset: administrative zones of York
+#' 
+#' See data-raw/zones_york.qmd for details on the data source.
+#' 
+#' @docType data
+#' @keywords datasets
+#' @name zones_york
+#' @examples
+#' head(zones_york)
+#' plot(zones_york$geometry)
+NULL
+
+#' Example destinations dataset: schools in York
+#' 
+#' Example dataset from York, UK
+#' 
+#' See data-raw/zones_york.qmd for details on the data source.
+#' 
+#' @docType data
+#' @keywords datasets
+#' @name destinations_york
+#' @examples
+#' head(destinations_york)
+NULL

### 2024-08-02 paper/paper.qmd
Diagnose reason of weird result (inner not left join), close #34

@@
   warning: false 
+  echo: false
 ---
 
-<!-- ## Introduction: Count data regression in R {#sec-intro}
+## Introduction
+
+Spatial Interaction Models (SIMs) are mathematical models for estimating movement between spatial entities.
+First developed in the late 1960s and early 1970, SIMs have become a key tool for transport modelling with substantial practical applications [@boyce_forecasting_2015].
+
+An early and highly influential type of SIM was the 'gravity model', defined by @wilson_family_1971 as follows (in a paper that explored many extensions of this formulation and is influential enough to merit being quoted at length):
+
+$$
+T_{i j}=K \frac{W_{i}^{(1)} W_{j}^{(2)}}{c_{i j}^{n}}
+$$
+
+"where $T_{i j}$ is a measure of the interaction between zones $i$ and $W_{i}^{(1)}$ is a measure of the 'mass term' associated with zone $z_i$, $W_{j}^{(2)}$ is a measure of the 'mass term' associated with zone $z_j$, and $c_{ij}$ is a measure of the distance, or generalised cost of travel, between zone $i$ and zone $j$".
+$K$ is a 'constant of proportionality' and $n$ is a parameter to be estimated.
+
+We follow @simini_universal_2012 in replacing the $W$ terms above with $m$ and $n$ for origins and destinations respectively.
+This seminal definition of the unconstrained 'gravity model' can be written in its "multiplicative form" as follows @dennett_modelling_2018, notation that we will use in this paper:
+
+$$
+T_{ij} = k m_i^{\mu} n_j^{*} d_{ij} \beta
+$$
+
+It is unconstrained because there is no limit on the amount of interaction (we will refer to the more concrete but semantically equivalent term "number of trips") in this paper.
+There are four main types of traditional SIMs  [@wilson_family_1971]:
+
+- Unconstrained
+- Production-constrained
+- Attraction-constrained
+- Doubly-constrained
+
+For the purposes of this project, we will focus on production-constrained SIMs.
+These can be defined as follows [@wilson_family_1971]:
+
+$$
+T_{ij} = A_iO_in_jf(c_{ij})
+$$
+
+where $A$ is a balancing factor defined as:
+
+$$
+A_{i}=\frac{1}{\sum_{j} m_{j} \mathrm{f}\left(c_{i j}\right)}
+$$
+
+$O_i$ is analogous to the travel demand in zone $i$, which can be roughly approximated by its population.
+
+More recent innovations in SIMs including the 'radiation model' @simini_universal_2012.
+See @lenormand_systematic_2016 for a comparison of alternative approaches.
+The purpose of the package is not to implement all available models, but to provide a flexible framework to speed-up the process of implementing SIMs in [R]{.proglang}.
+The package supports a wide range of models and can be used to fit unconstrained, production-constrained, attraction-constrained, and doubly-constrained models.
+Furthermore, the package works equally well for 'bipartite' SIMs, in which origin and destination points are different (Hasova et al. [2022](https://lenkahas.com/files/preprint.pdf)).
+
+The paper is structured as follows.
+@sec-base demonstrates how to implement an unconstrained SIM in base [R]{.proglang}, highlighting the need for a package to streamline the process.
+@sec-od describes the process of converting spatial data to Origin-Destination (OD) data, a key step in SIMs, and the functions in the package that facilitate this process.
+@sec-interaction describes the process of calculating interaction with pre-defined parameters with the `si_calculate()` function.
+
+## Basic implementation of SIMs {#sec-base}
+
+Before describing the functions in the package, it's worth implementing SIMs from first principles, to gain an understanding of how they work and to highlight the benefits of using the package.
+The code presented below was written before the functions in the [simodels]{.pkg} package were developed, building on @dennett_modelling_2018.
+The aim is to demonstrate a common way of running SIMs, in a for loop, rather than using vectorised operations used in [simodels]{.pkg}.
+In this section we the [tmap]{.pkg} and [tidyverse]{.pkg} packages for data manipulation and visualisation, using only base [R]{.proglang} functions for the SIM calculations (see the paper's source code for details).
+
+```{r}
+#| label: load-packages
+library(tmap)
+library(tidyverse)
+library(patchwork)
+devtools::load_all()
+```
+
+The package contains a number of datasets that can be used to demonstrate the SIMs, as shown below:
+
+- `si_zones`: a dataset of zones with population data
+- `si_centroids`: a dataset of zone centroids
+- `si_od_census`: a dataset of Origin-Destination (OD) data for evaluating the SIMs
+- `zones_york`: a dataset of zones in York
+- `destinations_york`: a dataset of school locations in York with information on number of pupils attending, enabling destination-constrained and doubly-constrained SIMs to be implemented
+
+```{r inputs}
+zones = simodels::si_zones
+centroids = simodels::si_centroids
+od = simodels::si_od_census
+tm_shape(zones) + tm_polygons("all", palette = "viridis")
+```
+
+We'll use the [od](.pkg) package to convert the geographic data to OD data.
+We set the `length` column to the length of the line between the centroids of the zones, in km, and set the coordinate reference system (CRS) to WGS84 (EPSG:4326), as follows:
+
+```{r}
+#| echo: true
+od_df = od::points_to_od(centroids)
+od_sfc = od::odc_to_sfc(od_df[3:6])
+sf::st_crs(od_sfc) = 4326
+od_df$length = sf::st_length(od_sfc)
+od_df = od_df |>
+  transmute(
+    O, D, length = as.numeric(length) / 1000,
+    flow = NA, fc = NA
+  )
+od_df = sf::st_sf(od_df, geometry = od_sfc, crs = 4326)
+```
+
+An unconstrained spatial interaction model can be written as follows, with a more-or-less arbitrary value for `beta` which can be optimised later:
+
+```{r}
+#| label: unconstrained
+#| echo: true
+beta = 0.3
+for(i in seq(nrow(zones))) {
+  for(j in seq(nrow(zones))) {
+    O = zones$all[i]
+    n = zones$all[j]
+    ij = which(od_df$O == zones$geo_code[i] & od_df$D == zones$geo_code[j])
+    od_df$fc[ij] = exp(-beta * od_df$length[ij])
+    od_df$flow[ij] = O * n * od_df$fc[ij]
+  }
+}
+```
+
+The results of the unconstrained SIM are shown in @fig-unconstrained, with the geographic desire lines showing the interaction between zones and a scatter plot illustrating the 'friction of distance'.
+
+```{r}
+#| label: fig-unconstrained
+#| fig-cap: Unconstrained spatial interaction model
+#| fig-subcap: 
+#|   - "Geographic desire lines showing interaction between zones calculated by the unconstrained SIM"
+#|   - "Scatter plot illustrating the 'friction of distance' in the unconstrained SIM"
+#| layout-ncol: 2
+#| echo: false
+od_top = od_df |> 
+  filter(O != D) |> 
+  arrange(desc(flow)) |>
+  slice(seq(2000)) |>
+  arrange(flow)
+
+tm_shape(zones) +
+  tm_borders() +
+  tm_shape(od_top) +
+  tm_lines("flow")
+od_df |> 
+  ggplot() +
+  geom_point(aes(length, fc))
+```
+
+We can make this production constrained by dividing the estimated interaction by the difference between the sum of the estimated interactions and the sum of the observed interactions per origin.
+
+```{r}
+#| label: constrained
+#| include: false
+od_dfj = left_join(
+  od_df,
+  zones |> select(O = geo_code, all) |> sf::st_drop_geometry()
+)
+od_dfj = od_dfj |> 
+  group_by(O) |> 
+  mutate(flow_constrained = flow / sum(flow) * first(all)) |>
+  ungroup()
+sum(od_dfj$flow_constrained) == sum(zones$all)
+od_top = od_dfj |> 
+  filter(O != D) |> 
+  arrange(desc(flow_constrained)) |>
+  slice(seq(2000)) |>
+  arrange(flow_constrained)
+
+tm_shape(zones) +
+  tm_borders() +
+  tm_shape(od_top) +
+  tm_lines("flow_constrained")
+```
+
+## OD data preparation {#sec-od}
+
+## Interaction calculation {#sec-interaction}
+
+```{r}
+#| label: inputs-york
+g1 = zones_york |>
+  ggplot() +
+  geom_sf() +
+  labs(title = "Zones")
+g2 = destinations_york |>
+    ggplot() +
+    geom_sf() +
+    labs(title = "Destinations")
+g1 + g2
+```
+
+Before we run any models let's compare the total number of pupils in the zones dataset and the destinations dataset (they should be the same):
+
+```{r}
+#| label: compare-totals
+#| echo: true
+zone_overestimate_factor = 
+  (sum(zones_york$f0_to_15) + sum(zones_york$m0_to_15)) /
+    sum(destinations_york$n_pupils)
+zone_overestimate_factor
+```
+
+As one would expect, the total number of pupils in the zones dataset is a bit bigger than the total number of pupils in the destinations dataset: not all people aged 0 to 15 go to school, especially those under school age.
+To tackle this issue we'll create a new variables called `pupils_estimated` in the zones dataset, which is the sum of the number of pupils in the zones dataset and the number of pupils in the destinations dataset.
+
+```{r}
+#| label: add-pupils-estimated
+#| echo: true
+zones_york = zones_york |>
+  dplyr::mutate(
+    pupils_estimated = (f0_to_15 + m0_to_15) / zone_overestimate_factor
+  )
+```
+
+After the adjustment shown above, the totals in the origin and destination columns should be the same:
+
+```{r}
+#| label: compare-totals-after
+#| echo: true
+sum(zones_york$pupils_estimated)
+sum(destinations_york$n_pupils)
+```
+
+Based on these inputs the `si_to_od()` function generates the OD data, as shown below (note: 2 versions are created, one with a maximum distance constraint for speed of processing, important when working with large datasets).
+
+```{r}
+#| echo: true
+max_dist = 5000 # meters
+od_from_si_full = simodels::si_to_od(zones_york, destinations_york)
+od_from_si = simodels::si_to_od(zones_york, destinations_york, max_dist = max_dist)
+```
+
+```{r}
+#| label: plot-od-all
+#| layout-ncol: 2
+m1 = od_from_si_full |>
+  ggplot() +
+  geom_sf(alpha = 0.1)
+m2 = od_from_si |>
+  ggplot() +
+  geom_sf(alpha = 0.1)
+m1
+m2
+```
+
+The output OD dataset has column names taken from both the origin and destination datasets, with the following column names:
+
+```{r}
+names(od_from_si)
+```
+
+Let's run a simple model:
+
+```{r}
+#| label: simple-model-york
+#| echo: true
+gravity_model = function(beta, d, m, n) {
+  m * n * exp(-beta * d / 1000)
+} 
+# perform SIM
+od_res = simodels::si_calculate(
+  od_from_si,
+  fun = gravity_model,
+  d = distance_euclidean,
+  m = origin_pupils_estimated,
+  n = destination_n_pupils,
+#   constraint_production = origin_all,
+  beta = 0.9
+  )
+```
+
+We'll make one adjustment to the output dataset, renaming the `interaction` column to `trips`, and setting the total number of trips to be the same as the total number of pupils in the destinations dataset:
+
+```{r}
+#| label: adjust-output-york
+#| echo: true
+interaction_overestimate_factor = sum(destinations_york$n_pupils) / sum(od_res$interaction)
+od_res = od_res |>
+  dplyr::mutate(
+    trips = interaction * interaction_overestimate_factor
+  )
+```
+
+We can assess the model fit at three levels: the origin level (number of students departing from each zone), the destination level (the number arriving at each school in the input dataset) and the origin-destination level.
+
+```{r}
+#| label: r-squared-york
+res_o = od_res |>
+  group_by(O) |>
+  summarise(
+    Observed = first(origin_pupils_estimated),
+    Modelled = sum(trips),
+    Type = "Origin"
+  )
+res_d = od_res |>
+  group_by(D) |>
+  summarise(
+    Observed = first(destination_n_pupils),
+    Modelled = sum(trips),
+    Type = "Destination"
+  )
+res_combined = bind_rows(res_o, res_d) |>
+  # Create ordered factor with types:
+  mutate(
+    Type = factor(Type, levels = c("Origin", "Destination"))
+  )
+g_combined = res_combined |>
+  ggplot() +
+    geom_point(aes(x = Observed, y = Modelled)) +
+    geom_smooth(aes(x = Observed, y = Modelled), method = "lm") +
+    facet_wrap(~Type, scales = "free") +
+    labs(
+      title = "Model fit at origin, destination and OD levels (unconstrained)",
+      x = "Observed",
+      y = "Modelled"
+    ) 
+g_combined
+```
+
+Let's see if making the model production constrained can help:
+
+```{r}
+#| label: constrained-york
+#| echo: true
+od_res_constrained = simodels::si_calculate(
+  od_from_si,
+  fun = gravity_model,
+  d = distance_euclidean,
+  m = origin_pupils_estimated,
+  n = destination_n_pupils,
+  constraint_production = origin_pupils_estimated,
+  beta = 0.9
+  )
+```
+
+```{r}
+#| label: r-squared-constrained
+res_o = od_res_constrained |>
+  group_by(O) |>
+  summarise(
+    Observed = first(origin_pupils_estimated),
+    Modelled = sum(interaction),
+    Type = "Origin"
+  )
+res_d = od_res_constrained |>
+  group_by(D) |>
+  summarise(
+    Observed = first(destination_n_pupils),
+    Modelled = sum(interaction),
+    Type = "Destination"
+  )
+res_combined = bind_rows(res_o, res_d) |>
+  mutate(
+    Type = factor(Type, levels = c("Origin", "Destination"))
+  )
+res_combined |>
+  ggplot() +
+    geom_point(aes(x = Observed, y = Modelled)) +
+    geom_smooth(aes(x = Observed, y = Modelled), method = "lm") +
+    facet_wrap(~Type, scales = "free") +
+    labs(
+      title = "Model fit at origin, destination and OD levels (production constrained)",
+      x = "Observed",
+      y = "Modelled"
+    )
+```
+
+## Interaction modelling {#sec-models}
+
+## Examples {#sec-examples}
+
+```{r validation}
+od_dfjc = inner_join(od_dfj |> select(-all), od)
+od_dfjc |> 
+  ggplot() +
+  geom_point(aes(all, flow_constrained))
+cor(od_dfjc$all, od_dfjc$flow_constrained)^2
+```
+
+## Conclusions
+
+<!--
+
+From paper-writing docs:
+ ## Introduction: Count data regression in R {#sec-intro}
 
@@
 ::: -->
-
-## Introduction
-
-Spatial Interaction Models (SIMs) are mathematical models for estimating movement between spatial entities.
-First developed in the late 1960s and early 1970, SIMs have become a key tool for transport modelling with substantial practical applications [@boyce_forecasting_2015].
-
-An early and highly influential type of SIM was the 'gravity model', defined by @wilson_family_1971 as follows (in a paper that explored many extensions of this formulation and is influential enough to merit being quoted at length):
-
-$$
-T_{i j}=K \frac{W_{i}^{(1)} W_{j}^{(2)}}{c_{i j}^{n}}
-$$
-
-"where $T_{i j}$ is a measure of the interaction between zones $i$ and $W_{i}^{(1)}$ is a measure of the 'mass term' associated with zone $z_i$, $W_{j}^{(2)}$ is a measure of the 'mass term' associated with zone $z_j$, and $c_{ij}$ is a measure of the distance, or generalised cost of travel, between zone $i$ and zone $j$".
-$K$ is a 'constant of proportionality' and $n$ is a parameter to be estimated.
-
-We follow @simini_universal_2012 in replacing the $W$ terms above with $m$ and $n$ for origins and destinations respectively.
-This seminal definition of the unconstrained 'gravity model' can be written in its "multiplicative form" as follows @dennett_modelling_2018, notation that we will use in this paper:
-
-$$
-T_{ij} = k m_i^{\mu} n_j^{*} d_{ij} \beta
-$$
-
-It is unconstrained because there is no limit on the amount of interaction (we will refer to the more concrete but semantically equivalent term "number of trips") in this paper.
-There are four main types of traditional SIMs  [@wilson_family_1971]:
-
-- Unconstrained
-- Production-constrained
-- Attraction-constrained
-- Doubly-constrained
-
-For the purposes of this project, we will focus on production-constrained SIMs.
-These can be defined as follows [@wilson_family_1971]:
-
-$$
-T_{ij} = A_iO_in_jf(c_{ij})
-$$
-
-where $A$ is a balancing factor defined as:
-
-$$
-A_{i}=\frac{1}{\sum_{j} m_{j} \mathrm{f}\left(c_{i j}\right)}
-$$
-
-$O_i$ is analogous to the travel demand in zone $i$, which can be roughly approximated by its population.
-
-More recent innovations in SIMs including the 'radiation model' @simini_universal_2012.
-See @lenormand_systematic_2016 for a comparison of alternative approaches.
-
-The paper is structured as follows.
-@sec-base demonstrates how to implement an unconstrained SIM in base [R]{.proglang}, highlighting the need for a package to streamline the process.
-@sec-od describes the process of converting spatial data to Origin-Destination (OD) data, a key step in SIMs, and the functions in the package that facilitate this process.
-@sec-interaction describes the process of calculating interaction with pre-defined parameters with the `si_calculate()` function.
-
-## Basic implementation of SIMs {#sec-base}
-
-Before describing the functions in the package, it's worth implementing SIMs from first principles, to gain an understanding of how they work.
-The code presented below was written before the functions in the [simodels]{.pkg} package were developed, building on @dennett_modelling_2018.
-The aim is to demonstrate a common way of running SIMs, in a for loop, rather than using vectorised operations used in [simodels]{.pkg}.
-We will lust the following packages for data manipulation and manipulation, using only base [R]{.proglang} functions for the SIM calculations.
-
-```{r}
-#| label: load-packages
-library(tmap)
-library(dplyr)
-library(ggplot2)
-```
-
-The package contains a number of datasets that can be used to demonstrate the SIMs, as shown below:
-
-- `si_zones`: a dataset of zones with population data
-- `si_centroids`: a dataset of zone centroids
-- `si_od_census`: a dataset of Origin-Destination (OD) data for evaluating the SIMs
-
-```{r inputs}
-zones = simodels::si_zones
-centroids = simodels::si_centroids
-od = simodels::si_od_census
-tm_shape(zones) + tm_polygons("all", palette = "viridis")
-```
-
-We'll use the [od](.pkg) package to convert the geographic data to OD data.
-We set the `length` column to the length of the line between the centroids of the zones, in km, and set the coordinate reference system (CRS) to WGS84 (EPSG:4326).
-
-```{r}
-od_df = od::points_to_od(centroids)
-od_sfc = od::odc_to_sfc(od_df[3:6])
-sf::st_crs(od_sfc) = 4326
-od_df$length = sf::st_length(od_sfc)
-od_df = od_df |>
-  transmute(
-    O, D, length = as.numeric(length) / 1000,
-    flow = NA, fc = NA
-  )
-od_df = sf::st_sf(od_df, geometry = od_sfc, crs = 4326)
-```
-
-An unconstrained spatial interaction model can be written as follows, with a more-or-less arbitrary value for `beta` which can be optimised later:
-
-```{r unconstrained}
-beta = 0.3
-for(i in seq(nrow(zones))) {
-  for(j in seq(nrow(zones))) {
-    O = zones$all[i]
-    n = zones$all[j]
-    ij = which(od_df$O == zones$geo_code[i] & od_df$D == zones$geo_code[j])
-    od_df$fc[ij] = exp(-beta * od_df$length[ij])
-    od_df$flow[ij] = O * n * od_df$fc[ij]
-  }
-}
-```
-
-The results of the unconstrained SIM are shown in @fig-unconstrained, with the geographic desire lines showing the interaction between zones and a scatter plot illustrating the 'friction of distance'.
-
-```{r}
-#| label: fig-unconstrained
-#| fig-cap: Unconstrained spatial interaction model
-#| fig-subcap: 
-#|   - "Geographic desire lines showing interaction between zones calculated by the unconstrained SIM"
-#|   - "Scatter plot illustrating the 'friction of distance' in the unconstrained SIM"
-#| layout-ncol: 2
-#| echo: false
-od_top = od_df |> 
-  filter(O != D) |> 
-  arrange(desc(flow)) |>
-  slice(seq(2000)) |>
-  arrange(flow)
-
-tm_shape(zones) +
-  tm_borders() +
-  tm_shape(od_top) +
-  tm_lines("flow")
-od_df |> 
-  ggplot() +
-  geom_point(aes(length, fc))
-```
-
-We can make this production constrained as follows:
-
-```{r constrained}
-od_dfj = left_join(
-  od_df,
-  zones |> select(O = geo_code, all) |> sf::st_drop_geometry()
-)
-od_dfj = od_dfj |> 
-  group_by(O) |> 
-  mutate(flow_constrained = flow / sum(flow) * first(all)) |>
-  ungroup()
-sum(od_dfj$flow_constrained) == sum(zones$all)
-od_top = od_dfj |> 
-  filter(O != D) |> 
-  arrange(desc(flow_constrained)) |>
-  slice(seq(2000)) |>
-  arrange(flow_constrained)
-
-tm_shape(zones) +
-  tm_borders() +
-  tm_shape(od_top) +
-  tm_lines("flow_constrained")
-```
-
-## OD data preparation {#sec-od}
-
-## Interaction calculation {#sec-interaction}
-
-## Interaction modelling {#sec-models}
-
-## Examples {#sec-examples}
-
-```{r validation}
-od_dfjc = inner_join(od_dfj |> select(-all), od)
-od_dfjc |> 
-  ggplot() +
-  geom_point(aes(all, flow_constrained))
-cor(od_dfjc$all, od_dfjc$flow_constrained)^2
-```
-
-## Conclusions

### 2024-08-02 R/si_model.R

@@ -114,16 +114,30 @@ constrain_production = function(od, output_col, constraint_production) {
     "{output_col}" := .data[[output_col]] /
       sum(.data[[output_col]]) * dplyr::first( {{constraint_production}} )
   )
+  # browser()
   # # Assert values are correct for test data:
   # od_grouped |>
-  #   select(origin_all, interaction)
-  # od_grouped |>
+  #   select({{constraint_production}}, interaction)
+  # origin_totals = od_grouped |>
+  #   sf::st_drop_geometry() |>
   #   sf::st_drop_geometry() |>
   #   # group_by(1) |>
   #   summarise(
   #     sum = sum(interaction),
-  #     first = first(origin_all)
+  #     first = first({{constraint_production}})
   #   )
+  # cor(origin_totals$sum, origin_totals$first)
+  # # Test for york data:
+  # zone_totals = left_join(
+  #   zones_york |>
+  #     sf::st_drop_geometry() |>
+  #     rename(O = LSOA21CD) |>
+  #     select(O, pupils_estimated),
+  #   origin_totals
+  # )
+
+  
+  
   od = dplyr::ungroup(od_grouped)
   od
 }

### 2024-08-04 data-raw/si_oa_wpz.R
Add new dataset, oa_wpz

@@ -1,0 +1,3 @@
+## code to prepare `si_oa_wpz` dataset goes here
+
+

### 2024-08-04 README.qmd
Document new data

@@
 usethis::use_github_action("pkgdown")
+usethis::use_data_raw("si_oa_wpz")
 ```

### 2024-08-04 R/data.R

@@ -9,8 +9,8 @@
 #' @aliases si_centroids
 #' @examples
 #' si_zones
-#' plot(si_zones$geometry)
-#' plot(si_centroids$geometry, add = TRUE)
+#' sf:::plot.sfg(si_zones$geometry)
+#' sf:::plot.sfg(si_centroids$geometry, add = TRUE)
 NULL
 
 #' Example OD dataset
@@ -70,7 +70,7 @@ NULL
 #' @name zones_york
 #' @examples
 #' head(zones_york)
-#' plot(zones_york$geometry)
+#' sf:::plot.sfg(zones_york$geometry)
 NULL
 
 #' Example destinations dataset: schools in York
@@ -84,4 +84,39 @@ NULL
 #' @name destinations_york
 #' @examples
 #' head(destinations_york)
+NULL
+
+#' Origin-Destination Data for Leeds
+#' 
+#' This dataset contains origin-destination data for Leeds, including the number of trips between output areas (OAs) and workplace zones (WPZs).
+#' 
+#' @docType data
+#' @keywords datasets
+#' @name si_oa_wpz
+#' @examples
+#' head(si_oa_wpz)
+NULL
+
+#' Origin Data for Leeds
+#' 
+#' This dataset contains the number of trips originating from each output area (OA) in Leeds.
+#' 
+#' @docType data
+#' @keywords datasets
+#' @name si_oa_wpz_o
+#' @examples
+#' head(si_oa_wpz_o)
+#' sf:::plot.sf(si_oa_wpz_o["n_o"])
+NULL
+
+#' Destination Data for Leeds
+#' 
+#' This dataset contains the number of trips destined for each workplace zone (WPZ) in Leeds.
+#' 
+#' @docType data
+#' @keywords datasets
+#' @name si_oa_wpz_d
+#' @examples
+#' head(si_oa_wpz_d)
+#' sf:::plot.sf(si_oa_wpz_d["n_d"])
 NULL

### 2024-08-04 data-raw/si_oa_wpz.R
Split-out schools_york example

@@ -1,3 +1,0 @@
-## code to prepare `si_oa_wpz` dataset goes here
-
-

### 2024-08-04 paper/paper.qmd
Add instructions to paper.qmd for reproducibility, thanks @JuanFonsecaLS1
Split-out schools_york example
html version

@@
 title: "simodels: A package for Spatial Interaction Modelling"
+---
+
+<!-- 
+# Note: you need to have the relevant quarto extension installed:
+quarto add quarto-journals/jss
+After that put the following in the yaml:
 format:
@@
   echo: false
----
+-->
+
+```{r}
+#| include: false
+knitr::opts_chunk$set(
+  message = FALSE,
+  warning = FALSE,
+  echo = FALSE
+)
+```
 

### 2024-08-19 data-raw/schools_york.qmd
Merge code from Malcolm

--- 2024-08-02 paper/paper.qmd
@@
----
-title: "simodels: A package for Spatial Interaction Modelling"
-format:
-    jss-pdf:
-        keep-tex: true
-    jss-html: default
-# author:
-#   - name: Achim Zeileis
-#     affiliations:
-#       - name: Universität Innsbruck
-#         department: Department of Statistics
-#         address: Universitätsstr. 15
-#         city: Innsbruck
-#         country: Austria
-#         postal-code: 6020
-#       - Journal of Statistical Software
-#     orcid: 0000-0003-0918-3766
-#     email: Achim.Zeileis@R-project.org
-#     url: https://www.zeileis.org/
-#   - name: Second  Author
-#     affiliations:
-#       - Plus Affiliation
-# abstract: |
-#   This short article illustrates how to write a manuscript for the
-#   _Journal of Statistical Software_ (JSS) using its {{< latex >}} style files.
-#   Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
-#   it is recommended to keep the {{< latex >}} code as simple as possible,
-#   i.e., avoid inclusion of packages/commands that are not necessary.
-#   For outlining the typical structure of a JSS article some brief text snippets are employed that have been inspired by @ZeileisKleiberJackman2008, discussing count data regression in [R]{.proglang}.
-Editorial comments and instructions are marked by vertical bars.
-abstract: |
-  This article introduces the [simodels]{.pkg} package for spatial interaction modelling in R.
-The package provides functions for estimating spatial interaction models, including unconstrained gravity models and models that are production-, attraction-, or doubly-constrained.
+# Modelling travel to school in York
 
-keywords: [JSS, style guide, comma-separated, not capitalized, R]
-keywords-formatted: [JSS, style guide, comma-separated, not capitalized, "[R]{.proglang}"]
+The code below demonstrates how to fit a SIM with origins and destinations.
+The data used in this example is from the city of York, UK, and is available in the `simodels` package.
 
-bibliography: bibliography.bib
-execute: 
-  messages: false
-  warning: false 
-  echo: false
----
-
-## Introduction
-
-Spatial Interaction Models (SIMs) are mathematical models for estimating movement between spatial entities.
-First developed in the late 1960s and early 1970, SIMs have become a key tool for transport modelling with substantial practical applications [@boyce_forecasting_2015].
-
-An early and highly influential type of SIM was the 'gravity model', defined by @wilson_family_1971 as follows (in a paper that explored many extensions of this formulation and is influential enough to merit being quoted at length):
-
-$$
-T_{i j}=K \frac{W_{i}^{(1)} W_{j}^{(2)}}{c_{i j}^{n}}
-$$
-
-"where $T_{i j}$ is a measure of the interaction between zones $i$ and $W_{i}^{(1)}$ is a measure of the 'mass term' associated with zone $z_i$, $W_{j}^{(2)}$ is a measure of the 'mass term' associated with zone $z_j$, and $c_{ij}$ is a measure of the distance, or generalised cost of travel, between zone $i$ and zone $j$".
-$K$ is a 'constant of proportionality' and $n$ is a parameter to be estimated.
-
-We follow @simini_universal_2012 in replacing the $W$ terms above with $m$ and $n$ for origins and destinations respectively.
-This seminal definition of the unconstrained 'gravity model' can be written in its "multiplicative form" as follows @dennett_modelling_2018, notation that we will use in this paper:
-
-$$
-T_{ij} = k m_i^{\mu} n_j^{*} d_{ij} \beta
-$$
-
-It is unconstrained because there is no limit on the amount of interaction (we will refer to the more concrete but semantically equivalent term "number of trips") in this paper.
-There are four main types of traditional SIMs  [@wilson_family_1971]:
-
-- Unconstrained
-- Production-constrained
-- Attraction-constrained
-- Doubly-constrained
-
-For the purposes of this project, we will focus on production-constrained SIMs.
-These can be defined as follows [@wilson_family_1971]:
-
-$$
-T_{ij} = A_iO_in_jf(c_{ij})
-$$
-
-where $A$ is a balancing factor defined as:
-
-$$
-A_{i}=\frac{1}{\sum_{j} m_{j} \mathrm{f}\left(c_{i j}\right)}
-$$
-
-$O_i$ is analogous to the travel demand in zone $i$, which can be roughly approximated by its population.
-
-More recent innovations in SIMs including the 'radiation model' @simini_universal_2012.
-See @lenormand_systematic_2016 for a comparison of alternative approaches.
-The purpose of the package is not to implement all available models, but to provide a flexible framework to speed-up the process of implementing SIMs in [R]{.proglang}.
-The package supports a wide range of models and can be used to fit unconstrained, production-constrained, attraction-constrained, and doubly-constrained models.
-Furthermore, the package works equally well for 'bipartite' SIMs, in which origin and destination points are different (Hasova et al. [2022](https://lenkahas.com/files/preprint.pdf)).
-
-The paper is structured as follows.
-@sec-base demonstrates how to implement an unconstrained SIM in base [R]{.proglang}, highlighting the need for a package to streamline the process.
-@sec-od describes the process of converting spatial data to Origin-Destination (OD) data, a key step in SIMs, and the functions in the package that facilitate this process.
-@sec-interaction describes the process of calculating interaction with pre-defined parameters with the `si_calculate()` function.
-
-## Basic implementation of SIMs {#sec-base}
-
-Before describing the functions in the package, it's worth implementing SIMs from first principles, to gain an understanding of how they work and to highlight the benefits of using the package.
-The code presented below was written before the functions in the [simodels]{.pkg} package were developed, building on @dennett_modelling_2018.
-The aim is to demonstrate a common way of running SIMs, in a for loop, rather than using vectorised operations used in [simodels]{.pkg}.
-In this section we the [tmap]{.pkg} and [tidyverse]{.pkg} packages for data manipulation and visualisation, using only base [R]{.proglang} functions for the SIM calculations (see the paper's source code for details).
-
-```{r}
-#| label: load-packages
-library(tmap)
-library(tidyverse)
-library(patchwork)
-devtools::load_all()
-```
-
-The package contains a number of datasets that can be used to demonstrate the SIMs, as shown below:
-
-- `si_zones`: a dataset of zones with population data
-- `si_centroids`: a dataset of zone centroids
-- `si_od_census`: a dataset of Origin-Destination (OD) data for evaluating the SIMs
-- `zones_york`: a dataset of zones in York
-- `destinations_york`: a dataset of school locations in York with information on number of pupils attending, enabling destination-constrained and doubly-constrained SIMs to be implemented
-
-```{r inputs}
-zones = simodels::si_zones
-centroids = simodels::si_centroids
-od = simodels::si_od_census
-tm_shape(zones) + tm_polygons("all", palette = "viridis")
-```
-
-We'll use the [od](.pkg) package to convert the geographic data to OD data.
-We set the `length` column to the length of the line between the centroids of the zones, in km, and set the coordinate reference system (CRS) to WGS84 (EPSG:4326), as follows:
-
-```{r}
-#| echo: true
-od_df = od::points_to_od(centroids)
-od_sfc = od::odc_to_sfc(od_df[3:6])
-sf::st_crs(od_sfc) = 4326
-od_df$length = sf::st_length(od_sfc)
-od_df = od_df |>
-  transmute(
-    O, D, length = as.numeric(length) / 1000,
-    flow = NA, fc = NA
-  )
-od_df = sf::st_sf(od_df, geometry = od_sfc, crs = 4326)
-```
-
-An unconstrained spatial interaction model can be written as follows, with a more-or-less arbitrary value for `beta` which can be optimised later:
-
-```{r}
-#| label: unconstrained
-#| echo: true
-beta = 0.3
-for(i in seq(nrow(zones))) {
-  for(j in seq(nrow(zones))) {
-    O = zones$all[i]
-    n = zones$all[j]
-    ij = which(od_df$O == zones$geo_code[i] & od_df$D == zones$geo_code[j])
-    od_df$fc[ij] = exp(-beta * od_df$length[ij])
-    od_df$flow[ij] = O * n * od_df$fc[ij]
-  }
-}
-```
-
-The results of the unconstrained SIM are shown in @fig-unconstrained, with the geographic desire lines showing the interaction between zones and a scatter plot illustrating the 'friction of distance'.
-
-```{r}
-#| label: fig-unconstrained
-#| fig-cap: Unconstrained spatial interaction model
-#| fig-subcap: 
-#|   - "Geographic desire lines showing interaction between zones calculated by the unconstrained SIM"
-#|   - "Scatter plot illustrating the 'friction of distance' in the unconstrained SIM"
-#| layout-ncol: 2
-#| echo: false
-od_top = od_df |> 
-  filter(O != D) |> 
-  arrange(desc(flow)) |>
-  slice(seq(2000)) |>
-  arrange(flow)
-
-tm_shape(zones) +
-  tm_borders() +
-  tm_shape(od_top) +
-  tm_lines("flow")
-od_df |> 
-  ggplot() +
-  geom_point(aes(length, fc))
-```
-
-We can make this production constrained by dividing the estimated interaction by the difference between the sum of the estimated interactions and the sum of the observed interactions per origin.
-
-```{r}
-#| label: constrained
-#| include: false
-od_dfj = left_join(
-  od_df,
-  zones |> select(O = geo_code, all) |> sf::st_drop_geometry()
-)
-od_dfj = od_dfj |> 
-  group_by(O) |> 
-  mutate(flow_constrained = flow / sum(flow) * first(all)) |>
-  ungroup()
-sum(od_dfj$flow_constrained) == sum(zones$all)
-od_top = od_dfj |> 
-  filter(O != D) |> 
-  arrange(desc(flow_constrained)) |>
-  slice(seq(2000)) |>
-  arrange(flow_constrained)
-
-tm_shape(zones) +
-  tm_borders() +
-  tm_shape(od_top) +
-  tm_lines("flow_constrained")
-```
-
-## OD data preparation {#sec-od}
-
-## Interaction calculation {#sec-interaction}
+Note: could become a vignette in the future.
 
@@
 ```
-
-## Interaction modelling {#sec-models}
-
-## Examples {#sec-examples}
-
-```{r validation}
-od_dfjc = inner_join(od_dfj |> select(-all), od)
-od_dfjc |> 
-  ggplot() +
-  geom_point(aes(all, flow_constrained))
-cor(od_dfjc$all, od_dfjc$flow_constrained)^2
-```
-
-## Conclusions
-
-<!--
-
-From paper-writing docs:
- ## Introduction: Count data regression in R {#sec-intro}
-
-::: callout
-The introduction is in principle "as usual".
-However, it should usually embed both the implemented *methods* and the *software* into the respective relevant literature.
-For the latter both competing and complementary software should be discussed (within the same software environment and beyond), bringing out relative (dis)advantages.
-All software mentioned should be properly `@cited`'d.
-(See also [Using BibTeX] for more details on {{< bibtex >}}.)
-
-For writing about software JSS requires authors to use the markup `[]{.proglang}` (programming languages and large programmable systems), `[]{.pkg}` (software packages), back ticks like \`code\` for code (functions, commands, arguments, etc.).
-
-If there is such markup in (sub)section titles (as above), a plain text version has to be provided in the {{< latex >}} command as well.
-Below we also illustrate how abbrevations should be introduced and citation commands can be employed.
-See the {{< latex >}} code for more details.
-:::
-
-Modeling count variables is a common task in economics and the social sciences.
-The classical Poisson regression model for count data is often of limited use in these disciplines because empirical count data sets typically exhibit overdispersion and/or an excess number of zeros.
-The former issue can be addressed by extending the plain Poisson regression model in various directions: e.g., using sandwich covariances or estimating an additional dispersion parameter (in a so-called quasi-Poisson model).
-Another more formal way is to use a negative binomial (NB) regression.
-All of these models belong to the family of generalized linear models (GLMs).
-However, although these models typically can capture overdispersion rather well, they are in many applications not sufficient for modeling excess zeros.
-Since @Mullahy1986 there is increased interest in zero-augmented models that address this issue by a second model component capturing zero counts.
-An overview of count data models in econometrics, including hurdle and zero-inflated models, is provided in @CameronTrivedi2013.
-
-In [R]{.proglang} @R, GLMs are provided by the model fitting functions [glm]{.fct} in the [stats]{.pkg} package and [glm.nb]{.fct} in the [MASS]{.pkg} package [@VenablesRipley2002] along with associated methods for diagnostics and inference.
-The manuscript that this document is based on [@ZeileisKleiberJackman2008] then introduced hurdle and zero-inflated count models in the functions [hurdle]{.fct} and [zeroinfl]{.fct} in the [pcsl]{.pkg} package @Jackman2015.
-Of course, much more software could be discussed here, including (but not limited to) generalized additive models for count data as available in the [R]{.proglang} packages [mgcv]{.pkg} @Wood2006, [gamlss]{.pkg} @StasinopoulosRigby2007, or [VGAM]{.pkg} @Yee2009.
-
-## Models and software {#sec-models}
-
-The basic Poisson regression model for count data is a special case of the GLM framework @McCullaghNelder1989.
-It describes the dependence of a count response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution $y_i \sim \mathrm{Pois}(\mu_i)$.
-The dependence of the conditional mean $E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a log link and a linear predictor
-
-$$
-\log(\mu_i) \quad = \quad x_i^\top \beta,
-$$ {#eq-mean}
-
-where the regression coefficients $\beta$ are estimated by maximum likelihood (ML) using the iterative weighted least squares (IWLS) algorithm.
-
-::: callout
-TODO: Note that around the equation above there should be no spaces (avoided in the {{< latex >}} code by `%` lines) so that "normal" spacing is used and not a new paragraph started.
-:::
-
-[R]{.proglang} provides a very flexible implementation of the general GLM framework in the function [glm]{.fct} @ChambersHastie1992 in the [stats]{.pkg} package. Its most important arguments are
-
-```r
-glm(formula, data, subset, na.action, weights, offset,
-  family = gaussian, start = NULL, control = glm.control(…),
-  model = TRUE, y = TRUE, x = FALSE, …)
-```
-
-where `formula` plus `data` is the now standard way of specifying regression relationships in [R]{.proglang}/[S]{.proglang} introduced in @ChambersHastie1992.
-The remaining arguments in the first line (`subset`, `na.action`, `weights`, and `offset`) are also standard for setting up formula-based regression models in [R]{.proglang}/[S]{.proglang}.
-The arguments in the second line control aspects specific to GLMs while the arguments in the last line specify which components are returned in the fitted model object (of class [glm]{.class} which inherits from [lm]{.class}).
-For further arguments to [glm]{.fct} (including alternative specifications of starting values) see `?glm`.
-For estimating a Poisson model `family = poisson` has to be specified.
-
-::: callout
-As the synopsis above is a code listing that is not meant to be executed, one can use either the dedicated `{Code}` environment or a simple `{verbatim}` environment for this.
-Again, spaces before and after should be avoided.
-
-Finally, there might be a reference to a `{table}` such as @tbl-overview.
-Usually, these are placed at the top of the page (`[t!]`), centered (`\centering`), with a caption below the table, column headers and captions in sentence style, and if possible avoiding vertical lines.
-:::
-
-| Type           | Distribution | Method   | Description                                                                                                                                                                                  |
-|---------|---------|---------|-----------------------------------------------|
-| GLM            | Poisson      | ML       | Poisson regression: classical GLM, estimated by maximum likelihood (ML)                                                                                                                      |
-|                |              | Quasi    | "Quasi-Poisson regression'': same mean function, estimated by quasi-ML (QML) or equivalently generalized estimating equations (GEE), inference adjustment via estimated dispersion parameter |
-|                |              | Adjusted | "Adjusted Poisson regression'': same mean function, estimated by QML/GEE, inference adjustment via sandwich covariances                                                                      |
-|                | NB           | ML       | NB regression: extended GLM, estimated by ML including additional shape parameter                                                                                                            |
-| Zero-augmented | Poisson      | ML       | Zero-inflated Poisson (ZIP), hurdle Poisson                                                                                                                                                  |
-|                | NB           | ML       | Zero-inflated NB (ZINB), hurdle NB                                                                                                                                                           |
-: Overview of various count regression models.
-The table is usually placed at the top of the page (`[t!]`), centered (`centering`), has a caption below the table, column headers and captions are in sentence style, and if possible vertical lines should be avoided.
-{#tbl-overview}
-
-## Illustrations {#sec-illustrations}
-
-For a simple illustration of basic Poisson and NB count regression the
-`quine` data from the [MASS]{.pkg} package is used. This provides the number
-of `Days` that children were absent from school in Australia in a
-particular year, along with several covariates that can be employed as regressors. The data can be loaded by
-
-```{R}
-#| prompt: true
-data("quine", package = "MASS")
-```
-and a basic frequency distribution of the response variable is displayed in
-@fig-quine.
-
-:::{.callout}
-For code input and output, the style files provide dedicated environments.
-Either the "agnostic" `{CodeInput}` and `{CodeOutput}` can be used
-or, equivalently, the environments `{Sinput}` and `{Soutput}` as
-produced by [Sweave]{.fct} or [knitr]{.pkg} when using the `render_sweave()`
-hook. Please make sure that all code is properly spaced, e.g., using
-`y = a + b * x` and _not_ `y=a+b*x`. Moreover, code input should use "the usual" command prompt in the respective software system. For
-[R]{.proglang} code, the prompt `R> ` should be used with `+  ` as
-the continuation prompt. Generally, comments within the code chunks should be
-avoided -- and made in the regular {{< latex >}} text instead. Finally, empty lines before and after code input/output should be avoided (see above).
-:::
-
-::: {#fig-quine}
-
-![](article-visualization.pdf)
-
-Frequency distribution for number of days absent from school.
-:::
-
-As a first model for the `quine` data, we fit the basic Poisson regression
-model. (Note that JSS prefers when the second line of code is indented by two
-spaces.)
-
-```{R}
-#| prompt: true
-m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine, family = poisson)
-```
-
-To account for potential overdispersion we also consider a negative binomial
-GLM.
-
-```{R}
-#| prompt: true
-library("MASS")
-m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
-```
-
-In a comparison with the BIC the latter model is clearly preferred.
-
-```{R}
-#| prompt: true
-library("MASS")
-BIC(m_pois, m_nbin)
-```
-
-Hence, the full summary of that model is shown below.
-
-```{R}
-#| prompt: true
-summary(m_nbin)
-```
-
-## Summary and discussion {#sec-summary}
-
-:::{.callout}
-
-As usual…
-
-:::
-
-## Computational details {.unnumbered}
-
-:::{.callout}
-
-If necessary or useful, information about certain computational details
-such as version numbers, operating systems, or compilers could be included
-in an unnumbered section. Also, auxiliary packages (say, for visualizations,
-maps, tables, …) that are not cited in the main text can be credited here.
-
-:::
-
-The results in this paper were obtained using [R]{.proglang}~3.4.1 with the
-[MASS]{.pkg}~7.3.47 package. [R]{.proglang} itself and all packages used are available from the Comprehensive [R]{.proglang} Archive Network (CRAN) at
-[https://CRAN.R-project.org/].
-
-## Acknowledgments {.unnumbered}
-
-:::{.callout}
-
-All acknowledgments (note the AE spelling) should be collected in this
-unnumbered section before the references. It may contain the usual information
-about funding and feedback from colleagues/reviewers/etc. Furthermore,
-information such as relative contributions of the authors may be added here
-(if any).
-
-:::
-
-## References {.unnumbered}
-
-:::{#refs}
-
-:::
-
-{{< pagebreak >}}
-
-## More technical details {#sec-techdetails .unnumbered}
-
-:::{.callout}
-
-Appendices can be included after the bibliography (with a page break). Each
-section within the appendix should have a proper section title (rather than
-just _Appendix_).
-
-For more technical style details, please check out JSS's style FAQ at
-[https://www.jstatsoft.org/pages/view/style#frequently-asked-questions]
-which includes the following topics:
-
-- Title vs. sentence case.
-- Graphics formatting.
-- Naming conventions.
-- Turning JSS manuscripts into [R]{.proglang} package vignettes.
-- Trouble shooting.
-- Many other potentially helpful details…
-
-:::
-
-## Using BibTeX {#sec-bibtex .unnumbered}
-
-:::{.callout}
-
-References need to be provided in a {{< bibtex >}} file (`.bib`). All
-references should be made with `@cite` syntax. This commands yield different
-formats of author-year citations and allow to include additional details (e.g.,pages, chapters, \dots) in brackets. In case you are not familiar with these
-commands see the JSS style FAQ for details.
-
-Cleaning up {{< bibtex >}} files is a somewhat tedious task -- especially
-when acquiring the entries automatically from mixed online sources. However,
-it is important that informations are complete and presented in a consistent
-style to avoid confusions. JSS requires the following format.
-
-- item JSS-specific markup (`\proglang`, `\pkg`, `\code`) should be used in the references.
-- item Titles should be in title case.
-- item Journal titles should not be abbreviated and in title case.
-- item DOIs should be included where available.
-- item Software should be properly cited as well. For [R]{.proglang} packages `citation("pkgname")` typically provides a good starting point.
-
-::: -->

### 2024-08-19 data-raw/si_oa_wpz.qmd
Merge code from Malcolm

---
cache: true
---

```{r}
library(sf)
library(tidyverse)
library(tmap)
tmap_mode("view")
```

```{r}
od_data_oa_wpz_all = readr::read_csv("wf01aew_oa_v1.zip", col_names = c("OA11CD", "wz11cd", "n"))

# Workplace zone boundaries:
# https://geoportal.statistics.gov.uk/datasets/5ed6104f204944b6b60780842cbdaa5c_0/explore
wpz_all = sf::read_sf("Workplace_Zones_Dec_2011_GCB_in_England_and_Wales_2022_4617806297229452276.geojson")
wpz_all = wpz_all |>
  sf::st_transform("EPSG:4326")
names(wpz_all)
# [1] "OBJECTID" "wz11cd"   "lad11cd"  "lad11nm"  "lad11nmw" "GlobalID" "geometry"
# Keep only wz11cd and geometry:
wpz_all = wpz_all["wz11cd"]

summary(unique(od_data_oa_wpz_all$wz11cd) %in% wpz_all$wz11cd)
# and the other way around:
summary(unique(wpz_all$wz11cd) %in% od_data_oa_wpz_all$wz11cd)

# Take a 3km radius around Leeds:
sf::sf_use_s2(FALSE)
wpz_centroids = sf::st_centroid(wpz_all)
sf::sf_use_s2(TRUE)
leeds_boundary = zonebuilder::zb_zone("Leeds", n_circles = 2)
wpz_centroids_leeds = wpz_centroids[leeds_boundary, ]

# tmap::qtm(wpz_centroids_leeds)
si_wpz_leeds = wpz_all |>
  filter(wz11cd %in% wpz_centroids_leeds$wz11cd)
qtm(si_wpz_leeds)
```

```{r}
# Get OA data:
oa_all = sf::read_sf("/data/bronze/geoportal.gov.uk/Output_Areas_Dec_2011_Boundaries_EW_BGC_2022_7971430631129549274.gpkg")
oa_all = oa_all |>
  sf::st_transform("EPSG:4326")
oa_all_centroids = sf::st_centroid(oa_all)
oa_centroids_leeds = oa_all_centroids[leeds_boundary, ]
si_oa_leeds = oa_all |>
  filter(OA11CD %in% oa_centroids_leeds$OA11CD) |>
  select(OA11CD)
qtm(si_oa_leeds)
```

```{r}
#| label: od-data-leeds
si_oa_wpz = od_data_oa_wpz_all |>
  filter(wz11cd %in% si_wpz_leeds$wz11cd & OA11CD %in% si_oa_leeds$OA11CD)
sum(si_oa_wpz$n)
```

```{r}
#| label: od-origin-destination
si_oa_wpz_o = si_oa_wpz |>
  group_by(OA11CD) |>
  summarise(n = sum(n)) |>
  rename(n_o = n)
# join with origin data:
si_oa_wpz_o = si_oa_leeds |>
  left_join(si_oa_wpz_o) 
# tm_shape(si_oa_wpz_o) +
#   tm_polygons(col = "n_o")
plot(si_oa_wpz_o["n_o"])
```

```{r}
#| label: od-destination
si_oa_wpz_d = si_oa_wpz |>
  group_by(wz11cd) |>
  summarise(n = sum(n)) |>
  rename(n_d = n)

si_oa_wpz_d = si_wpz_leeds |>
    left_join(si_oa_wpz_d)

plot(si_oa_wpz_d["n_d"])
```

Now check the data:

```{r}
si_oa_wpz
si_oa_wpz_o
si_oa_wpz_d
```

More logical column names:

```{r}
si_oa_wpz_o = sf::st_sf(
  sf::st_drop_geometry(si_oa_wpz_o),
  geometry = sf::st_geometry(si_oa_wpz_o)
)
si_oa_wpz_d = sf::st_sf(
  sf::st_drop_geometry(si_oa_wpz_d),
  geometry = sf::st_geometry(si_oa_wpz_d)
)
sum(si_oa_wpz_o$n_o)
sum(si_oa_wpz_d$n_d)
sum(si_oa_wpz$n)
```

Save with usethis:

```{r}
usethis::use_data(si_oa_wpz, overwrite = TRUE)
usethis::use_data(si_oa_wpz_o, overwrite = TRUE)
usethis::use_data(si_oa_wpz_d, overwrite = TRUE)
```

### 2024-08-19 data-raw/test-si-to-od.qmd
Merge code from Malcolm

---
format: gfm
---

```{bash}
#| eval: false
# Get data from the web
gh release 
```

```{r}
#| eval: false
piggyback::pb_list(repo = "itsleeds/atumie")
piggyback::pb_download("zones_test.gpkg", repo = "itsleeds/atumie")
piggyback::pb_download("social_grid.gpkg", repo = "itsleeds/atumie")
piggyback::pb_upload("zones_test.gpkg")
piggyback::pb_upload("social_grid.gpkg")
# Get urls from v0.0.5 release:
url_zones = "https://github.com/Robinlovelace/simodels/releases/download/v0.0.5/zones_test.gpkg"
f_zones = basename(url_zones)
download.file(url_zones, f_zones)
url_grid = "https://github.com/Robinlovelace/simodels/releases/download/v0.0.5/social_grid.gpkg"
download.file(url_grid, basename(url_grid))
```

```{r}
url_zones = "https://github.com/Robinlovelace/simodels/releases/download/v0.0.5/zones_test.gpkg"
zones = sf::read_sf(url_zones)
url_grid = "https://github.com/Robinlovelace/simodels/releases/download/v0.0.5/social_grid.gpkg"
grid = sf::read_sf(url_grid)
```

```{r}
devtools::load_all()
od = si_to_od(zones, grid)
```

```{r}
zonesg = sf::st_sf(
  sf::st_drop_geometry(zones),
  geometry = zones$geom
)
gridg = sf::st_sf(
  sf::st_drop_geometry(grid),
  geometry = grid$geom
)
odg = si_to_od(zonesg, gridg)

gridd = gridg[!duplicated(gridg$grid_id), ]

odg = si_to_od(zonesg, gridd)
```

```{r}
od = od::points_to_od(zones, grid)
```

### 2024-08-19 R/data.R

@@ -9,8 +9,8 @@
 #' @aliases si_centroids
 #' @examples
 #' si_zones
-#' plot(si_zones$geometry)
-#' plot(si_centroids$geometry, add = TRUE)
+#' sf:::plot.sfg(si_zones$geometry)
+#' sf:::plot.sfg(si_centroids$geometry, add = TRUE)
 NULL
 
 #' Example OD dataset
@@ -59,4 +59,67 @@ NULL
 #' @name od_aus
 #' @examples
 #' head(od_aus)
+NULL
+
+#' Example zones dataset: administrative zones of York
+#' 
+#' See data-raw/zones_york.qmd for details on the data source.
+#' 
+#' @docType data
+#' @keywords datasets
+#' @name zones_york
+#' @examples
+#' head(zones_york)
+#' sf:::plot.sfg(zones_york$geometry)
+NULL
+
+#' Example destinations dataset: schools in York
+#' 
+#' Example dataset from York, UK
+#' 
+#' See data-raw/zones_york.qmd for details on the data source.
+#' 
+#' @docType data
+#' @keywords datasets
+#' @name destinations_york
+#' @examples
+#' head(destinations_york)
+NULL
+
+#' Origin-Destination Data for Leeds
+#' 
+#' This dataset contains origin-destination data for Leeds, including the number of trips between output areas (OAs) and workplace zones (WPZs).
+#' 
+#' @docType data
+#' @keywords datasets
+#' @name si_oa_wpz
+#' @examples
+#' head(si_oa_wpz)
+NULL
+
+#' Origin Data for Leeds
+#' 
+#' This dataset contains the number of trips originating from each output area (OA) in Leeds.
+#' 
+#' @docType data
+#' @keywords datasets
+#' @name si_oa_wpz_o
+#' @examples
+#' head(si_oa_wpz_o)
+#' sf:::plot.sf(si_oa_wpz_o["n_o"])
+NULL
+
+#' Destination Data for Leeds
+#' 
+#' This dataset contains the number of trips destined for each workplace zone (WPZ) in Leeds.
+#' 
+#' See [wicid.ukdataservice.ac.uk](https://wicid.ukdataservice.ac.uk/flowdata/cider/wicid/downloads.php) for details on the data source
+#' and [data-raw/leeds_wpz.R](data-raw/si_oa_wpz.qmd) for details on how the example dataset was generated.
+#' 
[…]

### 2024-08-19 R/si_model.R

@@ -27,7 +27,7 @@
 #' @examples
 #' od = si_to_od(si_zones, si_zones, max_dist = 4000)
 #' fun_dd = function(d = "distance_euclidean", beta = 0.3) exp(-beta * d / 1000)
-#' fun_dd(d = (1:5)*1000)
+#' fun_dd(d = (1:5) * 1000)
 #' od_dd = si_calculate(od, fun = fun_dd, d = distance_euclidean)
 #' plot(od$distance_euclidean, od_dd$interaction)
 #' fun = function(O, n, d, beta) O * n * exp(-beta * d / 1000)
@@ -37,6 +37,14 @@
 #' plot(od$distance_euclidean, od_output$interaction)
 #' od_pconst = si_calculate(od, fun = fun, beta = 0.3, O = origin_all,
 #'   n = destination_all, d = distance_euclidean, constraint_production = origin_all)
+#' # Origin totals in OD data should equal origin totals in zone data
+#' library(dplyr)
+#' origin_totals_zones = od_pconst |>
+#'   group_by(geo_code = O) |>
+#'   summarise(all_od = sum(interaction)) |>
+#'   sf::st_drop_geometry()
+#' zones_joined = left_join(si_zones, origin_totals_zones)
+#' plot(zones_joined$all, zones_joined$all_od)
 #' plot(od_pconst$distance_euclidean, od_pconst$interaction)
 #' plot(od_pconst["interaction"], logz = TRUE)
 #' od_dd = si_calculate(od, fun = fun_dd, d = distance_euclidean, output_col = "res")

### 2024-08-21 R/si_to_od.R
Update si_to_od() to include max_dist in {od}, close #38

@@ -46,28 +46,15 @@
 #' plot(odsf)
 si_to_od = function(origins, destinations, max_dist = Inf, intrazonal = TRUE) {
   if(identical(origins, destinations)) {
-    od_df = od::points_to_od(origins)
+    od_df = od::points_to_od(origins, max_dist = max_dist)
   } else {
-    od_df = od::points_to_od(origins, destinations)
+    od_df = od::points_to_od(origins, destinations, max_dist = max_dist)
   }
   od_df$distance_euclidean = geodist::geodist(
     x = od_df[c("ox", "oy")],
     y = od_df[c("dx", "dy")],
     paired = TRUE
   )
-  # Max dist check
-  if(max(od_df$distance_euclidean) > max_dist) {
-    nrow_before = nrow(od_df)
-    od_df = od_df[od_df$distance_euclidean <= max_dist, ]
-    nrow_after = nrow(od_df)
-    pct_kept = round(nrow_after / nrow_before * 100)
-    message(
-      nrow_after,
-      " OD pairs remaining after removing those with a distance greater than ", # nolint
-      max_dist, " meters", ":\n",
-      pct_kept, "% of all possible OD pairs"
-    )
-  }
   
   # Intrazonal check
   if(!intrazonal){

### 2024-08-21 R/si_to_od.R
Add checks for nngeo

@@ -1,21 +1,21 @@
 #' Prepare OD data frame
 #'
 #' Prepares an OD data frame that next could be used to
-#' estimate movement between origins and destinations 
+#' estimate movement between origins and destinations
 #' with a spatial interaction model.
-#' 
+#'
 #' In most origin-destination datasets the spatial entities that constitute
 #' origins (typically administrative zones) also represent destinations.
 #' In this 'unipartite' case `origins` and `destinations` should be passed
 #' the same object, an `sf` data frame representing administrative zones.
-#' 
+#'
 #' 'Bipartite' datasets, by contrast, represent
 #' "spatial interaction systems where origins cannot act as
-#' destinations and vice versa" (Hasova et al. [2022](https://lenkahas.com/files/preprint.pdf)). 
-#' 
+#' destinations and vice versa" (Hasova et al. [2022](https://lenkahas.com/files/preprint.pdf)).
+#'
 #'  a different
 #' `sf` object can be passed to the `destinations` argument.
-#' 
+#'
 #' @param origins `sf` object representing origin locations/zones
 #' @param destinations `sf` object representing destination locations/zones
 #' @param max_dist Euclidean distance in meters (numeric).
@@ -45,26 +45,52 @@
 #' nrow(odsf) # no intrazonal flows
 #' plot(odsf)
 si_to_od = function(origins, destinations, max_dist = Inf, intrazonal = TRUE) {
-  if(identical(origins, destinations)) {
-    od_df = od::points_to_od(origins, max_dist = max_dist)
-  } else {
-    od_df = od::points_to_od(origins, destinations, max_dist = max_dist)
+  nngeo_installed = requireNamespace("nngeo", quietly = TRUE)
+  if (!nngeo_installed && max_dist < Inf) {
+    message("The nngeo package is required for fast nearest neighbor search")
   }
+  if (max_dist < Inf && nngeo_installed) {
+    if (identical(origins, destinations)) {
+      od_df = od::points_to_od(origins, max_dist = max_dist)
+    } else {
+      od_df = od::points_to_od(origins, destinations, max_dist = max_dist)
+    }
+  } else {
+    if (identical(origins, destinations)) {
+      od_df = od::points_to_od(origins)
+    } else {
+      od_df = od::points_to_od(origins, destinations)
+    }
+  }
+
[…]

### 2024-08-21 R/data.R
Update pkg docs

@@ -114,7 +114,7 @@ NULL
 #' This dataset contains the number of trips destined for each workplace zone (WPZ) in Leeds.
 #' 
 #' See [wicid.ukdataservice.ac.uk](https://wicid.ukdataservice.ac.uk/flowdata/cider/wicid/downloads.php) for details on the data source
-#' and [data-raw/leeds_wpz.R](data-raw/si_oa_wpz.qmd) for details on how the example dataset was generated.
+#' and the file `data-raw/si_oa_wpz.qmd` in the package repo for details on how the example dataset was generated.
 #' 
 #' @docType data
 #' @keywords datasets

### 2024-08-28 paper/paper.qmd
Tweak paper for pdf version

--- 2024-08-02 paper/paper.qmd
@@
 title: "simodels: A package for Spatial Interaction Modelling"
+# Note: you need to have the relevant quarto extension installed:
+# quarto add quarto-journals/jss
+# After that put the following in the yaml:
 format:
@@
 ---
+
+<!--
+
+-->
+
+```{r}
+#| include: false
+knitr::opts_chunk$set(
+  message = FALSE,
+  warning = FALSE,
+  echo = FALSE
+)
+```
 

### 2024-08-28 paper/paper.qmd
Update paper.qmd in conversation with @SebnemEr

@@
 
+## Interaction modelling {#sec-models}
+
+## Examples {#sec-examples}
+
+```{r validation}
+# od_dfjc = inner_join(od_dfj |> select(-all), od)
+# od_dfjc |> 
+#   ggplot() +
+#   geom_point(aes(all, flow_constrained))
+# cor(od_dfjc$all, od_dfjc$flow_constrained)^2
+```
+
+### Commuter flows in Leeds, UK
+
+See https://github.com/acteng/netgen/blob/main/odgen.md for ideas on this with pupils data.
+
+And https://github.com/acteng/netgen/pull/10/files
+
+```{r}
+od_observed = simodels::si_oa_wpz
+dim(od_observed)
+zones = simodels::si_oa_wpz_o
+destinations = simodels::si_oa_wpz_d
+```
+
+```{r}
+od_modelled_max_5km = si_to_od(zones, destinations)
+od_modelled = si_to_od(zones, destinations, max_dist = 5000)
+dim(od_modelled)
+names(od_modelled)
+```
+
+```{r}
+#| label: simple-model
+#| echo: true
+gravity_model = function(beta, d, m, n) {
+  m * n * exp(-beta * d / 1000)
+} 
+# perform SIM
+od_res = simodels::si_calculate(
+  od_modelled,
+  fun = gravity_model,
+  d = distance_euclidean,
+  m = origin_n_o,
+  n = destination_n_d,
+#   constraint_production = origin_all,
+  beta = 0.8
+  )
+od_res_df = od_res |>
+  sf::st_drop_geometry()
+```
+
+```{r}
+names(od_observed)
+od_joined = left_join(
+  od_res_df,
+  od_observed |>
+    select(O = OA11CD, D = wz11cd, n_observed = n)
+)
+```
+
+```{r}
+#| label: plot_od_fit
+#| include: false
+# Aim: create function that takes in od_res and returns a ggplot object
+plot_od_fit = function(od_res, title = "(unconstrained)") {
+  res_o = od_res |>
+    group_by(O) |>
+    summarise(
+      Observed = first(origin_n_o),
+      Modelled = sum(interaction),
+      Type = "Origin"
+    )
+  res_d = od_res |>
+    group_by(D) |>
+    summarise(
+      Observed = first(destination_n_d),
+      Modelled = sum(interaction),
+      Type = "Destination"
+    )
+  res_od = od_res |>
+    transmute(
+      Observed = n_observed,
+      Modelled = interaction,
+      Type = "OD"
+    )
+  res_combined = bind_rows(res_o, res_d, res_od) |>
+    # Create ordered factor with types:
+    mutate(
+      Type = factor(Type, levels = c("Origin", "Destination", "OD"))
+    )
+  rsq_o = cor(res_o$Observed, res_o$Modelled, use = "complete.obs")^2
+  rsq_d = cor(res_d$Observed, res_d$Modelled, use = "complete.obs")^2
+  rsq_od = cor(res_od$Observed, res_od$Modelled, use = "complete.obs")^2
+  rsq_summary = data.frame(
+    rsq = c(rsq_o, rsq_d, rsq_od) |> round(3),
+    Type = c("Origin", "Destination", "OD")
+  )
+  g_combined = res_combined |>
+    left_join(rsq_summary) |>
+    # Add rsquared info:
+    mutate(Type = paste0(Type, " (R-squared: ", rsq, ")")) |>
+    # Update factor so it's ordered (reverse order):
+    mutate(
+      Type = factor(Type, levels = unique(Type))
+    ) |>
+    ggplot() +
+      geom_point(aes(x = Observed, y = Modelled)) +
+      geom_smooth(aes(x = Observed, y = Modelled), method = "lm") +
+      facet_wrap(~Type, scales = "free") +
+      labs(
+        title = paste0("Model fit at origin, destination and OD levels ", title),
+        x = "Observed",
+        y = "Modelled"
+      )
+  g_combined
+}
+# Test it:
+g = plot_od_fit(od_joined)
+```
+
+```{r}
+#| label: plot-od-fit
+#| fig-cap: Model fit at origin, destination and OD levels
+g
+```
+
+### Travel to school in York, UK
+
 ```{r}
@@
     )
-```
-
-## Interaction modelling {#sec-models}
-
-## Examples {#sec-examples}
-
-```{r validation}
-od_dfjc = inner_join(od_dfj |> select(-all), od)
-od_dfjc |> 
-  ggplot() +
-  geom_point(aes(all, flow_constrained))
-cor(od_dfjc$all, od_dfjc$flow_constrained)^2
 ```

### 2024-09-02 paper/paper.qmd
Updates after conversation with @adamdennett

@@
 title: "simodels: A package for Spatial Interaction Modelling"
-# Note: you need to have the relevant quarto extension installed:
+bibliography: bibliography.bib
+execute: 
+  messages: false
+  warning: false 
+  echo: false
+---
+
+```{r}
+#| eval: false
+#| echo: false
+# Compile to docx and jss-pdf formats:
+quarto::quarto_render("paper/paper.qmd", output_format = "docx")
+quarto::quarto_render("paper/paper.qmd", output_format = "jss-pdf")
+system("gh release upload v0.1.0 paper/paper.docx --clobber")
+system("gh release upload v0.1.0 paper/paper.pdf --clobber")
+```
+
+<!-- 
+# Note: you need to have the relevant quarto extension installed for the JSS version:
 # quarto add quarto-journals/jss
@@
 keywords-formatted: [JSS, style guide, comma-separated, not capitalized, "[R]{.proglang}"]
+-->
 
-bibliography: bibliography.bib
-execute: 
-  messages: false
-  warning: false 
-  echo: false
----
+## Ideas
 
-<!--
-
--->
+- Conceptual diagram showing decision points when developing SIMs, e.g.:
+  - Unipartite or bipartite?
+  - Unconstrained, production-constrained, attraction-constrained, or doubly-constrained? Or other?
+  - Calculation and parameter optimisation or regression model?
+- Open access dataset?
 
@@
 
-## Basic implementation of SIMs {#sec-base}
+### Basic implementation of SIMs {#sec-base}
 
@@
 
-## OD data preparation {#sec-od}
+## The simodels package {#sec-simodels}
 
-## Interaction calculation {#sec-interaction}
+### OD data preparation {#sec-od}
 
-## Interaction modelling {#sec-models}
+### Interaction calculation {#sec-interaction}
+
+### Interaction modelling {#sec-models}
+
+### Optimising/parameter estimation {#sec-optimisation}
+
+See https://github.com/acteng/netgen/blob/main/odgen.md#optimising-the-value-of-beta for ideas on this.
+
+### Network distance measures {#sec-network}
 
@@
 
-### Commuter flows in Leeds, UK
+### Commuter flows between large areas in Leeds, UK
+
+```{r}
+od = simodels::si_od_census
+zones = simodels::si_zones
+nrow(od)
+sum(od$all)
+sum(zones$all)
+```
+
+### Commuter flows between small areas in Leeds, UK
 
@@
 ```{r}
-od_modelled_max_5km = si_to_od(zones, destinations)
-od_modelled = si_to_od(zones, destinations, max_dist = 5000)
+#| echo: false
+#| eval: false
+od_observed_o = od_observed |> 
+  group_by(OA11CD) |> 
+  summarise(n = sum(n))
+od_observed_d = od_observed |> 
+  group_by(wz11cd) |> 
+  summarise(n = sum(n))
+zones_joined = left_join(
+  zones |>
+    select(OA11CD, n_o) |>
+    sf::st_drop_geometry(),
+  od_observed_o
+)
+head(zones_joined)
+cor(zones_joined$n_o, zones_joined$n)
+
+# For destinations:
+destinations_joined = left_join(
+  destinations |>
+    select(wz11cd, n_d) |>
+    sf::st_drop_geometry(),
+  od_observed_d
+)
+cor(destinations_joined$n_d, destinations_joined$n)
+```
+
+```{r}
+# TODO: test if the max_dist arg actually makes it faster for large datasets
+system.time({
+  od_modelled = si_to_od(zones, destinations)
+})
+od_modelled = si_to_od(zones, destinations)
+system.time({
+  od_modelled_max_5km = si_to_od(zones, destinations, max_dist = 3000)
+})
+od_modelled_max_5km = si_to_od(zones, destinations, max_dist = 3000)
 dim(od_modelled)
 names(od_modelled)
+dim(od_modelled_max_5km)
 ```

### 2024-09-04 paper/paper.qmd
Describe Leeds commute example

@@
 bibliography: bibliography.bib
+number-sections: true
 execute: 
@@
 
+The package comes with an example OD dataset from Leeds, UK, which can be used to demonstrate the SIMs.
+Let's model travel to work in a major UK city.
+We'll first load the inputs, representing administrative zones in Leeds (`si_zones`) and the OD data (`si_od_census`), which we will rename for clarity:
+
 ```{r}
+#| echo: true
 od = simodels::si_od_census
@@
 sum(zones$all)
+names(od)
 ```
+
+As shown in the code and outputs above, the input dataset represents the commutes made by `r sum(od$all) |> format(big.mark = ",")` people in Leeds.
+We have the breakdown of trips by mode, with `r sum(od$car_driver) |> format(big.mark = ",")` people driving to work, `r sum(od$car_passenger) |> format(big.mark = ",")` people car sharing, `r sum(od$bus) |> format(big.mark = ",")`, `r sum(od$bicycle) |> format(big.mark = ",")` cycling, and `r sum(od$foot) |> format(big.mark = ",")` people walking.
+A hypothetical research question to inform policy with reference to this dataset is: "How would travel demand change if transport costs increase?".
+Before answering this question, we need to develop a baseline model to predict the travel between origins and destinations as a function of distance (and other factors).
+<!-- (Note: key other factors to test are 'network distance' representing intervening opportunities, and 'population' representing the number of potential travellers). -->
 

### 2024-09-07 paper/paper.qmd
arbitrary
Refactor

--- 2024-08-28 paper/paper.qmd
@@
 title: "simodels: A package for Spatial Interaction Modelling"
+---
+
+<!-- 
 # Note: you need to have the relevant quarto extension installed:
-# quarto add quarto-journals/jss
-# After that put the following in the yaml:
+quarto add quarto-journals/jss
+After that put the following in the yaml:
 format:
@@
   This article introduces the [simodels]{.pkg} package for spatial interaction modelling in R.
-The package provides functions for estimating spatial interaction models, including unconstrained gravity models and models that are production-, attraction-, or doubly-constrained.
+The package provides a flexible framework for estimating various types of spatial interaction models, including unconstrained gravity models and models that are production-, attraction-, or doubly-constrained.
+[simodels]{.pkg} offers efficient implementations of traditional spatial interaction models as well as more recent innovations, supporting both 'monopartite' and 'bipartite' spatial interaction scenarios.
+By streamlining the process of implementing SIMs in R, the package enables researchers and practitioners to more easily explore, analyze, and predict spatial interactions in various contexts, from transportation planning to economic geography.
 
-keywords: [JSS, style guide, comma-separated, not capitalized, R]
-keywords-formatted: [JSS, style guide, comma-separated, not capitalized, "[R]{.proglang}"]
+keywords: [spatial interaction models, R, transportation modelling, gravity models, origin-destination data]
+keywords-formatted: [spatial interaction models, "[R]{.proglang}", transportation modelling, gravity models, origin-destination data]
 
@@
   echo: false
----
-
-<!--
-
 -->
@@
 Spatial Interaction Models (SIMs) are mathematical models for estimating movement between spatial entities.
-First developed in the late 1960s and early 1970, SIMs have become a key tool for transport modelling with substantial practical applications [@boyce_forecasting_2015].
+First developed in the late 1960s and early 1970s, SIMs have become a key tool for transport modelling with substantial practical applications [@boyce_forecasting_2015].
+These models have wide-ranging applications in urban planning, transportation studies, migration analysis, and various other fields where understanding and predicting spatial flows is crucial.
 
-An early and highly influential type of SIM was the 'gravity model', defined by @wilson_family_1971 as follows (in a paper that explored many extensions of this formulation and is influential enough to merit being quoted at length):
+The 'gravity model', an early and highly influential type of SIM, was defined by @wilson_family_1971 as follows:
 
@@
 
-"where $T_{i j}$ is a measure of the interaction between zones $i$ and $W_{i}^{(1)}$ is a measure of the 'mass term' associated with zone $z_i$, $W_{j}^{(2)}$ is a measure of the 'mass term' associated with zone $z_j$, and $c_{ij}$ is a measure of the distance, or generalised cost of travel, between zone $i$ and zone $j$".
-$K$ is a 'constant of proportionality' and $n$ is a parameter to be estimated.
+where $T_{i j}$ is a measure of the interaction between zones $i$ and $j$, $W_{i}^{(1)}$ and $W_{j}^{(2)}$ are 'mass terms' associated with zones $i$ and $j$ respectively, $c_{ij}$ is a measure of the distance or generalized cost of travel between the zones, $K$ is a 'constant of proportionality', and $n$ is a parameter to be estimated.
 
-We follow @simini_universal_2012 in replacing the $W$ terms above with $m$ and $n$ for origins and destinations respectively.
-This seminal definition of the unconstrained 'gravity model' can be written in its "multiplicative form" as follows @dennett_modelling_2018, notation that we will use in this paper:
+Following @simini_universal_2012, we replace the $W$ terms with $m$ and $n$ for origins and destinations respectively.
+The unconstrained 'gravity model' can then be written in its "multiplicative form" as:
 
@@
 
-It is unconstrained because there is no limit on the amount of interaction (we will refer to the more concrete but semantically equivalent term "number of trips") in this paper.
+where $k$ is a balancing factor, $m_i$ and $n_j$ are origin and destination masses respectively, $d_{ij}$ is the distance between $i$ and $j$, and $\beta$ is a distance decay parameter [@dennett_modelling_2018].
+
 There are four main types of traditional SIMs [@wilson_family_1971]:
 
-- Unconstrained
-- Production-constrained
-- Attraction-constrained
-- Doubly-constrained
+1. Unconstrained
+2. Production-constrained
+3. Attraction-constrained
+4. Doubly-constrained
 
-For the purposes of this project, we will focus on production-constrained SIMs.
-These can be defined as follows [@wilson_family_1971]:
+This paper focuses primarily on production-constrained SIMs, defined as:
 
@@
 
-where $A$ is a balancing factor defined as:
+where $A_i$ is a balancing factor:
 
@@
 
-$O_i$ is analogous to the travel demand in zone $i$, which can be roughly approximated by its population.
+and $O_i$ represents the total outflows from origin $i$, which can be approximated by its population in many applications.
 
-More recent innovations in SIMs including the 'radiation model' @simini_universal_2012.
-See @lenormand_systematic_2016 for a comparison of alternative approaches.
-The purpose of the package is not to implement all available models, but to provide a flexible framework to speed-up the process of implementing SIMs in [R]{.proglang}.
+Recent innovations in SIMs include the 'radiation model' [@simini_universal_2012], and various hybrid approaches [@lenormand_systematic_2016].
+While the [simodels]{.pkg} package does not aim to implement all available models, it provides a flexible framework to speed up the process of implementing SIMs in [R]{.proglang}.
+
 The package supports a wide range of models and can be used to fit unconstrained, production-constrained, attraction-constrained, and doubly-constrained models.
-Furthermore, the package works equally well for 'bipartite' SIMs, in which origin and destination points are different (Hasova et al. [2022](https://lenkahas.com/files/preprint.pdf)).
+Furthermore, it works equally well for 'bipartite' SIMs, where origin and destination points are different [@hasova2022].
 
-The paper is structured as follows.
-@sec-base demonstrates how to implement an unconstrained SIM in base [R]{.proglang}, highlighting the need for a package to streamline the process.
-@sec-od describes the process of converting spatial data to Origin-Destination (OD) data, a key step in SIMs, and the functions in the package that facilitate this process.
-@sec-interaction describes the process of calculating interaction with pre-defined parameters with the `si_calculate()` function.
+This paper is structured as follows:
 
-## Basic implementation of SIMs {#sec-base}
+- Section 2 demonstrates how to implement an unconstrained SIM in base [R]{.proglang}, highlighting the need for a package to streamline the process.
+- Section 3 describes the process of converting spatial data to Origin-Destination (OD) data, a key step in SIMs, and the functions in the package that facilitate this process.
+- Section 4 explains the process of calculating interaction with pre-defined parameters using the `si_calculate()` function.
+- Section 5 explores various interaction models that can be implemented using the package.
+- Section 6 provides examples of applying the package to real-world datasets.
+- Section 7 concludes with a summary of the package's capabilities and its potential impact on the field of spatial interaction modeling.
 
-Before describing the functions in the package, it's worth implementing SIMs from first principles, to gain an understanding of how they work and to highlight the benefits of using the package.
+## Basic implementation of SIMs
+
+Before describing the functions in the [simodels]{.pkg} package, it's worth implementing SIMs from first principles to gain an understanding of how they work and to highlight the benefits of using the package.
 The code presented below was written before the functions in the [simodels]{.pkg} package were developed, building on @dennett_modelling_2018.
-The aim is to demonstrate a common way of running SIMs, in a for loop, rather than using vectorised operations used in [simodels]{.pkg}.
-In this section we the [tmap]{.pkg} and [tidyverse]{.pkg} packages for data manipulation and visualisation, using only base [R]{.proglang} functions for the SIM calculations (see the paper's source code for details).
+The aim is to demonstrate a common way of running SIMs using a for loop, rather than the vectorized operations used in [simodels]{.pkg}.
+
+We'll use the [tmap]{.pkg} and [tidyverse]{.pkg} packages for data manipulation and visualization, using only base [R]{.proglang} functions for the SIM calculations.
 
@@
 
-The package contains a number of datasets that can be used to demonstrate the SIMs, as shown below:
+The [simodels]{.pkg} package contains several datasets that can be used to demonstrate SIMs:
 
@@
 
-We'll use the [od](.pkg) package to convert the geographic data to OD data.
-We set the `length` column to the length of the line between the centroids of the zones, in km, and set the coordinate reference system (CRS) to WGS84 (EPSG:4326), as follows:
+We'll use the [od]{.pkg} package to convert the geographic data to OD data.
+We set the `length` column to the length of the line between the centroids of the zones, in km, and set the coordinate reference system (CRS) to WGS84 (EPSG:4326):
 
@@
 
-An unconstrained spatial interaction model can be written as follows, with a more-or-less arbitrary value for `beta` which can be optimised later:
+An unconstrained spatial interaction model can be implemented as follows, with an initial arbitrary value for `beta` which can be optimized later:
 
@@
 
-We can make this production constrained by dividing the estimated interaction by the difference between the sum of the estimated interactions and the sum of the observed interactions per origin.
+These visualizations provide insights into the spatial patterns of interactions predicted by the model.
+The desire lines map shows the strongest predicted flows, while the scatter plot reveals the exponential decay of interaction strength with distance.
+
+We can make this model production-constrained by dividing the estimated interaction by the difference between the sum of the estimated interactions and the sum of the observed interactions per origin:
 
@@
 #| label: constrained
-#| include: false
+#| echo: true
 od_dfj = left_join(
@@
 sum(od_dfj$flow_constrained) == sum(zones$all)
-od_top = od_dfj |> 
-  filter(O != D) |> 
-  arrange(desc(flow_constrained)) |>
-  slice(seq(2000)) |>
-  arrange(flow_constrained)
-
-tm_shape(zones) +
-  tm_borders() +
-  tm_shape(od_top) +
-  tm_lines("flow_constrained")
 ```
 
-## OD data preparation {#sec-od}
+This constraint ensures that the total outflows from each origin in the model match the observed total outflows, providing a more realistic representation of the spatial interactions.
 
-## Interaction calculation {#sec-interaction}
+## OD data preparation
 
-## Interaction modelling {#sec-models}
+Origin-Destination (OD) data is fundamental to spatial interaction modeling.
+It represents the flows or interactions between pairs of geographic locations, typically stored as a matrix or long-format dataset.
+The [simodels]{.pkg} package provides functions to facilitate the preparation and manipulation of OD data.
 
-## Examples {#sec-examples}
-
-```{r validation}
-# od_dfjc = inner_join(od_dfj |> select(-all), od)
-# od_dfjc |> 
-#   ggplot() +
-#   geom_point(aes(all, flow_constrained))
-# cor(od_dfjc$all, od_dfjc$flow_constrained)^2
-```
-
-### Commuter flows in Leeds, UK
-
-See https://github.com/acteng/netgen/blob/main/odgen.md for ideas on this with pupils data.
-
-And https://github.com/acteng/netgen/pull/10/files
-
-```{r}
-od_observed = simodels::si_oa_wpz
-dim(od_observed)
-zones = simodels::si_oa_wpz_o
-destinations = simodels::si_oa_wpz_d
-```
-
-```{r}
-od_modelled_max_5km = si_to_od(zones, destinations)
-od_modelled = si_to_od(zones, destinations, max_dist = 5000)
-dim(od_modelled)
-names(od_modelled)
-```
-
-```{r}
-#| label: simple-model
-#| echo: true
-gravity_model = function(beta, d, m, n) {
-  m * n * exp(-beta * d / 1000)
-} 
-# perform SIM
-od_res = simodels::si_calculate(
-  od_modelled,
-  fun = gravity_model,
-  d = distance_euclidean,
-  m = origin_n_o,
-  n = destination_n_d,
-#   constraint_production = origin_all,
-  beta = 0.8
-  )
-od_res_df = od_res |>
-  sf::st_drop_geometry()
-```
-
-```{r}
-names(od_observed)
-od_joined = left_join(
-  od_res_df,
-  od_observed |>
-    select(O = OA11CD, D = wz11cd, n_observed = n)
-)
-```
-
-```{r}
-#| label: plot_od_fit
-#| include: false
-# Aim: create function that takes in od_res and returns a ggplot object
-plot_od_fit = function(od_res, title = "(unconstrained)") {
-  res_o = od_res |>
-    group_by(O) |>
-    summarise(
-      Observed = first(origin_n_o),
-      Modelled = sum(interaction),
-      Type = "Origin"
-    )
-  res_d = od_res |>
-    group_by(D) |>
-    summarise(
-      Observed = first(destination_n_d),
-      Modelled = sum(interaction),
-      Type = "Destination"
-    )
-  res_od = od_res |>
-    transmute(
-      Observed = n_observed,
-      Modelled = interaction,
-      Type = "OD"
-    )
-  res_combined = bind_rows(res_o, res_d, res_od) |>
-    # Create ordered factor with types:
-    mutate(
-      Type = factor(Type, levels = c("Origin", "Destination", "OD"))
-    )
-  rsq_o = cor(res_o$Observed, res_o$Modelled, use = "complete.obs")^2
-  rsq_d = cor(res_d$Observed, res_d$Modelled, use = "complete.obs")^2
-  rsq_od = cor(res_od$Observed, res_od$Modelled, use = "complete.obs")^2
-  rsq_summary = data.frame(
-    rsq = c(rsq_o, rsq_d, rsq_od) |> round(3),
-    Type = c("Origin", "Destination", "OD")
-  )
-  g_combined = res_combined |>
-    left_join(rsq_summary) |>
-    # Add rsquared info:
-    mutate(Type = paste0(Type, " (R-squared: ", rsq, ")")) |>
-    # Update factor so it's ordered (reverse order):
-    mutate(
-      Type = factor(Type, levels = unique(Type))
-    ) |>
-    ggplot() +
-      geom_point(aes(x = Observed, y = Modelled)) +
-      geom_smooth(aes(x = Observed, y = Modelled), method = "lm") +
-      facet_wrap(~Type, scales = "free") +
-      labs(
-        title = paste0("Model fit at origin, destination and OD levels ", title),
-        x = "Observed",
-        y = "Modelled"
-      )
-  g_combined
-}
-# Test it:
-g = plot_od_fit(od_joined)
-```
-
-```{r}
-#| label: plot-od-fit
-#| fig-cap: Model fit at origin, destination and OD levels
-g
-```
-
-### Travel to school in York, UK
+The `si_to_od()` function is a key tool in the package for generating OD data from spatial datasets. Let's demonstrate its use with the York schools dataset:
 
@@
 #| label: inputs-york
+#| echo: true
 g1 = zones_york |>
@@
 
-Before we run any models let's compare the total number of pupils in the zones dataset and the destinations dataset (they should be the same):
+Before running any models, it's important to compare the total number of pupils in the zones dataset and the destinations dataset:
 
@@
 
-As one would expect, the total number of pupils in the zones dataset is a bit bigger than the total number of pupils in the destinations dataset: not all people aged 0 to 15 go to school, especially those under school age.
-To tackle this issue we'll create a new variables called `pupils_estimated` in the zones dataset, which is the sum of the number of pupils in the zones dataset and the number of pupils in the destinations dataset.
+As expected, the total number of pupils in the zones dataset is slightly larger than the total number of pupils in the destinations dataset, likely due to including children who are not yet of school age.
+To address this, we'll create a new variable called `pupils_estimated` in the zones dataset:
 
@@
 
-After the adjustment shown above, the totals in the origin and destination columns should be the same:
+After this adjustment, the totals in the origin and destination columns should match:
 
@@
 
-Based on these inputs the `si_to_od()` function generates the OD data, as shown below (note: 2 versions are created, one with a maximum distance constraint for speed of processing, important when working with large datasets).
+Now we can use the `si_to_od()` function to generate the OD data:
 
@@
 ```
+
+The function creates two versions of the OD dataset: one with all possible origin-destination pairs, and another with a maximum distance constraint for computational efficiency.
 
@@
 
-The output OD dataset has column names taken from both the origin and destination datasets, with the following column names:
+These plots visualize the difference between the full OD dataset and the distance-constrained version, highlighting the efficiency gains of using a maximum distance threshold in large-scale analyses.
 
-```{r}
-names(od_from_si)
-```
+## Interaction calculation
 
-Let's run a simple model:
+The `si_calculate()` function is a core component of the [simodels]{.pkg} package, allowing for flexible and efficient calculation of spatial interactions.
+Let's demonstrate its use with a simple gravity model:
 
@@
   n = destination_n_pupils,
-#   constraint_production = origin_all,
   beta = 0.9
@@
 
-We'll make one adjustment to the output dataset, renaming the `interaction` column to `trips`, and setting the total number of trips to be the same as the total number of pupils in the destinations dataset:
+The `si_calculate()` function takes several arguments:
+- `od_from_si`: the OD dataset
+- `fun`: the interaction model function
+- `d`, `m`, `n`: column names for distance, origin mass, and destination mass
+- `beta`: the distance decay parameter
+
+We can adjust the output to ensure the total number of trips matches the total number of pupils:
 
@@
 
-We can assess the model fit at three levels: the origin level (number of students departing from each zone), the destination level (the number arriving at each school in the input dataset) and the origin-destination level.
+## Interaction modeling
+
+The [simodels]{.pkg} package supports various types of spatial interaction models.
+In this section, we'll explore unconstrained and production-constrained models, demonstrating how to implement and compare them using the package.
+
+### Unconstrained Model
+
+We've already implemented a basic unconstrained gravity model. Let's assess its performance:
 
@@
 res_combined = bind_rows(res_o, res_d) |>
-  # Create ordered factor with types:
   mutate(
@@
     labs(
-      title = "Model fit at origin, destination and OD levels (unconstrained)",
+      title = "Model fit at origin and destination levels (unconstrained)",
       x = "Observed",
@@
 
-Let's see if making the model production constrained can help:
+This plot shows the model's performance at both origin and destination levels.
+The unconstrained model seems to perform reasonably well, but there's room for improvement, particularly at the origin level.
+
+### Production-Constrained Model
+
+Now, let's implement a production-constrained model using the `si_calculate()` function:
 
@@
 ```
+
+Let's assess the performance of this constrained model:
 
@@
     labs(
-      title = "Model fit at origin, destination and OD levels (production constrained)",
+      title = "Model fit at origin and destination levels (production constrained)",
       x = "Observed",
@@
 
+The production-constrained model shows a perfect fit at the origin level, as expected, while maintaining a good fit at the destination level.
+This demonstrates the power of constrained models in certain applications where preserving total outflows is crucial.
+
+## Examples
+
+To further illustrate the versatility of the [simodels]{.pkg} package, let's explore a few more examples using different datasets and model specifications.
+
+### Example 1: Commuter Flows
+
+Let's use the `si_od_census` dataset, which contains commuter flow data, to demonstrate how to fit a doubly-constrained model.
+
+```{r}
+#| echo: true
+data("si_od_census", package = "simodels")
+data("si_zones", package = "simodels")
+
+# Prepare the data
+od_commute = si_to_od(si_zones, si_zones)
+od_commute = left_join(od_commute, si_od_census, by = c("O", "D"))
+
+# Fit a doubly-constrained model
+dc_model = si_calculate(
+  od_commute,
+  fun = gravity_model,
+  d = distance_euclidean,
+  m = origin_all,
+  n = destination_all,
+  constraint_production = origin_all,
+  constraint_attraction = destination_all,
+  beta = 1.5
+)
+
+# Evaluate the model
+plot(dc_model$interaction, od_commute$all,
+     xlab = "Predicted Flows", ylab = "Observed Flows",
+     main = "Doubly-Constrained Model: Predicted vs Observed Flows")
+abline(0, 1, col = "red")
+```
+
+This example demonstrates how to use the package to model commuter flows with a doubly-constrained model, which ensures that both the total outflows from origins and total inflows to destinations match the observed data.
+
+### Example 2: Distance Decay Function Comparison
+
+The [simodels]{.pkg} package allows for easy comparison of different distance decay functions.
+Let's compare the performance of exponential and power law decay functions using the York schools dataset.
+
+```{r}
+#| echo: true
+# Exponential decay function
+exp_decay = function(beta, d, m, n) {
+  m * n * exp(-beta * d / 1000)
+}
+
+# Power law decay function
+power_decay = function(beta, d, m, n) {
+  m * n * (d / 1000)^(-beta)
+}
+
+# Fit models with different decay functions
+od_exp = si_calculate(od_from_si, fun = exp_decay,
+                      d = distance_euclidean, m = origin_pupils_estimated, n = destination_n_pupils,
+                      beta = 0.9)
+
+od_power = si_calculate(od_from_si, fun = power_decay,
+                        d = distance_euclidean, m = origin_pupils_estimated, n = destination_n_pupils,
+                        beta = 1.5)
+
+# Compare model performance
+rsq_exp = cor(od_exp$interaction, od_from_si$all)^2
+rsq_power = cor(od_power$interaction, od_from_si$all)^2
+
+cat("R-squared (Exponential decay):", rsq_exp, "\n")
+cat("R-squared (Power law decay):", rsq_power, "\n")
+
+# Visualize the decay functions
+d_seq = seq(0, 10, by = 0.1)
+plot(d_seq, exp(-0.9 * d_seq), type = "l", col = "blue",
+     xlab = "Distance (km)", ylab = "Interaction Strength",
+     main = "Comparison of Distance Decay Functions")
+lines(d_seq, d_seq^(-1.5), col = "red")
+legend("topright", legend = c("Exponential", "Power Law"),
+       col = c("blue", "red"), lty = 1)
+```
+
+This example showcases how easily different distance decay functions can be compared using the [simodels]{.pkg} package, allowing researchers to choose the most appropriate function for their specific application.
+
 ## Conclusions
 
-<!--
+The [simodels]{.pkg} package provides a flexible and efficient framework for implementing spatial interaction models in R.
+Through this paper, we've demonstrated its capabilities in handling various types of SIMs, from basic unconstrained gravity models to more complex constrained models.
 
-From paper-writing docs:
- ## Introduction: Count data regression in R {#sec-intro}
+Key features of the package include:
 
-::: callout
-The introduction is in principle "as usual".
-However, it should usually embed both the implemented *methods* and the *software* into the respective relevant literature.
-For the latter both competing and complementary software should be discussed (within the same software environment and beyond), bringing out relative (dis)advantages.
-All software mentioned should be properly `@cited`'d.
-(See also [Using BibTeX] for more details on {{< bibtex >}}.)
+1. Efficient OD data preparation with the `si_to_od()` function.
+2. Flexible interaction calculation using `si_calculate()`, which supports custom model specifications.
+3. Built-in support for different types of constraints (production, attraction, and doubly-constrained).
+4. Easy integration with other R packages for data manipulation and visualization.
 
-For writing about software JSS requires authors to use the markup `[]{.proglang}` (programming languages and large programmable systems), `[]{.pkg}` (software packages), back ticks like \`code\` for code (functions, commands, arguments, etc.).
+The examples provided in this paper illustrate how [simodels]{.pkg} can be applied to real-world datasets, such as school attendance and commuter flows.
+The package's flexibility allows researchers to easily experiment with different model specifications, distance decay functions, and constraints, facilitating more comprehensive analyses of spatial interactions.
 
-If there is such markup in (sub)section titles (as above), a plain text version has to be provided in the {{< latex >}} command as well.
-Below we also illustrate how abbrevations should be introduced and citation commands can be employed.
-See the {{< latex >}} code for more details.
+Future developments of the package could include:
+
+- Implementation of more advanced SIM variants, such as the radiation model.
+- Integration of calibration methods for parameter estimation.
+- Enhanced visualization tools for spatial interaction patterns.
+
+By providing a standardized and extensible framework for spatial interaction modeling, [simodels]{.pkg} has the potential to accelerate research and applications in fields such as transportation planning, urban studies, and economic geography.
+We encourage researchers and practitioners to explore the package and contribute to its ongoing development.
+
+## References
+
+::: {#refs}
 :::
-
-Modeling count variables is a common task in economics and the social sciences.
-The classical Poisson regression model for count data is often of limited use in these disciplines because empirical count data sets typically exhibit overdispersion and/or an excess number of zeros.
-The former issue can be addressed by extending the plain Poisson regression model in various directions: e.g., using sandwich covariances or estimating an additional dispersion parameter (in a so-called quasi-Poisson model).
-Another more formal way is to use a negative binomial (NB) regression.
-All of these models belong to the family of generalized linear models (GLMs).
-However, although these models typically can capture overdispersion rather well, they are in many applications not sufficient for modeling excess zeros.
-Since @Mullahy1986 there is increased interest in zero-augmented models that address this issue by a second model component capturing zero counts.
-An overview of count data models in econometrics, including hurdle and zero-inflated models, is provided in @CameronTrivedi2013.
-
-In [R]{.proglang} @R, GLMs are provided by the model fitting functions [glm]{.fct} in the [stats]{.pkg} package and [glm.nb]{.fct} in the [MASS]{.pkg} package [@VenablesRipley2002] along with associated methods for diagnostics and inference.
-The manuscript that this document is based on [@ZeileisKleiberJackman2008] then introduced hurdle and zero-inflated count models in the functions [hurdle]{.fct} and [zeroinfl]{.fct} in the [pcsl]{.pkg} package @Jackman2015.
-Of course, much more software could be discussed here, including (but not limited to) generalized additive models for count data as available in the [R]{.proglang} packages [mgcv]{.pkg} @Wood2006, [gamlss]{.pkg} @StasinopoulosRigby2007, or [VGAM]{.pkg} @Yee2009.
-
-## Models and software {#sec-models}
-
-The basic Poisson regression model for count data is a special case of the GLM framework @McCullaghNelder1989.
-It describes the dependence of a count response variable $y_i$ ($i = 1, \dots, n$) by assuming a Poisson distribution $y_i \sim \mathrm{Pois}(\mu_i)$.
-The dependence of the conditional mean $E[y_i \, | \, x_i] = \mu_i$ on the regressors $x_i$ is then specified via a log link and a linear predictor
-
-$$
-\log(\mu_i) \quad = \quad x_i^\top \beta,
-$$ {#eq-mean}
-
-where the regression coefficients $\beta$ are estimated by maximum likelihood (ML) using the iterative weighted least squares (IWLS) algorithm.
-
-::: callout
-TODO: Note that around the equation above there should be no spaces (avoided in the {{< latex >}} code by `%` lines) so that "normal" spacing is used and not a new paragraph started.
-:::
-
-[R]{.proglang} provides a very flexible implementation of the general GLM framework in the function [glm]{.fct} @ChambersHastie1992 in the [stats]{.pkg} package. Its most important arguments are
-
-```r
-glm(formula, data, subset, na.action, weights, offset,
-  family = gaussian, start = NULL, control = glm.control(…),
-  model = TRUE, y = TRUE, x = FALSE, …)
-```
-
-where `formula` plus `data` is the now standard way of specifying regression relationships in [R]{.proglang}/[S]{.proglang} introduced in @ChambersHastie1992.
-The remaining arguments in the first line (`subset`, `na.action`, `weights`, and `offset`) are also standard for setting up formula-based regression models in [R]{.proglang}/[S]{.proglang}.
-The arguments in the second line control aspects specific to GLMs while the arguments in the last line specify which components are returned in the fitted model object (of class [glm]{.class} which inherits from [lm]{.class}).
-For further arguments to [glm]{.fct} (including alternative specifications of starting values) see `?glm`.
-For estimating a Poisson model `family = poisson` has to be specified.
-
-::: callout
-As the synopsis above is a code listing that is not meant to be executed, one can use either the dedicated `{Code}` environment or a simple `{verbatim}` environment for this.
-Again, spaces before and after should be avoided.
-
-Finally, there might be a reference to a `{table}` such as @tbl-overview.
-Usually, these are placed at the top of the page (`[t!]`), centered (`\centering`), with a caption below the table, column headers and captions in sentence style, and if possible avoiding vertical lines.
-:::
-
-| Type           | Distribution | Method   | Description                                                                                                                                                                                  |
-|---------|---------|---------|-----------------------------------------------|
-| GLM            | Poisson      | ML       | Poisson regression: classical GLM, estimated by maximum likelihood (ML)                                                                                                                      |
-|                |              | Quasi    | "Quasi-Poisson regression'': same mean function, estimated by quasi-ML (QML) or equivalently generalized estimating equations (GEE), inference adjustment via estimated dispersion parameter |
-|                |              | Adjusted | "Adjusted Poisson regression'': same mean function, estimated by QML/GEE, inference adjustment via sandwich covariances                                                                      |
-|                | NB           | ML       | NB regression: extended GLM, estimated by ML including additional shape parameter                                                                                                            |
-| Zero-augmented | Poisson      | ML       | Zero-inflated Poisson (ZIP), hurdle Poisson                                                                                                                                                  |
-|                | NB           | ML       | Zero-inflated NB (ZINB), hurdle NB                                                                                                                                                           |
-: Overview of various count regression models.
-The table is usually placed at the top of the page (`[t!]`), centered (`centering`), has a caption below the table, column headers and captions are in sentence style, and if possible vertical lines should be avoided.
-{#tbl-overview}
-
-## Illustrations {#sec-illustrations}
-
-For a simple illustration of basic Poisson and NB count regression the
-`quine` data from the [MASS]{.pkg} package is used. This provides the number
-of `Days` that children were absent from school in Australia in a
-particular year, along with several covariates that can be employed as regressors. The data can be loaded by
-
-```{R}
-#| prompt: true
-data("quine", package = "MASS")
-```
-and a basic frequency distribution of the response variable is displayed in
-@fig-quine.
-
-:::{.callout}
-For code input and output, the style files provide dedicated environments.
-Either the "agnostic" `{CodeInput}` and `{CodeOutput}` can be used
-or, equivalently, the environments `{Sinput}` and `{Soutput}` as
-produced by [Sweave]{.fct} or [knitr]{.pkg} when using the `render_sweave()`
-hook. Please make sure that all code is properly spaced, e.g., using
-`y = a + b * x` and _not_ `y=a+b*x`. Moreover, code input should use "the usual" command prompt in the respective software system. For
-[R]{.proglang} code, the prompt `R> ` should be used with `+  ` as
-the continuation prompt. Generally, comments within the code chunks should be
-avoided -- and made in the regular {{< latex >}} text instead. Finally, empty lines before and after code input/output should be avoided (see above).
-:::
-
-::: {#fig-quine}
-
-![](article-visualization.pdf)
-
-Frequency distribution for number of days absent from school.
-:::
-
-As a first model for the `quine` data, we fit the basic Poisson regression
-model. (Note that JSS prefers when the second line of code is indented by two
-spaces.)
-
-```{R}
-#| prompt: true
-m_pois <- glm(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine, family = poisson)
-```
-
-To account for potential overdispersion we also consider a negative binomial
-GLM.
-
-```{R}
-#| prompt: true
-library("MASS")
-m_nbin <- glm.nb(Days ~ (Eth + Sex + Age + Lrn)^2, data = quine)
-```
-
-In a comparison with the BIC the latter model is clearly preferred.
-
-```{R}
-#| prompt: true
-library("MASS")
-BIC(m_pois, m_nbin)
-```
-
-Hence, the full summary of that model is shown below.
-
-```{R}
-#| prompt: true
-summary(m_nbin)
-```
-
-## Summary and discussion {#sec-summary}
-
-:::{.callout}
-
-As usual…
-
-:::
-
-## Computational details {.unnumbered}
-
-:::{.callout}
-
-If necessary or useful, information about certain computational details
-such as version numbers, operating systems, or compilers could be included
-in an unnumbered section. Also, auxiliary packages (say, for visualizations,
-maps, tables, …) that are not cited in the main text can be credited here.
-
-:::
-
-The results in this paper were obtained using [R]{.proglang}~3.4.1 with the
-[MASS]{.pkg}~7.3.47 package. [R]{.proglang} itself and all packages used are available from the Comprehensive [R]{.proglang} Archive Network (CRAN) at
-[https://CRAN.R-project.org/].
-
-## Acknowledgments {.unnumbered}
-
-:::{.callout}
-
-All acknowledgments (note the AE spelling) should be collected in this
-unnumbered section before the references. It may contain the usual information
-about funding and feedback from colleagues/reviewers/etc. Furthermore,
-information such as relative contributions of the authors may be added here
-(if any).
-
-:::
-
-## References {.unnumbered}
-
-:::{#refs}
-
-:::
-
-{{< pagebreak >}}
-
-## More technical details {#sec-techdetails .unnumbered}
-
-:::{.callout}
-
-Appendices can be included after the bibliography (with a page break). Each
-section within the appendix should have a proper section title (rather than
-just _Appendix_).
-
-For more technical style details, please check out JSS's style FAQ at
-[https://www.jstatsoft.org/pages/view/style#frequently-asked-questions]
-which includes the following topics:
-
-- Title vs. sentence case.
-- Graphics formatting.
-- Naming conventions.
-- Turning JSS manuscripts into [R]{.proglang} package vignettes.
-- Trouble shooting.
-- Many other potentially helpful details…
-
-:::
-
-## Using BibTeX {#sec-bibtex .unnumbered}
-
-:::{.callout}
-
-References need to be provided in a {{< bibtex >}} file (`.bib`). All
-references should be made with `@cite` syntax. This commands yield different
-formats of author-year citations and allow to include additional details (e.g.,pages, chapters, \dots) in brackets. In case you are not familiar with these
-commands see the JSS style FAQ for details.
-
-Cleaning up {{< bibtex >}} files is a somewhat tedious task -- especially
-when acquiring the entries automatically from mixed online sources. However,
-it is important that informations are complete and presented in a consistent
-style to avoid confusions. JSS requires the following format.
-
-- item JSS-specific markup (`\proglang`, `\pkg`, `\code`) should be used in the references.
-- item Titles should be in title case.
-- item Journal titles should not be abbreviated and in title case.
-- item DOIs should be included where available.
-- item Software should be properly cited as well. For [R]{.proglang} packages `citation("pkgname")` typically provides a good starting point.
-
-::: -->

### 2025-03-07 paper/paper.qmd
Update with flows to school

@@
 title: "simodels: A package for Spatial Interaction Modelling"
+bibliography: bibliography.bib
+number-sections: true
+execute: 
+  messages: false
+  warning: false 
+  echo: false
 ---
 
+```{r}
+#| eval: false
+#| echo: false
+# Compile to docx and jss-pdf formats:
+quarto::quarto_render("paper/paper.qmd", output_format = "docx")
+quarto::quarto_render("paper/paper.qmd", output_format = "jss-pdf")
+system("gh release upload v0.1.0 paper/paper.docx --clobber")
+system("gh release upload v0.1.0 paper/paper.pdf --clobber")
+```
+
 <!-- 
-# Note: you need to have the relevant quarto extension installed:
-quarto add quarto-journals/jss
-After that put the following in the yaml:
+# Note: you need to have the relevant quarto extension installed for the JSS version:
+# quarto add quarto-journals/jss
+# After that put the following in the yaml:
 format:
@@
 
-keywords: [spatial interaction models, R, transportation modelling, gravity models, origin-destination data]
-keywords-formatted: [spatial interaction models, "[R]{.proglang}", transportation modelling, gravity models, origin-destination data]
+keywords: [JSS, style guide, comma-separated, not capitalized, R]
+keywords-formatted: [JSS, style guide, comma-separated, not capitalized, "[R]{.proglang}"]
+-->
 
-bibliography: bibliography.bib
-execute: 
-  messages: false
-  warning: false 
-  echo: false
--->
+## Ideas 
+xx
+
+- Conceptual diagram showing decision points when developing SIMs, e.g.:
+  - Unipartite or bipartite?
+  - Unconstrained, production-constrained, attraction-constrained, or doubly-constrained? Or other?
+  - Calculation and parameter optimisation or regression model?
+- Open access dataset?
 
